---
title: "Cleaning the Dataset"
---
Import the data.
```{r}
#for reading dta
library(haven)
#for easier file navigation
library(here)

data17_18_fv <- read_dta(here("Data", "17-18", "hh_per_fv_2017-18.dta"))

data18_19_fv <- read_dta(here("Data","18-19","PerV1_2018-19.dta"))

data19_20_fv <- read_dta(here("Data","19-20","PERFV_2019-20.dta"))

data20_21_fv <- read_dta(here("Data","20-21","perv1.dta"))

data21_22_fv <- readr::read_csv(here("Data","21-22","perv1.csv"))

data22_23 <- read_dta(here("Data","22-23","perv1 (2).dta"))

data23_24 <- read_dta(here("Data","23-24","perv1.dta"))

data24 <- read_dta(here("Data","24","cperv1.dta"))

```

Creating analysis dataframe:
```{r}
library(dplyr)
analysis17_18 <- data17_18_fv
analysis18_19 <- data18_19_fv
analysis19_20 <- data19_20_fv
analysis20_21 <- data20_21_fv
analysis21_22 <- data21_22_fv
analysis22_23 <- data22_23
analysis23_24 <- data23_24
analysis24 <- data24

#Zapping stata labels because they get in the way of dataframe manipulation
analysis17_18 <- analysis17_18 %>% mutate(across(where(haven::is.labelled), haven::zap_labels))
analysis18_19 <- analysis18_19 %>% mutate(across(where(haven::is.labelled), haven::zap_labels))
analysis19_20 <- analysis19_20 %>% mutate(across(where(haven::is.labelled), haven::zap_labels))
analysis20_21 <- analysis20_21 %>% mutate(across(where(haven::is.labelled), haven::zap_labels))
analysis21_22 <- analysis21_22 %>% mutate(across(where(haven::is.labelled), haven::zap_labels))
analysis22_23 <- analysis22_23 %>% mutate(across(where(haven::is.labelled), haven::zap_labels))
analysis23_24 <- analysis23_24 %>% mutate(across(where(haven::is.labelled), haven::zap_labels))
analysis24 <- analysis24 %>% mutate(across(where(haven::is.labelled), haven::zap_labels))

#get rid of "data" so I don't even accidentally write over actual dataset
rm(list=c("data17_18_fv","data18_19_fv","data19_20_fv","data20_21_fv","data21_22_fv","data22_23","data23_24","data24","data25"))
```

Matching quarters in 2024 data to consolidate the two:
```{r}
aggregate(data.frame(count = analysis22_23$qtr_perv1), list(value = analysis22_23$qtr_perv1), length)

aggregate(data.frame(count = analysis23_24$qtr_perv1), list(value = analysis23_24$qtr_perv1), length)

aggregate(data.frame(count = analysis24$Quarter), list(value = analysis24$Quarter), length)

#Task now is to add the Q1 and Q2 from 23-24 into the 22 data, so it becomes 22 to end of 23 data.Then, get rid of 23_24 because Q3 and Q4 from 23_24 is already in 24.

Q1Q2_2324 <- analysis23_24 %>% filter(qtr_perv1 %in% c("Q1","Q2"))
analysis22Jul_23Dec <- rbind(analysis22_23,Q1Q2_2324)
rm(analysis22_23)
rm(analysis23_24)
rm(Q1Q2_2324)

#Before this, I checked that the columns/variables recorded only differ by 1 and that is panel. Given observations in 23_24 and 24 have identical 24 components, no need to correct for those. 22_23 and 23_24 also had identical columns, so no need to check.
```

1.Informal and Transport binaries.
1.1Matching the NIC and NCO columns:
```{r}
NIC17 <- analysis17_18$b5pt1q5_per_fv
NCO17 <- analysis17_18$b5pt1q6_per_fv

NIC18 <- analysis18_19$b5pt1q5_per_fv
NCO18 <- analysis18_19$b5pt1q6_per_fv

NIC19 <- analysis19_20$b5pt1q5_per_fv
NCO19 <- analysis19_20$b5pt1q6_per_fv

NIC20 <- analysis20_21$b5pt1q5_perv1
NCO20 <- analysis20_21$b5pt1q6_perv1

NIC21 <- analysis21_22$b5pt1q5_perv1
NCO21 <- analysis21_22$b5pt1q6_perv1

NIC22 <- analysis22Jul_23Dec$b5pt1q5_perv1
NCO22 <- analysis22Jul_23Dec$b5pt1q6_perv1

NIC24 <- analysis24$Principal_Industry_Code
NCO24 <- analysis24$Principal_Occupation_Code
```

1.2 Checking which NIC values are present and whether there are empty rows:
```{r}
aggregate(data.frame(count = NIC19), list(value = NIC19), length)
```
#We find that all NIC values in all data frames are 6 digits long.

1.3 Creating transport binary:
```{r}
library(dplyr)


analysis17_18$transport <- ifelse(
  NIC17 %in% c("49110", "49211","49212","49213","49219",
               "49222","49223","49224","49226","49229" ),
  1,
  0
)

analysis18_19$transport <- ifelse(
  NIC18 %in% c("49110", "49211","49212","49213","49219",
               "49222","49223","49224","49226","49229" ),
  1,
  0
)

analysis19_20$transport <- ifelse(
  NIC19 %in% c("49110", "49211","49212","49213","49219",
               "49222","49223","49224","49226","49229" ),
  1,
  0
)

analysis20_21$transport <- ifelse(
  NIC20 %in% c("49110", "49211","49212","49213","49219",
               "49222","49223","49224","49226","49229" ),
  1,
  0
)

analysis21_22$transport <- ifelse(
  NIC21 %in% c("49110", "49211","49212","49213","49219",
               "49222","49223","49224","49226","49229"  ),
  1,
  0
)

analysis22Jul_23Dec$transport <- ifelse(
  NIC22 %in% c("49110", "49211","49212","49213","49219",
               "49222","49223","49224","49226","49229" ),
  1,
  0
)

analysis24$transport <- ifelse(
  NIC24 %in% c("49110", "49211","49212","49213","49219",
               "49222","49223","49224","49226","49229" ),
  1,
  0
)

sum(analysis17_18$transport == 1, na.rm = TRUE) #4924
sum(analysis18_19$transport==1, na.rm=TRUE) #4727
sum(analysis19_20$transport==1, na.rm=TRUE) #4691
sum(analysis20_21$transport==1, na.rm=TRUE) #4719
sum(analysis21_22$transport==1, na.rm=TRUE) #4604
sum(analysis22Jul_23Dec$transport==1, na.rm=TRUE) #7581
sum(analysis24$transport==1, na.rm=TRUE) #5169

```

**1.4 Checking how many NCO codes are there for transport per year
NCO-2004 used from 2017 until 2021
NCO-2015 used from 2022 to 2024
#Will look at these and check if share of drivers changed later
```{r}

NCO17_transport <- NCO17[analysis17_18$transport == 1]

aggregate(
  data.frame(count = NCO17_transport),
  list(value = NCO17_transport),
  length
)

#NCO codes 980 and 990 do not fit the documentation, and make up 0.01% of entries. Treat as missing/ignore



NCO18_transport <- NCO18[analysis18_19$transport == 1]

aggregate(
  data.frame(count = NCO18_transport),
  list(value = NCO18_transport),
  length
)


NCO19_transport <- NCO19[analysis19_20$transport == 1]

aggregate(
  data.frame(count = NCO19_transport),
  list(value = NCO19_transport),
  length
)


NCO20_transport <- NCO20[analysis20_21$transport == 1]

aggregate(
  data.frame(count = NCO20_transport),
  list(value = NCO20_transport),
  length
)

NCO21_transport <- NCO21[analysis21_22$transport == 1]

aggregate(
  data.frame(count = NCO21_transport),
  list(value = NCO21_transport),
  length
)

NCO22_transport <- NCO22[analysis22Jul_23Dec$transport == 1]

aggregate(
  data.frame(count = NCO22_transport),
  list(value = NCO22_transport),
  length
)

NCO24_transport <- NCO24[analysis24$transport == 1]

aggregate(
  data.frame(count = NCO24_transport),
  list(value = NCO24_transport),
  length
)
```

1.5 Creating informal binary
```{r}
# STATUS = 11 or 12
# │
# ├── Enterprise Type = 01, 02, 03, 04
# │       → INFORMAL = 1
# │
# ├── Enterprise Type = clearly non-informal category
# │       → INFORMAL = 0
# │
# └── Enterprise Type missing / ambiguous
#         → INFORMAL = NA
# 
# 
# STATUS = 21
# │
# └── INFORMAL = 1
# 
# 
# STATUS = 31, 41 or 51
# │
# ├── Social Security = 8
# │       → INFORMAL = 1
# │
# ├── Social Security = 1–7
# │       → INFORMAL = 0
# │
# └── Social Security = 9 / missing
#         → INFORMAL = NA
# 
# 
# ALL OTHER STATUS CODES
# │
# └── INFORMAL = NA
#     (not an employed worker covered by this definition)


analysis17_18 <- analysis17_18 |>
  dplyr::mutate(
    informal = dplyr::case_when(

      # Own-account workers / employers in informal-sector enterprises
      b5pt1q3_per_fv %in% c("11", "12") &
        b5pt1q9_per_fv %in% c("01", "02", "03", "04") ~ 1,

      # Own-account workers / employers in clearly non-informal enterprises
      b5pt1q3_per_fv %in% c("11", "12") &
        !is.na(b5pt1q9_per_fv) &
        b5pt1q9_per_fv != "" &
        !b5pt1q9_per_fv %in% c("01", "02", "03", "04", "19") ~ 0,

      # Contributing/unpaid family workers
      b5pt1q3_per_fv == "21" ~ 1,

      # Employees with no employer-linked social security
      b5pt1q3_per_fv %in% c("31", "41", "51") &
        b5pt1q13_per_fv == "8" ~ 1,

      # Employees with some social-security coverage
      b5pt1q3_per_fv %in% c("31", "41", "51") &
        b5pt1q13_per_fv %in% as.character(1:7) ~ 0,

      # Unknown / not classifiable / non-workers
      TRUE ~ NA_real_
    )
  )

analysis18_19 <- analysis18_19 |>
  dplyr::mutate(
    informal = dplyr::case_when(

      # Own-account workers / employers in informal-sector enterprises
      b5pt1q3_per_fv %in% c("11", "12") &
        b5pt1q9_per_fv %in% c("01", "02", "03", "04") ~ 1,

      # Own-account workers / employers in clearly non-informal enterprises
      b5pt1q3_per_fv %in% c("11", "12") &
        !is.na(b5pt1q9_per_fv) &
        b5pt1q9_per_fv != "" &
        !b5pt1q9_per_fv %in% c("01", "02", "03", "04", "19") ~ 0,

      # Contributing/unpaid family workers
      b5pt1q3_per_fv == "21" ~ 1,

      # Employees with no employer-linked social security
      b5pt1q3_per_fv %in% c("31", "41", "51") &
        b5pt1q13_per_fv == "8" ~ 1,

      # Employees with some social-security coverage
      b5pt1q3_per_fv %in% c("31", "41", "51") &
        b5pt1q13_per_fv %in% as.character(1:7) ~ 0,

      # Unknown / not classifiable / non-workers
      TRUE ~ NA_real_
    )
  )

analysis19_20 <- analysis19_20 |>
  dplyr::mutate(
    informal = dplyr::case_when(

      # Own-account workers / employers in informal-sector enterprises
      b5pt1q3_per_fv %in% c("11", "12") &
        b5pt1q9_per_fv %in% c("01", "02", "03", "04") ~ 1,

      # Own-account workers / employers in clearly non-informal enterprises
      b5pt1q3_per_fv %in% c("11", "12") &
        !is.na(b5pt1q9_per_fv) &
        b5pt1q9_per_fv != "" &
        !b5pt1q9_per_fv %in% c("01", "02", "03", "04", "19") ~ 0,

      # Contributing/unpaid family workers
      b5pt1q3_per_fv == "21" ~ 1,

      # Employees with no employer-linked social security
      b5pt1q3_per_fv %in% c("31", "41", "51") &
        b5pt1q13_per_fv == "8" ~ 1,

      # Employees with some social-security coverage
      b5pt1q3_per_fv %in% c("31", "41", "51") &
        b5pt1q13_per_fv %in% as.character(1:7) ~ 0,

      # Unknown / not classifiable / non-workers
      TRUE ~ NA_real_
    )
  )

analysis20_21 <- analysis20_21 |>
  dplyr::mutate(
    informal = dplyr::case_when(

      # Own-account workers / employers in informal-sector enterprises
      b5pt1q3_perv1 %in% c("11", "12") &
        b5pt1q9_perv1 %in% c("01", "02", "03", "04") ~ 1,

      # Own-account workers / employers in clearly non-informal enterprises
      b5pt1q3_perv1 %in% c("11", "12") &
        !is.na(b5pt1q9_perv1) &
        b5pt1q9_perv1 != "" &
        !b5pt1q9_perv1 %in% c("01", "02", "03", "04", "19") ~ 0,

      # Contributing/unpaid family workers
      b5pt1q3_perv1 == "21" ~ 1,

      # Employees with no employer-linked social security
      b5pt1q3_perv1 %in% c("31", "41", "51") &
        b5pt1q13_perv1 == "8" ~ 1,

      # Employees with some social-security coverage
      b5pt1q3_perv1 %in% c("31", "41", "51") &
        b5pt1q13_perv1 %in% as.character(1:7) ~ 0,

      # Unknown / not classifiable / non-workers
      TRUE ~ NA_real_
    )
  )

analysis21_22 <- analysis21_22 |>
  dplyr::mutate(
    informal = dplyr::case_when(

      # Own-account workers / employers in informal-sector enterprises
      b5pt1q3_perv1 %in% c("11", "12") &
        b5pt1q9_perv1 %in% c("01", "02", "03", "04") ~ 1,

      # Own-account workers / employers in clearly non-informal enterprises
      b5pt1q3_perv1 %in% c("11", "12") &
        !is.na(b5pt1q9_perv1) &
        b5pt1q9_perv1 != "" &
        !b5pt1q9_perv1 %in% c("01", "02", "03", "04", "19") ~ 0,

      # Contributing/unpaid family workers
      b5pt1q3_perv1 == "21" ~ 1,

      # Employees with no employer-linked social security
      b5pt1q3_perv1 %in% c("31", "41", "51") &
        b5pt1q13_perv1 == "8" ~ 1,

      # Employees with some social-security coverage
      b5pt1q3_perv1 %in% c("31", "41", "51") &
        b5pt1q13_perv1 %in% as.character(1:7) ~ 0,

      # Unknown / not classifiable / non-workers
      TRUE ~ NA_real_
    )
  )

analysis22Jul_23Dec <- analysis22Jul_23Dec |>
  dplyr::mutate(
    informal = dplyr::case_when(

      # Own-account workers / employers in informal-sector enterprises
      b5pt1q3_perv1 %in% c("11", "12") &
        b5pt1q9_perv1 %in% c("01", "02", "03", "04") ~ 1,

      # Own-account workers / employers in clearly non-informal enterprises
      b5pt1q3_perv1 %in% c("11", "12") &
        !is.na(b5pt1q9_perv1) &
        b5pt1q9_perv1 != "" &
        !b5pt1q9_perv1 %in% c("01", "02", "03", "04", "19") ~ 0,

      # Contributing/unpaid family workers
      b5pt1q3_perv1 == "21" ~ 1,

      # Employees with no employer-linked social security
      b5pt1q3_perv1 %in% c("31", "41", "51") &
        b5pt1q13_perv1 == "8" ~ 1,

      # Employees with some social-security coverage
      b5pt1q3_perv1 %in% c("31", "41", "51") &
        b5pt1q13_perv1 %in% as.character(1:7) ~ 0,

      # Unknown / not classifiable / non-workers
      TRUE ~ NA_real_
    )
  )

analysis24 <- analysis24 |>
  dplyr::mutate(
    informal = dplyr::case_when(

      # Own-account workers / employers in informal-sector enterprises
      Principal_Status_Code %in% c("11", "12") &
        Principal_Enterprise_Type %in% c("01", "02", "03", "04") ~ 1,

      # Own-account workers / employers in clearly non-informal enterprises
      Principal_Status_Code %in% c("11", "12") &
        !is.na(Principal_Enterprise_Type) &
        Principal_Enterprise_Type != "" &
        !Principal_Enterprise_Type %in% c("01", "02", "03", "04", "19") ~ 0,

      # Contributing/unpaid family workers
      Principal_Status_Code == "21" ~ 1,

      # Employees with no employer-linked social security
      Principal_Status_Code %in% c("31", "41", "51") &
        Principal_Social_Security == "8" ~ 1,

      # Employees with some social-security coverage
      Principal_Status_Code %in% c("31", "41", "51") &
        Principal_Social_Security %in% as.character(1:7) ~ 0,

      # Unknown / not classifiable / non-workers
      TRUE ~ NA_real_
    )
  )

sum(analysis17_18$informal=="1", na.rm = TRUE)

sum(analysis18_19$informal=="1", na.rm = TRUE)

sum(analysis19_20$informal=="1", na.rm = TRUE)

sum(analysis20_21$informal=="1", na.rm = TRUE)

sum(analysis21_22$informal=="1", na.rm = TRUE)

sum(analysis22Jul_23Dec$informal=="1", na.rm = TRUE)

sum(analysis24$informal=="1", na.rm = TRUE)
```


2.Creating periods based on quarter and year
```{r}
analysis17_18$period <- match(
  analysis17_18$quarter_per_fv,
  c("Q1", "Q2", "Q3", "Q4")
)

analysis18_19$period <- match(
  analysis18_19$quarter_per_fv,
  c("Q5", "Q6", "Q7", "Q8")
) + 4

analysis19_20$period <- match(
  analysis19_20$quarter_per_fv,
  c("Q1", "Q2", "Q3", "Q4")
) + 8

analysis20_21$period <- match(
  analysis20_21$qtr_perv1,
  c("Q5", "Q6", "Q7", "Q8")
) + 12

analysis21_22$period <- match(
  analysis21_22$qtr_perv1,
  c("Q1", "Q2", "Q3", "Q4")
) + 16

analysis22Jul_23Dec$period <- match(
  analysis22Jul_23Dec$qtr_perv1,
  c("Q5", "Q6", "Q7", "Q8", "Q1", "Q2")
) + 20

analysis24$period <- match(
  analysis24$Quarter, 
  c("Q3","Q4","Q5","Q6")
) +26



aggregate(data.frame(count = analysis17_18$period), list(value = analysis17_18$period), length)
aggregate(data.frame(count = analysis18_19$period), list(value = analysis18_19$period), length)
aggregate(data.frame(count = analysis19_20$period), list(value = analysis19_20$period), length)
aggregate(data.frame(count = analysis20_21$period), list(value = analysis20_21$period), length)
aggregate(data.frame(count = analysis21_22$period), list(value = analysis21_22$period), length)
aggregate(data.frame(count = analysis22Jul_23Dec$period), list(value = analysis22Jul_23Dec$period), length)
aggregate(data.frame(count = analysis24$period), list(value = analysis24$period), length)

#period successfully made
```

3.Treatment variable
```{r}
# For:
#   Q1 2019 and onwards for state:24 District:7 
#   Q1 2019 and onwards for state:27 district:9
#   Q4 2021 and onwards for state:9 district: 33
#   Q1 2022 and onwards for state:27 district:25
# Treatment is 1

#Adding stratum to limit treatment to cities, only cities with stratum>4 in each treatment district has the metro, so better identification. i.e no different city is identified.

analysis17_18$Treatment <- 0
analysis18_19$Treatment <- 0
analysis19_20$Treatment <- 0
analysis20_21$Treatment <- 0
analysis21_22$Treatment <- 0
analysis22Jul_23Dec$Treatment <- 0
analysis24$Treatment <- 0

# ================================================================
# 3. TREATMENT VARIABLE
# ================================================================

library(dplyr)


# ----------------------------------------------------------------
# Helper function for harmonised metro treatment assignment
#
# IMPORTANT:
# - 2017-18 to 2020-21: city-size stratum is stored in b1q6
# - 2021-22 to 2024: city-size stratum is stored in b1q5
# - Urban strata >= 4 represent individual Census-2011 cities
#   with populations of at least 15 lakh.
#
# Primary treatment timing uses the first fully exposed quarter:
#   Ahmedabad: period 8
#   Nagpur:    period 8
#   Kanpur:    period 19
#   Pune/PCMC: period 20
#
# The opening-quarter specification is retained for sensitivity:
#   Ahmedabad: period 7
#   Nagpur:    period 7
#   Kanpur:    period 18
#   Pune/PCMC: period 19
# ----------------------------------------------------------------

add_metro_treatment <- function(df,
                                urban_var,
                                city_stratum_var,
                                state_var,
                                district_var) {

  required_variables <- c(
    urban_var,
    city_stratum_var,
    state_var,
    district_var,
    "period"
  )

  missing_variables <- setdiff(required_variables, names(df))

  if (length(missing_variables) > 0) {
    stop(
      "The following variables are missing: ",
      paste(missing_variables, collapse = ", ")
    )
  }


  df %>%

    mutate(

      # ----------------------------------------------------------
      # Standardised versions of the raw geography variables
      # ----------------------------------------------------------

      .urban_code = suppressWarnings(
        as.integer(as.character(.data[[urban_var]]))
      ),

      .state_code_treatment = suppressWarnings(
        as.integer(as.character(.data[[state_var]]))
      ),

      .district_code_treatment = suppressWarnings(
        as.integer(as.character(.data[[district_var]]))
      ),


      # ----------------------------------------------------------
      # Harmonised Census-2011 city-size stratum
      #
      # The appropriate raw field is supplied separately for
      # each survey wave below.
      # ----------------------------------------------------------

      city_size_stratum = suppressWarnings(
        as.integer(as.character(.data[[city_stratum_var]]))
      ),


      # ----------------------------------------------------------
      # Comparable large-city analysis sample
      #
      # This identifies urban city strata with a Census-2011
      # population of at least 15 lakh.
      # ----------------------------------------------------------

      eligible_city = as.integer(
        .urban_code == 2L &
          !is.na(city_size_stratum) &
          city_size_stratum >= 4L
      ),


      # ----------------------------------------------------------
      # Metro system label
      #
      # Pune includes the Pune/Pimpri-Chinchwad urban system.
      # ----------------------------------------------------------

      metro_system = case_when(

        .state_code_treatment == 24L &
          .district_code_treatment == 7L ~
          "Ahmedabad",

        .state_code_treatment == 27L &
          .district_code_treatment == 9L ~
          "Nagpur",

        .state_code_treatment == 9L &
          .district_code_treatment == 33L ~
          "Kanpur",

        .state_code_treatment == 27L &
          .district_code_treatment == 25L ~
          "Pune-PCMC",

        TRUE ~ NA_character_
      ),


      # ----------------------------------------------------------
      # PRIMARY COHORT:
      # first fully exposed quarter
      #
      # 0 = never treated
      # 8 = Ahmedabad and Nagpur: April-June 2019
      # 19 = Kanpur: January-March 2022
      # 20 = Pune/PCMC: April-June 2022
      #
      # Cohort assignment depends only on the metro district.
      # The eligible-city restriction is applied separately.
      # ----------------------------------------------------------

      first_treat = case_when(

        .state_code_treatment == 24L &
          .district_code_treatment == 7L ~ 8L,

        .state_code_treatment == 27L &
          .district_code_treatment == 9L ~ 8L,

        .state_code_treatment == 9L &
          .district_code_treatment == 33L ~ 19L,

        .state_code_treatment == 27L &
          .district_code_treatment == 25L ~ 20L,

        TRUE ~ 0L
      ),


      # ----------------------------------------------------------
      # SENSITIVITY COHORT:
      # quarter in which the metro opened
      #
      # 7 = Ahmedabad and Nagpur
      # 18 = Kanpur
      # 19 = Pune/PCMC
      # ----------------------------------------------------------

      first_treat_opening_quarter = case_when(

        .state_code_treatment == 24L &
          .district_code_treatment == 7L ~ 7L,

        .state_code_treatment == 27L &
          .district_code_treatment == 9L ~ 7L,

        .state_code_treatment == 9L &
          .district_code_treatment == 33L ~ 18L,

        .state_code_treatment == 27L &
          .district_code_treatment == 25L ~ 19L,

        TRUE ~ 0L
      ),


      # ----------------------------------------------------------
      # PRIMARY BINARY TREATMENT
      #
      # Smaller towns and rural areas inside the four districts
      # are not treated in this city-level specification.
      # ----------------------------------------------------------

      Treatment = case_when(

        is.na(period) ~ NA_integer_,

        eligible_city == 1L &
          first_treat > 0L &
          period >= first_treat ~ 1L,

        TRUE ~ 0L
      ),


      # ----------------------------------------------------------
      # OPENING-QUARTER SENSITIVITY TREATMENT
      # ----------------------------------------------------------

      Treatment_opening_quarter = case_when(

        is.na(period) ~ NA_integer_,

        eligible_city == 1L &
          first_treat_opening_quarter > 0L &
          period >= first_treat_opening_quarter ~ 1L,

        TRUE ~ 0L
      )

    ) %>%

    # Remove temporary harmonisation variables
    select(
      -.urban_code,
      -.state_code_treatment,
      -.district_code_treatment
    )
}


# ================================================================
# APPLY THE TREATMENT FUNCTION TO EACH SURVEY WAVE
# ================================================================


# 2017-18
# In the early files, b1q6 contains the effective urban
# city-size stratum.
analysis17_18 <- add_metro_treatment(
  df                = analysis17_18,
  urban_var         = "b1q3_per_fv",
  city_stratum_var  = "b1q6_per_fv",
  state_var         = "state_per_fv",
  district_var      = "b1q4_per_fv"
)


# 2018-19
analysis18_19 <- add_metro_treatment(
  df                = analysis18_19,
  urban_var         = "b1q3_per_fv",
  city_stratum_var  = "b1q6_per_fv",
  state_var         = "state_per_fv",
  district_var      = "b1q4_per_fv"
)


# 2019-20
analysis19_20 <- add_metro_treatment(
  df                = analysis19_20,
  urban_var         = "b1q3_per_fv",
  city_stratum_var  = "b1q6_per_fv",
  state_var         = "state_per_fv",
  district_var      = "b1q4_per_fv"
)


# 2020-21
analysis20_21 <- add_metro_treatment(
  df                = analysis20_21,
  urban_var         = "b1q3_perv1",
  city_stratum_var  = "b1q6_perv1",
  state_var         = "state_perv1",
  district_var      = "distcode_perv1"
)


# 2021-22
# From this wave onward, b1q5 contains the urban
# city-size stratum.
analysis21_22 <- add_metro_treatment(
  df                = analysis21_22,
  urban_var         = "b1q3_perv1",
  city_stratum_var  = "b1q5_perv1",
  state_var         = "state_perv1",
  district_var      = "distcode_perv1"
)


# July 2022 to December 2023
analysis22Jul_23Dec <- add_metro_treatment(
  df                = analysis22Jul_23Dec,
  urban_var         = "b1q3_perv1",
  city_stratum_var  = "b1q5_perv1",
  state_var         = "state_perv1",
  district_var      = "distcode_perv1"
)


# Calendar year 2024
analysis24 <- add_metro_treatment(
  df                = analysis24,
  urban_var         = "Sector",
  city_stratum_var  = "Stratum",
  state_var         = "State_UT_Code",
  district_var      = "District_Code"
)


# ================================================================
# BASIC VALIDATION CHECKS
# ================================================================

treatment_dataframes <- list(
  "2017-18" = analysis17_18,
  "2018-19" = analysis18_19,
  "2019-20" = analysis19_20,
  "2020-21" = analysis20_21,
  "2021-22" = analysis21_22,
  "2022-23" = analysis22Jul_23Dec,
  "2024"    = analysis24
)


# Confirm that treatment is never assigned outside the eligible
# urban large-city sample.
invisible(
  lapply(
    treatment_dataframes,
    function(x) {

      stopifnot(
        !any(
          x$Treatment == 1L &
            x$eligible_city != 1L,
          na.rm = TRUE
        )
      )

      stopifnot(
        !any(
          x$Treatment_opening_quarter == 1L &
            x$eligible_city != 1L,
          na.rm = TRUE
        )
      )
    }
  )
)


# Unweighted person-row diagnostic.
# These counts are only cleaning checks; they are not estimates.
treatment_diagnostic <- bind_rows(
  lapply(
    names(treatment_dataframes),
    function(wave_name) {

      x <- treatment_dataframes[[wave_name]]

      tibble(
        survey_wave = wave_name,
        total_rows = nrow(x),
        eligible_large_city_rows =
          sum(x$eligible_city == 1L, na.rm = TRUE),
        primary_treated_rows =
          sum(x$Treatment == 1L, na.rm = TRUE),
        opening_quarter_treated_rows =
          sum(x$Treatment_opening_quarter == 1L, na.rm = TRUE)
      )
    }
  )
)

print(treatment_diagnostic)


# ================================================================
# IMPORTANT FOR THE FINAL DiD DATA
# ================================================================
#
# Do not filter the analysis dataframes at this point because the
# survey-weight calculations later in the cleaning file require the
# complete survey sample.
#
# After the weights have been constructed, the final DiD dataset
# must be restricted as follows:
#
#     filter(eligible_city == 1L)
#
# Section 8 should retain the variables created here:
#
#     city_size_stratum
#     eligible_city
#     metro_system
#     first_treat
#     first_treat_opening_quarter
#     Treatment
#     Treatment_opening_quarter
#
# It should not recreate city eligibility or first_treat from the
# original b1q5 variable.
#
# Also remember that the informal-employment outcome uses usual
# principal status and therefore refers to the preceding 365 days.
# The early post-treatment quarters should consequently be
# interpreted as transition periods.
# ================================================================


sum(analysis17_18$Treatment=="1" & analysis17_18$transport=="1" & analysis17_18$informal=="1",na.rm=TRUE)
sum(analysis18_19$Treatment=="1" & analysis18_19$transport=="1" & analysis18_19$informal=="1",na.rm=TRUE)
sum(analysis19_20$Treatment=="1" & analysis19_20$transport=="1" & analysis19_20$informal=="1",na.rm=TRUE)
sum(analysis20_21$Treatment=="1" & analysis20_21$transport=="1" & analysis20_21$informal=="1",na.rm=TRUE)
sum(analysis21_22$Treatment=="1" & analysis21_22$transport=="1" & analysis21_22$informal=="1",na.rm=TRUE)
sum(analysis22Jul_23Dec$Treatment=="1" & analysis22Jul_23Dec$transport=="1" & analysis22Jul_23Dec$informal=="1",na.rm=TRUE)
sum(analysis24$Treatment=="1" & analysis24$transport=="1" & analysis24$informal=="1",na.rm=TRUE)


```

4.Already Treated Variables
```{r}
#2017

analysis17_18$NotControl <- 0

# Ernakulam
analysis17_18$NotControl[
  analysis17_18$state_per_fv == "32" &
  analysis17_18$b1q4_per_fv == "8"
] <- 1

# Chennai
analysis17_18$NotControl[
  analysis17_18$state_per_fv == "33" &
  analysis17_18$b1q4_per_fv == "2"
] <- 1

# Bengaluru Urban
analysis17_18$NotControl[
  analysis17_18$state_per_fv == "29" &
  analysis17_18$b1q4_per_fv == "18"
] <- 1

# Hyderabad
analysis17_18$NotControl[
  analysis17_18$state_per_fv == "36" &
  (
    (analysis17_18$period < 9 &
       analysis17_18$b1q4_per_fv == "5") |
      
    (analysis17_18$period >= 9 &
       analysis17_18$b1q4_per_fv == "22")
  )
] <- 1

# Mumbai Suburban
analysis17_18$NotControl[
  analysis17_18$state_per_fv == "27" &
  analysis17_18$b1q4_per_fv == "22"
] <- 1

# Kolkata
analysis17_18$NotControl[
  analysis17_18$state_per_fv == "19" &
  analysis17_18$b1q4_per_fv == "16"
] <- 1

# Jaipur
analysis17_18$NotControl[
  analysis17_18$state_per_fv == "8" &
  analysis17_18$b1q4_per_fv == "12"
] <- 1

# Agra
analysis17_18$NotControl[
  analysis17_18$state_per_fv == "9" &
  analysis17_18$b1q4_per_fv == "15"
] <- 1

# Lucknow
analysis17_18$NotControl[
  analysis17_18$state_per_fv == "9" &
  analysis17_18$b1q4_per_fv == "26"
] <- 1

# All Delhi
analysis17_18$NotControl[
  analysis17_18$state_per_fv == "7"
] <- 1

# Gurugram
analysis17_18$NotControl[
  analysis17_18$state_per_fv == "6" &
  analysis17_18$b1q4_per_fv == "18"
] <- 1

# Faridabad
analysis17_18$NotControl[
  analysis17_18$state_per_fv == "6" &
  analysis17_18$b1q4_per_fv == "20"
] <- 1

# Gautam Buddha Nagar
analysis17_18$NotControl[
  analysis17_18$state_per_fv == "9" &
  analysis17_18$b1q4_per_fv == "10"
] <- 1

# Ghaziabad
analysis17_18$NotControl[
  analysis17_18$state_per_fv == "9" &
  analysis17_18$b1q4_per_fv == "9"
] <- 1

# Haryana - Jhajjar
analysis17_18$NotControl[
  analysis17_18$state_per_fv == "6" &
  analysis17_18$b1q4_per_fv == "15"
] <- 1



#2018
analysis18_19$NotControl <- 0

# Ernakulam
analysis18_19$NotControl[
  analysis18_19$state_per_fv == "32" &
  analysis18_19$b1q4_per_fv == "8"
] <- 1

# Chennai
analysis18_19$NotControl[
  analysis18_19$state_per_fv == "33" &
  analysis18_19$b1q4_per_fv == "2"
] <- 1

# Bengaluru Urban
analysis18_19$NotControl[
  analysis18_19$state_per_fv == "29" &
  analysis18_19$b1q4_per_fv == "18"
] <- 1

# Hyderabad
analysis18_19$NotControl[
  analysis18_19$state_per_fv == "36" &
  (
    (analysis18_19$period < 9 &
       analysis18_19$b1q4_per_fv == "5") |
      
    (analysis18_19$period >= 9 &
       analysis18_19$b1q4_per_fv == "22")
  )
] <- 1

# Mumbai Suburban
analysis18_19$NotControl[
  analysis18_19$state_per_fv == "27" &
  analysis18_19$b1q4_per_fv == "22"
] <- 1

# Kolkata
analysis18_19$NotControl[
  analysis18_19$state_per_fv == "19" &
  analysis18_19$b1q4_per_fv == "16"
] <- 1

# Jaipur
analysis18_19$NotControl[
  analysis18_19$state_per_fv == "8" &
  analysis18_19$b1q4_per_fv == "12"
] <- 1

# Agra
analysis18_19$NotControl[
  analysis18_19$state_per_fv == "9" &
  analysis18_19$b1q4_per_fv == "15"
] <- 1

# Lucknow
analysis18_19$NotControl[
  analysis18_19$state_per_fv == "9" &
  analysis18_19$b1q4_per_fv == "26"
] <- 1

# All Delhi
analysis18_19$NotControl[
  analysis18_19$state_per_fv == "7"
] <- 1

# Gurugram
analysis18_19$NotControl[
  analysis18_19$state_per_fv == "6" &
  analysis18_19$b1q4_per_fv == "18"
] <- 1

# Faridabad
analysis18_19$NotControl[
  analysis18_19$state_per_fv == "6" &
  analysis18_19$b1q4_per_fv == "20"
] <- 1

# Gautam Buddha Nagar
analysis18_19$NotControl[
  analysis18_19$state_per_fv == "9" &
  analysis18_19$b1q4_per_fv == "10"
] <- 1

# Ghaziabad
analysis18_19$NotControl[
  analysis18_19$state_per_fv == "9" &
  analysis18_19$b1q4_per_fv == "9"
] <- 1

# Haryana - Jhajjar
analysis18_19$NotControl[
  analysis18_19$state_per_fv == "6" &
  analysis18_19$b1q4_per_fv == "15"
] <- 1



#2019
analysis19_20$NotControl <- 0

# Ernakulam
analysis19_20$NotControl[
  analysis19_20$state_per_fv == "32" &
  analysis19_20$b1q4_per_fv == "8"
] <- 1

# Chennai
analysis19_20$NotControl[
  analysis19_20$state_per_fv == "33" &
  analysis19_20$b1q4_per_fv == "2"
] <- 1

# Bengaluru Urban
analysis19_20$NotControl[
  analysis19_20$state_per_fv == "29" &
  analysis19_20$b1q4_per_fv == "18"
] <- 1

# Hyderabad
analysis19_20$NotControl[
  analysis19_20$state_per_fv == "36" &
  (
    (analysis19_20$period < 9 &
       analysis19_20$b1q4_per_fv == "5") |
      
    (analysis19_20$period >= 9 &
       analysis19_20$b1q4_per_fv == "22")
  )
] <- 1

# Mumbai Suburban
analysis19_20$NotControl[
  analysis19_20$state_per_fv == "27" &
  analysis19_20$b1q4_per_fv == "22"
] <- 1

# Kolkata
analysis19_20$NotControl[
  analysis19_20$state_per_fv == "19" &
  analysis19_20$b1q4_per_fv == "16"
] <- 1

# Jaipur
analysis19_20$NotControl[
  analysis19_20$state_per_fv == "8" &
  analysis19_20$b1q4_per_fv == "12"
] <- 1

# Agra
analysis19_20$NotControl[
  analysis19_20$state_per_fv == "9" &
  analysis19_20$b1q4_per_fv == "15"
] <- 1

# Lucknow
analysis19_20$NotControl[
  analysis19_20$state_per_fv == "9" &
  analysis19_20$b1q4_per_fv == "26"
] <- 1

# All Delhi
analysis19_20$NotControl[
  analysis19_20$state_per_fv == "7"
] <- 1

# Gurugram
analysis19_20$NotControl[
  analysis19_20$state_per_fv == "6" &
  analysis19_20$b1q4_per_fv == "18"
] <- 1

# Faridabad
analysis19_20$NotControl[
  analysis19_20$state_per_fv == "6" &
  analysis19_20$b1q4_per_fv == "20"
] <- 1

# Gautam Buddha Nagar
analysis19_20$NotControl[
  analysis19_20$state_per_fv == "9" &
  analysis19_20$b1q4_per_fv == "10"
] <- 1

# Ghaziabad
analysis19_20$NotControl[
  analysis19_20$state_per_fv == "9" &
  analysis19_20$b1q4_per_fv == "9"
] <- 1

# Haryana - Jhajjar
analysis19_20$NotControl[
  analysis19_20$state_per_fv == "6" &
  analysis19_20$b1q4_per_fv == "15"
] <- 1



#2020
analysis20_21$NotControl <- 0

# Ernakulam
analysis20_21$NotControl[
  analysis20_21$state_perv1 == "32" &
  analysis20_21$distcode_perv1 == "8"
] <- 1

# Chennai
analysis20_21$NotControl[
  analysis20_21$state_perv1 == "33" &
  analysis20_21$distcode_perv1 == "2"
] <- 1

# Bengaluru Urban
analysis20_21$NotControl[
  analysis20_21$state_perv1 == "29" &
  analysis20_21$distcode_perv1 == "18"
] <- 1

# Hyderabad
analysis20_21$NotControl[
  analysis20_21$state_perv1 == "36" &
  (
    (analysis20_21$period < 9 &
       analysis20_21$distcode_perv1 == "5") |
      
    (analysis20_21$period >= 9 &
       analysis20_21$distcode_perv1 == "22")
  )
] <- 1

# Mumbai Suburban
analysis20_21$NotControl[
  analysis20_21$state_perv1 == "27" &
  analysis20_21$distcode_perv1 == "22"
] <- 1

# Kolkata
analysis20_21$NotControl[
  analysis20_21$state_perv1 == "19" &
  analysis20_21$distcode_perv1 == "16"
] <- 1

# Jaipur
analysis20_21$NotControl[
  analysis20_21$state_perv1 == "8" &
  analysis20_21$distcode_perv1 == "12"
] <- 1

# Agra
analysis20_21$NotControl[
  analysis20_21$state_perv1 == "9" &
  analysis20_21$distcode_perv1 == "15"
] <- 1

# Lucknow
analysis20_21$NotControl[
  analysis20_21$state_perv1 == "9" &
  analysis20_21$distcode_perv1 == "26"
] <- 1

# All Delhi
analysis20_21$NotControl[
  analysis20_21$state_perv1 == "7"
] <- 1

# Gurugram
analysis20_21$NotControl[
  analysis20_21$state_perv1 == "6" &
  analysis20_21$distcode_perv1 == "18"
] <- 1

# Faridabad
analysis20_21$NotControl[
  analysis20_21$state_perv1 == "6" &
  analysis20_21$distcode_perv1 == "20"
] <- 1

# Gautam Buddha Nagar
analysis20_21$NotControl[
  analysis20_21$state_perv1 == "9" &
  analysis20_21$distcode_perv1 == "10"
] <- 1

# Ghaziabad
analysis20_21$NotControl[
  analysis20_21$state_perv1 == "9" &
  analysis20_21$distcode_perv1 == "9"
] <- 1

# Haryana - Jhajjar
analysis20_21$NotControl[
  analysis20_21$state_perv1 == "6" &
  analysis20_21$distcode_perv1 == "15"
] <- 1



#2021
analysis21_22$NotControl <- 0

# Ernakulam
analysis21_22$NotControl[
  analysis21_22$state_perv1 == "32" &
  analysis21_22$distcode_perv1 == "8"
] <- 1

# Chennai
analysis21_22$NotControl[
  analysis21_22$state_perv1 == "33" &
  analysis21_22$distcode_perv1 == "2"
] <- 1

# Bengaluru Urban
analysis21_22$NotControl[
  analysis21_22$state_perv1 == "29" &
  analysis21_22$distcode_perv1 == "18"
] <- 1

# Hyderabad
analysis21_22$NotControl[
  analysis21_22$state_perv1 == "36" &
  (
    (analysis21_22$period < 9 &
       analysis21_22$distcode_perv1 == "5") |
      
    (analysis21_22$period >= 9 &
       analysis21_22$distcode_perv1 == "22")
  )
] <- 1

# Mumbai Suburban
analysis21_22$NotControl[
  analysis21_22$state_perv1 == "27" &
  analysis21_22$distcode_perv1 == "22"
] <- 1

# Kolkata
analysis21_22$NotControl[
  analysis21_22$state_perv1 == "19" &
  analysis21_22$distcode_perv1 == "16"
] <- 1

# Jaipur
analysis21_22$NotControl[
  analysis21_22$state_perv1 == "8" &
  analysis21_22$distcode_perv1 == "12"
] <- 1

# Agra
analysis21_22$NotControl[
  analysis21_22$state_perv1 == "9" &
  analysis21_22$distcode_perv1 == "15"
] <- 1

# Lucknow
analysis21_22$NotControl[
  analysis21_22$state_perv1 == "9" &
  analysis21_22$distcode_perv1 == "26"
] <- 1

# All Delhi
analysis21_22$NotControl[
  analysis21_22$state_perv1 == "7"
] <- 1

# Gurugram
analysis21_22$NotControl[
  analysis21_22$state_perv1 == "6" &
  analysis21_22$distcode_perv1 == "18"
] <- 1

# Faridabad
analysis21_22$NotControl[
  analysis21_22$state_perv1 == "6" &
  analysis21_22$distcode_perv1 == "20"
] <- 1

# Gautam Buddha Nagar
analysis21_22$NotControl[
  analysis21_22$state_perv1 == "9" &
  analysis21_22$distcode_perv1 == "10"
] <- 1

# Ghaziabad
analysis21_22$NotControl[
  analysis21_22$state_perv1 == "9" &
  analysis21_22$distcode_perv1 == "9"
] <- 1

# Haryana - Jhajjar
analysis21_22$NotControl[
  analysis21_22$state_perv1 == "6" &
  analysis21_22$distcode_perv1 == "15"
] <- 1




#22
analysis22Jul_23Dec$NotControl <- 0

# Ernakulam
analysis22Jul_23Dec$NotControl[
  analysis22Jul_23Dec$state_perv1 == "32" &
  analysis22Jul_23Dec$distcode_perv1 == "8"
] <- 1

# Chennai
analysis22Jul_23Dec$NotControl[
  analysis22Jul_23Dec$state_perv1 == "33" &
  analysis22Jul_23Dec$distcode_perv1 == "2"
] <- 1

# Bengaluru Urban
analysis22Jul_23Dec$NotControl[
  analysis22Jul_23Dec$state_perv1 == "29" &
  analysis22Jul_23Dec$distcode_perv1 == "18"
] <- 1

# Hyderabad
analysis22Jul_23Dec$NotControl[
  analysis22Jul_23Dec$state_perv1 == "36" &
  (
    (analysis22Jul_23Dec$period < 9 &
       analysis22Jul_23Dec$distcode_perv1 == "5") |
      
    (analysis22Jul_23Dec$period >= 9 &
       analysis22Jul_23Dec$distcode_perv1 == "22")
  )
] <- 1



# Mumbai Suburban
analysis22Jul_23Dec$NotControl[
  analysis22Jul_23Dec$state_perv1 == "27" &
  analysis22Jul_23Dec$distcode_perv1 == "22"
] <- 1

# Kolkata
analysis22Jul_23Dec$NotControl[
  analysis22Jul_23Dec$state_perv1 == "19" &
  analysis22Jul_23Dec$distcode_perv1 == "16"
] <- 1

# Jaipur
analysis22Jul_23Dec$NotControl[
  analysis22Jul_23Dec$state_perv1 == "8" &
  analysis22Jul_23Dec$distcode_perv1 == "12"
] <- 1

# Agra
analysis22Jul_23Dec$NotControl[
  analysis22Jul_23Dec$state_perv1 == "9" &
  analysis22Jul_23Dec$distcode_perv1 == "15"
] <- 1

# Lucknow
analysis22Jul_23Dec$NotControl[
  analysis22Jul_23Dec$state_perv1 == "9" &
  analysis22Jul_23Dec$distcode_perv1 == "26"
] <- 1

# All Delhi
analysis22Jul_23Dec$NotControl[
  analysis22Jul_23Dec$state_perv1 == "7"
] <- 1

# Gurugram
analysis22Jul_23Dec$NotControl[
  analysis22Jul_23Dec$state_perv1 == "6" &
  analysis22Jul_23Dec$distcode_perv1 == "18"
] <- 1

# Faridabad
analysis22Jul_23Dec$NotControl[
  analysis22Jul_23Dec$state_perv1 == "6" &
  analysis22Jul_23Dec$distcode_perv1 == "20"
] <- 1

# Gautam Buddha Nagar
analysis22Jul_23Dec$NotControl[
  analysis22Jul_23Dec$state_perv1 == "9" &
  analysis22Jul_23Dec$distcode_perv1 == "10"
] <- 1

# Ghaziabad
analysis22Jul_23Dec$NotControl[
  analysis22Jul_23Dec$state_perv1 == "9" &
  analysis22Jul_23Dec$distcode_perv1 == "9"
] <- 1


# Haryana - Jhajjar
analysis22Jul_23Dec$NotControl[
  analysis22Jul_23Dec$state_perv1 == "6" &
  analysis22Jul_23Dec$distcode_perv1 == "15"
] <- 1




#24
analysis24$NotControl <- 0

# Ernakulam
analysis24$NotControl[
  analysis24$State_UT_Code == "32" &
  analysis24$District_Code == "8"
] <- 1

# Chennai
analysis24$NotControl[
  analysis24$State_UT_Code == "33" &
  analysis24$District_Code == "2"
] <- 1

# Bengaluru Urban
analysis24$NotControl[
  analysis24$State_UT_Code == "29" &
  analysis24$District_Code == "18"
] <- 1

# Hyderabad
analysis24$NotControl[
  analysis24$State_UT_Code == "36" &
  (
    (analysis24$period < 9 &
       analysis24$District_Code == "5") |
      
    (analysis24$period >= 9 &
       analysis24$District_Code == "22")
  )
] <- 1

# Mumbai Suburban
analysis24$NotControl[
  analysis24$State_UT_Code == "27" &
  analysis24$District_Code == "22"
] <- 1

# Kolkata
analysis24$NotControl[
  analysis24$State_UT_Code == "19" &
  analysis24$District_Code == "16"
] <- 1

# Jaipur
analysis24$NotControl[
  analysis24$State_UT_Code == "8" &
  analysis24$District_Code == "12"
] <- 1

# Agra
analysis24$NotControl[
  analysis24$State_UT_Code == "9" &
  analysis24$District_Code == "15"
] <- 1

# Lucknow
analysis24$NotControl[
  analysis24$State_UT_Code == "9" &
  analysis24$District_Code == "26"
] <- 1

# All Delhi
analysis24$NotControl[
  analysis24$State_UT_Code == "7"
] <- 1

# Gurugram
analysis24$NotControl[
  analysis24$State_UT_Code == "6" &
  analysis24$District_Code == "18"
] <- 1

# Faridabad
analysis24$NotControl[
  analysis24$State_UT_Code == "6" &
  analysis24$District_Code == "20"
] <- 1

# Gautam Buddha Nagar
analysis24$NotControl[
  analysis24$State_UT_Code == "9" &
  analysis24$District_Code == "10"
] <- 1

# Ghaziabad
analysis24$NotControl[
  analysis24$State_UT_Code == "9" &
  analysis24$District_Code == "9"
] <- 1

# Haryana - Jhajjar
analysis24$NotControl[
  analysis24$State_UT_Code == "6" &
  analysis24$District_Code == "15"
] <- 1


```


5.Weight for each person
```{r}
library(dplyr)

# 2017-18
analysis17_18 <- analysis17_18 %>%
  mutate(
    weight = case_when(
      is.na(NSS_per_fv) | is.na(NSC_per_fv) | is.na(MULT_per_fv) ~ NA_real_,
      NSS_per_fv == NSC_per_fv ~ MULT_per_fv / 100,
      NSS_per_fv != NSC_per_fv ~ MULT_per_fv / 200
    )
  )


# 2018-19
analysis18_19 <- analysis18_19 %>%
  mutate(
    weight = case_when(
      is.na(NSS_per_fv) | is.na(NSC_per_fv) | is.na(MULT_per_fv) ~ NA_real_,
      NSS_per_fv == NSC_per_fv ~ MULT_per_fv / 100,
      NSS_per_fv != NSC_per_fv ~ MULT_per_fv / 200
    )
  )


# 2019-20
analysis19_20 <- analysis19_20 %>%
  mutate(
    weight = case_when(
      is.na(NSS_per_fv) | is.na(NSC_per_fv) | is.na(MULT_per_fv) ~ NA_real_,
      NSS_per_fv == NSC_per_fv ~ MULT_per_fv / 100,
      NSS_per_fv != NSC_per_fv ~ MULT_per_fv / 200
    )
  )


# 2020-21
analysis20_21 <- analysis20_21 %>%
  mutate(
    weight = case_when(
      is.na(NSS_perv1) | is.na(NSC_perv1) | is.na(mult_perv1) ~ NA_real_,
      NSS_perv1 == NSC_perv1 ~ mult_perv1 / 100,
      NSS_perv1 != NSC_perv1 ~ mult_perv1 / 200
    )
  )


# 2021-22
analysis21_22 <- analysis21_22 %>%
  mutate(
    weight = case_when(
      is.na(NSS_perv1) | is.na(NSC_perv1) | is.na(mult_perv1) ~ NA_real_,
      NSS_perv1 == NSC_perv1 ~ mult_perv1 / 100,
      NSS_perv1 != NSC_perv1 ~ mult_perv1 / 200
    )
  )


# July 2022 - December 2023
analysis22Jul_23Dec <- analysis22Jul_23Dec %>%
  mutate(
    weight = case_when(
      is.na(NSS_perv1) | is.na(NSC_perv1) | is.na(mult_perv1) ~ NA_real_,
      NSS_perv1 == NSC_perv1 ~ mult_perv1 / 100,
      NSS_perv1 != NSC_perv1 ~ mult_perv1 / 200
    )
  )


# 2024
analysis24 <- analysis24 %>%
  mutate(
    weight = case_when(
      is.na(Ns_Count_Sector_Subsample) |
        is.na(Ns_Count_Sector_Substratum) |
        is.na(Subsample_Multiplier) ~ NA_real_,

      Ns_Count_Sector_Subsample == Ns_Count_Sector_Substratum ~
        Subsample_Multiplier / 100,

      Ns_Count_Sector_Subsample != Ns_Count_Sector_Substratum ~
        Subsample_Multiplier / 200
    )
  )



```


6.District/State identifier
```{r}
# ================================================================
# PLFS DISTRICT UID HARMONISATION: 2017-18 TO 2024
# ================================================================

library(dplyr)
library(tibble)
library(purrr)


# ================================================================
# 1. DATASET + VARIABLE CONFIGURATION
# ================================================================
#
# Variable names differ across PLFS waves.
#
# 2017-18 to 2019-20:
#   state_per_fv
#   b1q4_per_fv
#
# 2020-21 to 2022/23:
#   state_perv1
#   distcode_perv1
#
# 2024:
#   STATE_UT_Code
#   District_Code
#
# ================================================================

code_vars <- tribble(
  ~dataset,                ~period,    ~state_var,        ~district_var,      ~baseline,
  
  "analysis17_18",         "2017-18",  "state_per_fv",    "b1q4_per_fv",      TRUE,
  "analysis18_19",         "2018-19",  "state_per_fv",    "b1q4_per_fv",      TRUE,
  
  "analysis19_20",         "2019-20",  "state_per_fv",    "b1q4_per_fv",      FALSE,
  
  "analysis20_21",         "2020-21",  "state_perv1",     "distcode_perv1",   FALSE,
  "analysis21_22",         "2021-22",  "state_perv1",     "distcode_perv1",   FALSE,
  "analysis22Jul_23Dec",   "2022-23",  "state_perv1",     "distcode_perv1",   FALSE,
  
  "analysis24",            "2024",     "State_UT_Code",   "District_Code",    FALSE
)


# ================================================================
# 2. HELPER FUNCTIONS
# ================================================================

safe_integer <- function(x) {
  
  if (is.factor(x)) {
    x <- as.character(x)
  }
  
  suppressWarnings(as.integer(x))
}


make_raw_key <- function(state, district) {
  
  ifelse(
    is.na(state) | is.na(district),
    NA_character_,
    sprintf("%02d-%02d", state, district)
  )
}


standardise_codes <- function(df,
                              state_var,
                              district_var,
                              period) {
  
  # Check state variable exists
  if (!(state_var %in% names(df))) {
    stop(
      paste0(
        "State variable '",
        state_var,
        "' does not exist in ",
        period
      )
    )
  }
  
  # Check district variable exists
  if (!(district_var %in% names(df))) {
    stop(
      paste0(
        "District variable '",
        district_var,
        "' does not exist in ",
        period
      )
    )
  }
  
  
  # Extract original variables
  state_vector <- safe_integer(df[[state_var]])
  district_vector <- safe_integer(df[[district_var]])
  
  
  # Remove harmonisation variables if script is re-run
  variables_to_remove <- setdiff(
    c(
      "state_code",
      "district_code",
      "raw_district_key",
      "canonical_key",
      "district_uid",
      "survey_period"
    ),
    c(state_var, district_var)
  )
  
  
  df <- df %>%
    select(-any_of(variables_to_remove))
  
  
  # Create consistent variables for every PLFS wave
  df %>%
    mutate(
      state_code = state_vector,
      district_code = district_vector,
      
      raw_district_key =
        make_raw_key(
          state_code,
          district_code
        ),
      
      survey_period = period
    )
}


# ================================================================
# 3. LOAD DATASETS INTO A LIST
# ================================================================

datasets <- setNames(
  
  lapply(
    code_vars$dataset,
    function(x) get(x, envir = .GlobalEnv)
  ),
  
  code_vars$dataset
)


# ================================================================
# 4. STANDARDISE STATE/DISTRICT VARIABLES
# ================================================================

for (i in seq_len(nrow(code_vars))) {
  
  dataset_name <- code_vars$dataset[i]
  
  datasets[[dataset_name]] <-
    standardise_codes(
      
      df = datasets[[dataset_name]],
      
      state_var =
        code_vars$state_var[i],
      
      district_var =
        code_vars$district_var[i],
      
      period =
        code_vars$period[i]
    )
}


# ================================================================
# 5. CREATE 2017-19 BASELINE DISTRICT SYSTEM
# ================================================================
#
# 2017-18 + 2018-19 constitute the baseline.
#
# Every unique state-district pair appearing during this period
# receives its own district_uid.
# ================================================================

baseline_lookup <- bind_rows(
  
  datasets$analysis17_18 %>%
    distinct(
      state_code,
      district_code,
      raw_district_key
    ),
  
  datasets$analysis18_19 %>%
    distinct(
      state_code,
      district_code,
      raw_district_key
    )
  
) %>%
  
  filter(
    !is.na(state_code),
    !is.na(district_code)
  ) %>%
  
  distinct(
    state_code,
    district_code,
    raw_district_key
  ) %>%
  
  arrange(
    state_code,
    district_code
  ) %>%
  
  mutate(
    canonical_key = raw_district_key,
    district_uid = row_number()
  )


# ================================================================
# 6. DISTRICTS WHOSE CODE / STATE CHANGED
# ================================================================
#
# These later codes refer to districts already present in
# the 2017-19 baseline.
#
# They inherit the OLD canonical key and UID.
# ================================================================

continuity_map <- tribble(
  
  ~state_code, ~district_code, ~canonical_key, ~district_name,
  
  
  # ------------------------------------------------
  # LADAKH
  # ------------------------------------------------
  
  # Leh:
  # old Jammu & Kashmir 01-03
  # later Ladakh 37-01
  
  37,  1, "01-03", "Leh-Ladakh",
  
  
  # Kargil:
  # old Jammu & Kashmir 01-04
  # later Ladakh 37-02
  
  37,  2, "01-04", "Kargil",
  
  
  # ------------------------------------------------
  # DADRA & NAGAR HAVELI
  # ------------------------------------------------
  
  # old state 26 district 1
  # later state 25 district 3
  
  25,  3, "26-01", "Dadra & Nagar Haveli",
  
  
  # ------------------------------------------------
  # TELANGANA
  # ------------------------------------------------
  
  # Old Nizamabad 36-02 -> later 36-05
  36,  5, "36-02", "Nizamabad",
  
  # Old Karimnagar 36-03 -> later 36-13
  36, 13, "36-03", "Karimnagar",
  
  # Old Medak 36-04 -> later 36-17
  36, 17, "36-04", "Medak",
  
  # Old Hyderabad 36-05 -> later 36-22
  36, 22, "36-05", "Hyderabad",
  
  # Old Rangareddy 36-06 -> later 36-23
  36, 23, "36-06", "Rangareddy",
  
  # Old Mahbubnagar 36-07 -> later 36-25
  36, 25, "36-07", "Mahbubnagar",
  
  # Old Nalgonda 36-08 -> later 36-29
  36, 29, "36-08", "Nalgonda",
  
  # Old Khammam 36-10 -> later 36-31
  36, 31, "36-10", "Khammam"
)


# ================================================================
# NOTE ABOUT WARANGAL
# ================================================================
#
# Old:
#   36-09 = Warangal
#
# Later:
#   36-11 = Warangal Rural
#   36-12 = Warangal Urban
#
# Old Warangal was split.
#
# Neither new district is treated as geographically identical
# to the old district.
#
# Therefore both later districts receive NEW UIDs.
# ================================================================


# ================================================================
# 7. TELANGANA NEW DISTRICTS THAT REUSE OLD NUMERIC CODES
# ================================================================
#
# Example:
#
# 2017-19:
#   36-04 = Medak
#
# 2019+:
#   36-04 = Nirmal
#
# Later Nirmal therefore cannot simply use "36-04",
# because that canonical key belongs to old Medak.
# ================================================================

telangana_new_reused <- tribble(
  
  ~state_code, ~district_code, ~canonical_key, ~district_name,
  
  36,  2, "NEW-36-02", "Kumaram Bheem",
  36,  3, "NEW-36-03", "Mancherial",
  36,  4, "NEW-36-04", "Nirmal",
  36,  6, "NEW-36-06", "Jagtial",
  36,  7, "NEW-36-07", "Peddapalli",
  36,  8, "NEW-36-08", "Jayashankar",
  36,  9, "NEW-36-09", "Bhadradri Kothagudem",
  36, 10, "NEW-36-10", "Mahabubabad"
)


# ================================================================
# 8. FUNCTION TO HARMONISE 2019+ DISTRICT CODES
# ================================================================

harmonise_post2019 <- function(df) {
  
  df %>%
    
    # Check for continuing districts whose code/state changed
    left_join(
      
      continuity_map %>%
        select(
          state_code,
          district_code,
          continuity_key = canonical_key
        ),
      
      by = c(
        "state_code",
        "district_code"
      )
    ) %>%
    
    
    # Check for genuinely new Telangana districts
    # occupying recycled codes
    left_join(
      
      telangana_new_reused %>%
        select(
          state_code,
          district_code,
          reused_new_key = canonical_key
        ),
      
      by = c(
        "state_code",
        "district_code"
      )
    ) %>%
    
    
    mutate(
      
      canonical_key = case_when(
        
        # Existing district with changed code/state
        !is.na(continuity_key) ~
          continuity_key,
        
        # Genuinely new Telangana district using recycled code
        !is.na(reused_new_key) ~
          reused_new_key,
        
        # Everything else:
        #
        # existing unchanged district retains its pair
        #
        # genuinely new district gets its new pair
        TRUE ~ raw_district_key
      )
    ) %>%
    
    
    select(
      -continuity_key,
      -reused_new_key
    )
}


# ================================================================
# 9. BASELINE CANONICAL KEYS
# ================================================================

datasets$analysis17_18 <-
  datasets$analysis17_18 %>%
  mutate(
    canonical_key = raw_district_key
  )

datasets$analysis18_19 <-
  datasets$analysis18_19 %>%
  mutate(
    canonical_key = raw_district_key
  )


# ================================================================
# 10. APPLY HARMONISATION TO 2019+
# ================================================================

post2019_datasets <- c(
  "analysis19_20",
  "analysis20_21",
  "analysis21_22",
  "analysis22Jul_23Dec",
  "analysis24"
)


for (dataset_name in post2019_datasets) {
  
  datasets[[dataset_name]] <-
    harmonise_post2019(
      datasets[[dataset_name]]
    )
}


# ================================================================
# 11. FIND ALL DISTRICTS APPEARING FROM 2019-20 ONWARD
# ================================================================

all_post2019_keys <- bind_rows(
  
  lapply(
    post2019_datasets,
    
    function(dataset_name) {
      
      datasets[[dataset_name]] %>%
        distinct(canonical_key)
    }
  )
  
) %>%
  
  filter(
    !is.na(canonical_key)
  ) %>%
  
  distinct()


# ================================================================
# 12. IDENTIFY GENUINELY NEW DISTRICTS
# ================================================================
#
# Any canonical key that did not exist in the 2017-19 baseline
# receives a new UID.
#
# This automatically picks up:
#
# - Tripura additions
# - Assam additions
# - West Bengal additions
# - Gujarat additions
# - Telangana additions
# - any other genuinely new pair
# ================================================================

new_district_lookup <- all_post2019_keys %>%
  
  anti_join(
    
    baseline_lookup %>%
      select(canonical_key),
    
    by = "canonical_key"
  ) %>%
  
  arrange(canonical_key)


# ================================================================
# 13. ASSIGN NEW UIDs AFTER BASELINE IDs
# ================================================================

max_baseline_uid <-
  max(
    baseline_lookup$district_uid,
    na.rm = TRUE
  )


new_district_lookup <- new_district_lookup %>%
  
  mutate(
    
    district_uid =
      max_baseline_uid +
      row_number()
  )


# ================================================================
# 14. COMPLETE DISTRICT UID LOOKUP TABLE
# ================================================================

district_uid_lookup <- bind_rows(
  
  baseline_lookup %>%
    select(
      canonical_key,
      district_uid
    ),
  
  new_district_lookup %>%
    select(
      canonical_key,
      district_uid
    )
  
) %>%
  
  distinct(
    canonical_key,
    .keep_all = TRUE
  ) %>%
  
  arrange(
    district_uid
  )


# ================================================================
# 15. ADD UID BACK TO EVERY PLFS DATASET
# ================================================================

add_district_uid <- function(df) {
  
  df %>%
    
    select(
      -any_of("district_uid")
    ) %>%
    
    left_join(
      
      district_uid_lookup,
      
      by = "canonical_key"
    )
}


for (dataset_name in names(datasets)) {
  
  datasets[[dataset_name]] <-
    add_district_uid(
      datasets[[dataset_name]]
    )
}


# ================================================================
# 16. WRITE UPDATED DATASETS BACK INTO GLOBAL ENVIRONMENT
# ================================================================

for (dataset_name in names(datasets)) {
  
  assign(
    dataset_name,
    datasets[[dataset_name]],
    envir = .GlobalEnv
  )
}


# ================================================================
# 17. AUDIT 1:
# CHECK FOR MISSING DISTRICT UIDs
# ================================================================

missing_uid_audit <- bind_rows(
  
  lapply(
    names(datasets),
    
    function(dataset_name) {
      
      datasets[[dataset_name]] %>%
        
        summarise(
          
          dataset = dataset_name,
          
          observations = n(),
          
          missing_state =
            sum(is.na(state_code)),
          
          missing_district =
            sum(is.na(district_code)),
          
          missing_uid =
            sum(is.na(district_uid))
        )
    }
  )
)


print(
  missing_uid_audit
)


# ================================================================
# 18. AUDIT 2:
# INSPECT TELANGANA OVER TIME
# ================================================================

telangana_audit <- bind_rows(
  
  lapply(
    names(datasets),
    
    function(dataset_name) {
      
      datasets[[dataset_name]] %>%
        
        filter(
          state_code == 36
        ) %>%
        
        distinct(
          survey_period,
          state_code,
          district_code,
          raw_district_key,
          canonical_key,
          district_uid
        )
    }
  )
  
) %>%
  
  arrange(
    district_uid,
    survey_period
  )


print(
  telangana_audit,
  n = Inf
)


# ================================================================
# 19. AUDIT 3:
# SHOW CODES THAT WERE CHANGED DURING HARMONISATION
# ================================================================

changed_code_audit <- bind_rows(
  
  lapply(
    names(datasets),
    
    function(dataset_name) {
      
      datasets[[dataset_name]] %>%
        
        filter(
          raw_district_key != canonical_key
        ) %>%
        
        distinct(
          survey_period,
          state_code,
          district_code,
          raw_district_key,
          canonical_key,
          district_uid
        )
    }
  )
  
) %>%
  
  arrange(
    canonical_key,
    survey_period
  )


print(
  changed_code_audit,
  n = Inf
)


# ================================================================
# 20. AUDIT 4:
# WHICH UIDs ARE REPRESENTED BY MULTIPLE RAW CODES?
# ================================================================
#
# Expected examples:
#
# Medak:
#   36-04 and 36-17
#
# Hyderabad:
#   36-05 and 36-22
#
# Leh:
#   01-03 and 37-01
#
# ================================================================

uid_raw_code_audit <- bind_rows(
  
  lapply(
    names(datasets),
    
    function(dataset_name) {
      
      datasets[[dataset_name]] %>%
        
        distinct(
          district_uid,
          canonical_key,
          raw_district_key
        )
    }
  )
  
) %>%
  
  filter(
    !is.na(district_uid)
  ) %>%
  
  group_by(
    district_uid,
    canonical_key
  ) %>%
  
  summarise(
    
    n_raw_codes =
      n_distinct(raw_district_key),
    
    raw_codes =
      paste(
        sort(
          unique(raw_district_key)
        ),
        collapse = ", "
      ),
    
    .groups = "drop"
  ) %>%
  
  filter(
    n_raw_codes > 1
  ) %>%
  
  arrange(
    district_uid
  )


print(
  uid_raw_code_audit,
  n = Inf
)


# ================================================================
# 21. AUDIT 5:
# IDENTIFY RAW STATE-DISTRICT CODES REUSED FOR DIFFERENT PLACES
# ================================================================
#
# Expected particularly in Telangana:
#
# 36-04:
#   old Medak
#   later Nirmal
#
# ================================================================

raw_code_reuse_audit <- bind_rows(
  
  lapply(
    names(datasets),
    
    function(dataset_name) {
      
      datasets[[dataset_name]] %>%
        
        distinct(
          survey_period,
          raw_district_key,
          canonical_key,
          district_uid
        )
    }
  )
  
) %>%
  
  filter(
    !is.na(raw_district_key)
  ) %>%
  
  group_by(
    raw_district_key
  ) %>%
  
  summarise(
    
    n_canonical =
      n_distinct(canonical_key),
    
    canonical_keys =
      paste(
        sort(
          unique(canonical_key)
        ),
        collapse = ", "
      ),
    
    district_uids =
      paste(
        sort(
          unique(district_uid)
        ),
        collapse = ", "
      ),
    
    .groups = "drop"
  ) %>%
  
  filter(
    n_canonical > 1
  )


print(
  raw_code_reuse_audit,
  n = Inf
)


# ================================================================
# 22. AUDIT 6:
# SHOW ALL GENUINELY NEW DISTRICTS
# ================================================================

print(
  new_district_lookup,
  n = Inf
)


# ================================================================
# 23. CHECK TREATMENT DISTRICTS
# ================================================================
#
# Ahmedabad      = 24-07
# Nagpur         = 27-09
# Kanpur Nagar   = 09-33
# Pune           = 27-25
#
# ================================================================

treatment_district_audit <- bind_rows(
  
  lapply(
    names(datasets),
    
    function(dataset_name) {
      
      datasets[[dataset_name]] %>%
        
        filter(
          raw_district_key %in%
            c(
              "24-07",
              "27-09",
              "09-33",
              "27-25"
            )
        ) %>%
        
        distinct(
          survey_period,
          state_code,
          district_code,
          raw_district_key,
          canonical_key,
          district_uid
        )
    }
  )
  
) %>%
  
  arrange(
    raw_district_key,
    survey_period
  )


print(
  treatment_district_audit,
  n = Inf
)


# ================================================================
# FINISHED
# ================================================================

message(
  "District UID harmonisation complete."
)
```

7.Clear memory
```{r}
# ================================================================
# CLEAN UP TEMPORARY OBJECTS FROM DISTRICT HARMONISATION
# ================================================================

rm(
  # Temporary data / lookup objects
  datasets,
  code_vars,
  baseline_lookup,
  continuity_map,
  telangana_new_reused,
  all_post2019_keys,
  new_district_lookup,
  max_baseline_uid,

  # Audit tables
  missing_uid_audit,
  telangana_audit,
  changed_code_audit,
  uid_raw_code_audit,
  raw_code_reuse_audit,
  treatment_district_audit,

  # Helper functions
  safe_integer,
  make_raw_key,
  standardise_codes,
  harmonise_post2019,
  add_district_uid
)

# Optional: force R to reclaim unused memory
gc()
```


8.Combining the analysis dataframes
```{r}
# ================================================================
# 8. CREATE COMPACT HARMONISED DATA FOR DiD
# ================================================================

library(dplyr)
library(haven)
library(readr)
library(here)

employed_status_codes <- c("11", "12", "21", "31", "41", "51")


# ================================================================
# 8.1 HOUSEHOLD CONTROLS
# ================================================================

normalise_key_piece <- function(x) {
  out <- trimws(as.character(x))
  out[out %in% c("", "NA", "NaN")] <- NA_character_

  numeric_value <- suppressWarnings(as.numeric(out))
  numeric_piece <- !is.na(out) & !is.na(numeric_value)

  out[numeric_piece] <- format(
    numeric_value[numeric_piece],
    scientific = FALSE,
    trim = TRUE,
    drop0trailing = TRUE
  )

  out
}


make_household_key <- function(df, key_vars) {

  missing_vars <- setdiff(key_vars, names(df))

  if (length(missing_vars) > 0L) {
    stop(
      "Missing household-key variables: ",
      paste(missing_vars, collapse = ", ")
    )
  }

  key_parts <- lapply(
    key_vars,
    function(v) normalise_key_piece(df[[v]])
  )

  do.call(paste, c(key_parts, sep = "|"))
}


build_household_controls <- function(df,
                                     key_vars,
                                     household_size_var,
                                     religion_var,
                                     social_group_var) {

  controls <- tibble(
    household_join_key = make_household_key(df, key_vars),

    household_size = suppressWarnings(
      as.integer(as.character(df[[household_size_var]]))
    ),

    religion = suppressWarnings(
      as.integer(as.character(df[[religion_var]]))
    ),

    social_group = suppressWarnings(
      as.integer(as.character(df[[social_group_var]]))
    ),

    .household_match = 1L
  )

  if (anyDuplicated(controls$household_join_key)) {
    stop("A household control file contains duplicate household keys.")
  }

  controls
}


read_household_controls_dta <- function(path,
                                        key_vars,
                                        household_size_var,
                                        religion_var,
                                        social_group_var) {

  variables_needed <- unique(
    c(
      key_vars,
      household_size_var,
      religion_var,
      social_group_var
    )
  )

  household_data <- read_dta(
    path,
    col_select = all_of(variables_needed)
  )

  build_household_controls(
    household_data,
    key_vars,
    household_size_var,
    religion_var,
    social_group_var
  )
}


read_household_controls_csv <- function(path,
                                        key_vars,
                                        household_size_var,
                                        religion_var,
                                        social_group_var) {

  variables_needed <- unique(
    c(
      key_vars,
      household_size_var,
      religion_var,
      social_group_var
    )
  )

  household_data <- read_csv(
    path,
    col_select = all_of(variables_needed),
    show_col_types = FALSE,
    progress = FALSE
  )

  build_household_controls(
    household_data,
    key_vars,
    household_size_var,
    religion_var,
    social_group_var
  )
}


# Person-file household identifiers: 2017-18 to 2019-20
person_keys_early <- c(
  "quarter_per_fv",
  "visit_per_fv",
  "b1q3_per_fv",
  "state_per_fv",
  "b1q4_per_fv",
  "nss_region_per_fv",
  "b1q5_per_fv",
  "b1q6_per_fv",
  "b1q11_per_fv",
  "fsu_per_fv",
  "b1q13_per_fv",
  "b1q14_per_fv",
  "b1q15_per_fv"
)


# Household-file identifiers: 2017-18 and 2019-20
household_keys_early_fv <- c(
  "qtr_hh_fv",
  "visit_hh_fv",
  "b1q3_hh_fv",
  "state_hh_fv",
  "dist_code_hh_fv",
  "nss_region_hh_fv",
  "b1q5_hh_fv",
  "b1q6_hh_fv",
  "b1q11_hh_fv",
  "b1q1_hh_fv",
  "b1q13_hh_fv",
  "b1q14_hh_fv",
  "b1q15_hh_fv"
)


# The 2018-19 household file uses the _rv suffix even though it is
# the household file corresponding to the first-visit person file.
household_keys_2018 <- c(
  "qtr_hh_rv",
  "visit_hh_rv",
  "b1q3_hh_rv",
  "state_hh_rv",
  "dist_code_hh_rv",
  "nss_region_hh_rv",
  "b1q5_hh_rv",
  "b1q6_hh_rv",
  "b1q11_hh_rv",
  "b1q1_hh_rv",
  "b1q13_hh_rv",
  "b1q14_hh_rv",
  "b1q15_hh_rv"
)


# Person-file identifiers: 2020-21 to 2023-24
person_keys_later <- c(
  "qtr_perv1",
  "visit_perv1",
  "b1q3_perv1",
  "state_perv1",
  "distcode_perv1",
  "nss_region_perv1",
  "b1q5_perv1",
  "b1q6_perv1",
  "b1q11_perv1",
  "b1q1_perv1",
  "b1q13_perv1",
  "b1q14_perv1",
  "b1q15_perv1"
)


# Household-file identifiers: 2020-21 to 2023-24
household_keys_later <- c(
  "qtr_hhv1",
  "visit_hhv1",
  "b1q3_hhv1",
  "state_hhv1",
  "distcode_hhv1",
  "nss_region_hhv1",
  "b1q5_hhv1",
  "b1q6_hhv1",
  "b1q11_hhv1",
  "b1q1_hhv1",
  "b1q13_hhv1",
  "b1q14_hhv1",
  "b1q15_hhv1"
)


# Calendar-year 2024 identifiers
person_keys_2024 <- c(
  "Quarter",
  "Visit",
  "Sector",
  "State_UT_Code",
  "District_Code",
  "NSS_Region",
  "Stratum",
  "Sub_Stratum",
  "Sub_Sample",
  "FSU",
  "Sample_Sg_Sb_No",
  "Second_Stage_Stratum_No",
  "Sample_Household_Number"
)


household_keys_2024 <- c(
  "Quarter",
  "Visit",
  "Sector",
  "State_Ut_Code",
  "District_Code",
  "NSS_Region",
  "Stratum",
  "Sub_Stratum",
  "Sub_Sample",
  "FSU",
  "Sample_Sg_Sb_No",
  "Second_Stage_Stratum_No",
  "Sample_Household_Number"
)


# Load only the household variables required for the controls.
hh_controls_17_18 <- read_household_controls_dta(
  here("Data", "17-18", "hhfv_2017-18.dta"),
  household_keys_early_fv,
  "b3q1_hh_fv",
  "b3q3_hh_fv",
  "b3q4_hh_fv"
)


hh_controls_18_19 <- read_household_controls_dta(
  here("Data", "18-19", "HHV1_2018-19 (1).dta"),
  household_keys_2018,
  "b3q1_hh_rv",
  "b3q3_hh_rv",
  "b3q4_hh_rv"
)


hh_controls_19_20 <- read_household_controls_dta(
  here("Data", "19-20", "HHFV_2019-20.dta"),
  household_keys_early_fv,
  "b3q1_hh_fv",
  "b3q3_hh_fv",
  "b3q4_hh_fv"
)


hh_controls_20_21 <- read_household_controls_dta(
  here("Data", "20-21", "hhv1.dta"),
  household_keys_later,
  "b3q1_hhv1",
  "b3q3_hhv1",
  "b3q4_hhv1"
)


hh_controls_21_22 <- read_household_controls_csv(
  here("Data", "21-22", "hhv1.csv"),
  household_keys_later,
  "b3q1_hhv1",
  "b3q3_hhv1",
  "b3q4_hhv1"
)


hh_controls_22_23 <- read_household_controls_dta(
  here("Data", "22-23", "hhv1 (1).dta"),
  household_keys_later,
  "b3q1_hhv1",
  "b3q3_hhv1",
  "b3q4_hhv1"
)


# Add Q1 and Q2 from the 2023-24 household file so that the
# household-control periods match analysis22Jul_23Dec.
hh_23_24_variables <- unique(
  c(
    household_keys_later,
    "b3q1_hhv1",
    "b3q3_hhv1",
    "b3q4_hhv1"
  )
)


hh_23_24_q1q2 <- read_dta(
  here("Data", "23-24", "hhv1.dta"),
  col_select = all_of(hh_23_24_variables)
) %>%
  filter(qtr_hhv1 %in% c("Q1", "Q2"))


hh_controls_23_24_q1q2 <- build_household_controls(
  hh_23_24_q1q2,
  household_keys_later,
  "b3q1_hhv1",
  "b3q3_hhv1",
  "b3q4_hhv1"
)


hh_controls_22Jul_23Dec <- bind_rows(
  hh_controls_22_23,
  hh_controls_23_24_q1q2
)


if (anyDuplicated(hh_controls_22Jul_23Dec$household_join_key)) {
  stop("Duplicate household keys after combining 2022-23 and 2023-24.")
}


hh_controls_24 <- read_household_controls_dta(
  here("Data", "24", "chhv1.dta"),
  household_keys_2024,
  "Household_Size",
  "Religion",
  "Social_Group"
)


# ================================================================
# 8.2 CONVERT EACH PERSON FILE TO THE SAME DiD VARIABLES
# ================================================================

make_did_wave <- function(df,
                          household_controls,
                          person_key_vars,
                          urban_var,
                          status_var,
                          age_var,
                          sex_var,
                          marital_var,
                          education_var) {

  required_variables <- unique(
    c(
      person_key_vars,
      urban_var,
      status_var,
      age_var,
      sex_var,
      marital_var,
      education_var,
      "survey_period",
      "period",
      "state_code",
      "district_code",
      "district_uid",
      "city_size_stratum",
      "eligible_city",
      "metro_system",
      "first_treat",
      "first_treat_opening_quarter",
      "Treatment",
      "Treatment_opening_quarter",
      "weight",
      "transport",
      "informal",
      "NotControl"
    )
  )

  missing_variables <- setdiff(required_variables, names(df))

  if (length(missing_variables) > 0L) {
    stop(
      "The following variables are missing from a person wave: ",
      paste(missing_variables, collapse = ", ")
    )
  }

  original_rows <- nrow(df)

  joined_data <- df %>%
    mutate(
      household_join_key = make_household_key(., person_key_vars)
    ) %>%
    left_join(
      household_controls,
      by = "household_join_key",
      relationship = "many-to-one"
    )

  if (nrow(joined_data) != original_rows) {
    stop("The household join changed the number of person observations.")
  }

  if (any(is.na(joined_data$.household_match))) {
    stop("At least one person observation did not match a household record.")
  }

  joined_data %>%

    transmute(

      # Time and geography
      survey_period = as.character(survey_period),
      period = as.integer(period),

      state_code = as.integer(state_code),
      district_code = as.integer(district_code),
      district_uid = as.integer(district_uid),

      urban = suppressWarnings(
        as.integer(as.character(.data[[urban_var]]))
      ),

      city_size_stratum = as.integer(city_size_stratum),
      eligible_city = as.integer(eligible_city),


      # Treatment variables constructed in Section 3
      metro_system = as.character(metro_system),
      first_treat = as.integer(first_treat),

      first_treat_opening_quarter = as.integer(
        first_treat_opening_quarter
      ),

      Treatment = as.integer(Treatment),

      Treatment_opening_quarter = as.integer(
        Treatment_opening_quarter
      ),


      # Survey information and existing classifications
      weight = as.numeric(weight),
      transport = as.integer(transport),
      informal = as.integer(informal),
      NotControl = as.integer(NotControl),


      # Individual controls
      age = suppressWarnings(
        as.integer(as.character(.data[[age_var]]))
      ),

      sex = suppressWarnings(
        as.integer(as.character(.data[[sex_var]]))
      ),

      marital_status = suppressWarnings(
        as.integer(as.character(.data[[marital_var]]))
      ),

      general_education = suppressWarnings(
        as.integer(as.character(.data[[education_var]]))
      ),


      # Household controls
      household_size = as.integer(household_size),
      religion = as.integer(religion),
      social_group = as.integer(social_group),


      # Temporary usual-principal-status variable
      status = as.character(.data[[status_var]])
    ) %>%

    mutate(

      # Main and robustness age samples
      working_age_15_64 = as.integer(
        !is.na(age) & age >= 15L & age <= 64L
      ),

      age_15_plus = as.integer(
        !is.na(age) & age >= 15L
      ),


      # Flexible age categories for the conditional specification
      age_group = case_when(
        age >= 15L & age <= 24L ~ "15-24",
        age >= 25L & age <= 34L ~ "25-34",
        age >= 35L & age <= 44L ~ "35-44",
        age >= 45L & age <= 54L ~ "45-54",
        age >= 55L & age <= 64L ~ "55-64",
        age >= 65L ~ "65+",
        TRUE ~ NA_character_
      ),

      age_group = factor(
        age_group,
        levels = c(
          "15-24",
          "25-34",
          "35-44",
          "45-54",
          "55-64",
          "65+"
        )
      ),


      # Employment status based on usual principal status
      employed = case_when(
        is.na(status) | status == "" ~ NA_integer_,
        status %in% employed_status_codes ~ 1L,
        TRUE ~ 0L
      ),


      # Total passenger-transport employment among the population
      transport_employment = case_when(
        is.na(employed) ~ NA_integer_,
        employed == 0L ~ 0L,
        employed == 1L & transport == 1L ~ 1L,
        employed == 1L & transport == 0L ~ 0L,
        TRUE ~ NA_integer_
      ),


      # Employment outside the selected passenger-transport industries
      nontransport_employment = case_when(
        is.na(employed) ~ NA_integer_,
        employed == 0L ~ 0L,
        employed == 1L & transport == 0L ~ 1L,
        employed == 1L & transport == 1L ~ 0L,
        TRUE ~ NA_integer_
      ),


      # PRIMARY OUTCOME:
      # probability that a working-age resident is employed in
      # informal passenger transport
      informal_transport = case_when(
        is.na(employed) ~ NA_integer_,
        employed == 0L ~ 0L,
        employed == 1L & transport == 0L ~ 0L,
        employed == 1L & transport == 1L & informal == 1L ~ 1L,
        employed == 1L & transport == 1L & informal == 0L ~ 0L,
        TRUE ~ NA_integer_
      ),


      # Formal passenger-transport employment among the population
      formal_transport = case_when(
        is.na(employed) ~ NA_integer_,
        employed == 0L ~ 0L,
        employed == 1L & transport == 0L ~ 0L,
        employed == 1L & transport == 1L & informal == 0L ~ 1L,
        employed == 1L & transport == 1L & informal == 1L ~ 0L,
        TRUE ~ NA_integer_
      ),


      # Secondary composition outcome.
      # This is only observed for transport workers and should not
      # replace informal_transport as the primary outcome.
      informal_among_transport_workers = case_when(
        transport_employment == 1L & informal == 1L ~ 1L,
        transport_employment == 1L & informal == 0L ~ 0L,
        TRUE ~ NA_integer_
      ),


      # Event time for descriptive checks.
      # The did package will construct its own dynamic aggregation.
      event_time = case_when(
        eligible_city == 1L & first_treat > 0L ~
          period - first_treat,
        TRUE ~ NA_integer_
      ),

      event_time_opening_quarter = case_when(
        eligible_city == 1L & first_treat_opening_quarter > 0L ~
          period - first_treat_opening_quarter,
        TRUE ~ NA_integer_
      )
    ) %>%

    select(-status)
}


# ================================================================
# 8.3 HARMONISE EACH WAVE
# ================================================================

did17_18 <- make_did_wave(
  df = analysis17_18,
  household_controls = hh_controls_17_18,
  person_key_vars = person_keys_early,
  urban_var = "b1q3_per_fv",
  status_var = "b5pt1q3_per_fv",
  age_var = "b4q6_per_fv",
  sex_var = "b4q5_per_fv",
  marital_var = "b4q7_per_fv",
  education_var = "b4q8_per_fv"
)


did18_19 <- make_did_wave(
  df = analysis18_19,
  household_controls = hh_controls_18_19,
  person_key_vars = person_keys_early,
  urban_var = "b1q3_per_fv",
  status_var = "b5pt1q3_per_fv",
  age_var = "b4q6_per_fv",
  sex_var = "b4q5_per_fv",
  marital_var = "b4q7_per_fv",
  education_var = "b4q8_per_fv"
)


did19_20 <- make_did_wave(
  df = analysis19_20,
  household_controls = hh_controls_19_20,
  person_key_vars = person_keys_early,
  urban_var = "b1q3_per_fv",
  status_var = "b5pt1q3_per_fv",
  age_var = "b4q6_per_fv",
  sex_var = "b4q5_per_fv",
  marital_var = "b4q7_per_fv",
  education_var = "b4q8_per_fv"
)


did20_21 <- make_did_wave(
  df = analysis20_21,
  household_controls = hh_controls_20_21,
  person_key_vars = person_keys_later,
  urban_var = "b1q3_perv1",
  status_var = "b5pt1q3_perv1",
  age_var = "b4q6_perv1",
  sex_var = "b4q5_perv1",
  marital_var = "b4q7_perv1",
  education_var = "b4q8_perv1"
)


did21_22 <- make_did_wave(
  df = analysis21_22,
  household_controls = hh_controls_21_22,
  person_key_vars = person_keys_later,
  urban_var = "b1q3_perv1",
  status_var = "b5pt1q3_perv1",
  age_var = "b4q6_perv1",
  sex_var = "b4q5_perv1",
  marital_var = "b4q7_perv1",
  education_var = "b4q8_perv1"
)


did22_23 <- make_did_wave(
  df = analysis22Jul_23Dec,
  household_controls = hh_controls_22Jul_23Dec,
  person_key_vars = person_keys_later,
  urban_var = "b1q3_perv1",
  status_var = "b5pt1q3_perv1",
  age_var = "b4q6_perv1",
  sex_var = "b4q5_perv1",
  marital_var = "b4q7_perv1",
  education_var = "b4q8_perv1"
)


did24 <- make_did_wave(
  df = analysis24,
  household_controls = hh_controls_24,
  person_key_vars = person_keys_2024,
  urban_var = "Sector",
  status_var = "Principal_Status_Code",
  age_var = "Age",
  sex_var = "Sex",
  marital_var = "Marital_Status",
  education_var = "General_Education_Level"
)


# ================================================================
# 8.4 STACK THE REPEATED CROSS-SECTIONS
# ================================================================

did_data_all <- bind_rows(
  did17_18,
  did18_19,
  did19_20,
  did20_21,
  did21_22,
  did22_23,
  did24
)


# Keep only comparable Census-2011 large-city strata.
did_data_large_cities <- did_data_all %>%
  filter(eligible_city == 1L)


# Remove districts that already had metro exposure and therefore
# cannot be clean comparison units.
did_data_analysis_pool <- did_data_large_cities %>%
  filter(!is.na(NotControl), NotControl == 0L)


# Main population: working-age residents aged 15-64.
did_data <- did_data_analysis_pool %>%
  filter(working_age_15_64 == 1L)


# Robustness population: everyone aged 15 and above.
did_data_15plus <- did_data_analysis_pool %>%
  filter(age_15_plus == 1L)


# ================================================================
# 8.5 PRE-TREATMENT CITY-LEVEL CONTROLS
# ================================================================

weighted_mean_safe <- function(x, w) {
  valid <- !is.na(x) & !is.na(w) & w > 0

  if (!any(valid)) {
    return(NA_real_)
  }

  weighted.mean(x[valid], w[valid])
}


# Periods 1-6 end before the Ahmedabad and Nagpur openings.
# These variables are predetermined and may be used selectively
# in sensitivity specifications.
baseline_city_controls <- did_data %>%
  filter(period <= 6L) %>%
  group_by(state_code, district_uid) %>%
  summarise(
    baseline_informal_transport_share = weighted_mean_safe(
      informal_transport,
      weight
    ),

    baseline_transport_employment_share = weighted_mean_safe(
      transport_employment,
      weight
    ),

    baseline_employment_rate = weighted_mean_safe(
      employed,
      weight
    ),

    .groups = "drop"
  )


did_data <- did_data %>%
  left_join(
    baseline_city_controls,
    by = c("state_code", "district_uid"),
    relationship = "many-to-one"
  )


did_data_15plus <- did_data_15plus %>%
  left_join(
    baseline_city_controls,
    by = c("state_code", "district_uid"),
    relationship = "many-to-one"
  )


# ================================================================
# 8.6 DISTRICT-QUARTER ROBUSTNESS DATA
# ================================================================

# This is a robustness/descriptive dataset. The individual-level
# repeated-cross-section data above remain the primary specification.
did_district_quarter <- did_data %>%
  group_by(
    survey_period,
    period,
    state_code,
    district_code,
    district_uid,
    metro_system,
    first_treat
  ) %>%
  summarise(
    unweighted_n = n(),
    population_weight = sum(weight, na.rm = TRUE),

    informal_transport_rate = weighted_mean_safe(
      informal_transport,
      weight
    ),

    transport_employment_rate = weighted_mean_safe(
      transport_employment,
      weight
    ),

    formal_transport_rate = weighted_mean_safe(
      formal_transport,
      weight
    ),

    overall_employment_rate = weighted_mean_safe(
      employed,
      weight
    ),

    informal_share_among_transport_workers = weighted_mean_safe(
      informal_among_transport_workers,
      weight
    ),

    .groups = "drop"
  )


# ================================================================
# 8.7 FINAL VALIDATION
# ================================================================

stopifnot(
  all(did_data$eligible_city == 1L),
  all(did_data$NotControl == 0L),
  all(did_data$age >= 15L & did_data$age <= 64L),
  !any(is.na(did_data$weight)),
  !any(did_data$weight <= 0, na.rm = TRUE)
)


did_sample_check <- did_data %>%
  summarise(
    observations = n(),
    periods = n_distinct(period),
    districts = n_distinct(district_uid),
    treated_metro_systems = n_distinct(
      metro_system[!is.na(metro_system)]
    ),
    missing_age = sum(is.na(age)),
    missing_sex = sum(is.na(sex)),
    missing_education = sum(is.na(general_education)),
    missing_marital_status = sum(is.na(marital_status)),
    missing_religion = sum(is.na(religion)),
    missing_social_group = sum(is.na(social_group)),
    missing_household_size = sum(is.na(household_size))
  )


print(did_sample_check)


# Recommended later regression formulas:
#
# Unconditional benchmark:
#   xformla = ~ 1
#
# Parsimonious conditional specification:
#   xformla = ~ age_group + factor(sex) +
#     factor(general_education) + factor(marital_status) +
#     factor(social_group) + factor(religion) +
#     factor(state_code)
#
# Household-size sensitivity:
#   add household_size to the conditional formula.
#
# Do not control for transport, informal, enterprise type,
# social-security coverage, contract type, earnings or current
# household expenditure: these define the outcome or may be affected
# by metro opening.


# Optional memory clean-up
rm(
  hh_controls_17_18,
  hh_controls_18_19,
  hh_controls_19_20,
  hh_controls_20_21,
  hh_controls_21_22,
  hh_controls_22_23,
  hh_controls_23_24_q1q2,
  hh_controls_22Jul_23Dec,
  hh_controls_24,
  hh_23_24_q1q2
)

gc()

```


9.DID

