---
title: "Cleaning the Dataset"
---
Import the data.
```{r}
#for reading dta
library(haven)
#for easier file navigation
library(here)

data17_18_fv <- read_dta(here("Data", "17-18", "hh_per_fv_2017-18.dta"))

data18_19_fv <- read_dta(here("Data","18-19","PerV1_2018-19.dta"))

data19_20_fv <- read_dta(here("Data","19-20","PERFV_2019-20.dta"))

data20_21_fv <- read_dta(here("Data","20-21","perv1.dta"))

data21_22_fv <- readr::read_csv(here("Data","21-22","perv1.csv"))

data22_23 <- read_dta(here("Data","22-23","perv1 (2).dta"))

data23_24 <- read_dta(here("Data","23-24","perv1.dta"))

data24 <- read_dta(here("Data","24","cperv1.dta"))

data25 <- read_dta(here("Data","25","cperv12025.dta"))

```

Creating analysis dataframe:
```{r}
library(dplyr)
analysis17_18 <- data17_18_fv
analysis18_19 <- data18_19_fv
analysis19_20 <- data19_20_fv
analysis20_21 <- data20_21_fv
analysis21_22 <- data21_22_fv
analysis22_23 <- data22_23
analysis23_24 <- data23_24
analysis24 <- data24
analysis25 <- data25

#Zapping stata labels because they get in the way of dataframe manipulation
analysis17_18 <- analysis17_18 %>% mutate(across(where(haven::is.labelled), haven::zap_labels))
analysis18_19 <- analysis18_19 %>% mutate(across(where(haven::is.labelled), haven::zap_labels))
analysis19_20 <- analysis19_20 %>% mutate(across(where(haven::is.labelled), haven::zap_labels))
analysis20_21 <- analysis20_21 %>% mutate(across(where(haven::is.labelled), haven::zap_labels))
analysis21_22 <- analysis21_22 %>% mutate(across(where(haven::is.labelled), haven::zap_labels))
analysis22_23 <- analysis22_23 %>% mutate(across(where(haven::is.labelled), haven::zap_labels))
analysis23_24 <- analysis23_24 %>% mutate(across(where(haven::is.labelled), haven::zap_labels))
analysis24 <- analysis24 %>% mutate(across(where(haven::is.labelled), haven::zap_labels))
analysis25 <- analysis25 %>% mutate(across(where(haven::is.labelled), haven::zap_labels))

#get rid of "data" so I don't even accidentally write over actual dataset
rm(list=c("data17_18_fv","data18_19_fv","data19_20_fv","data20_21_fv","data21_22_fv","data22_23","data23_24","data24","data25"))
```

Matching quarters in 2024 data to consolidate the two:
```{r}
aggregate(data.frame(count = analysis22_23$qtr_perv1), list(value = analysis22_23$qtr_perv1), length)

aggregate(data.frame(count = analysis23_24$qtr_perv1), list(value = analysis23_24$qtr_perv1), length)

aggregate(data.frame(count = analysis24$Quarter), list(value = analysis24$Quarter), length)

#Task now is to add the Q1 and Q2 from 23-24 into the 22 data, so it becomes 22 to end of 23 data.Then, get rid of 23_24 because Q3 and Q4 from 23_24 is already in 24.

Q1Q2_2324 <- analysis23_24 %>% filter(qtr_perv1 %in% c("Q1","Q2"))
analysis22J_23D <- rbind(analysis22_23,Q1Q2_2324)
rm(analysis22_23)
rm(analysis23_24)
rm(Q1Q2_2324)

#Before this, I checked that the columns/variables recorded only differ by 1 and that is panel. Given observations in 23_24 and 24 have identical 24 components, no need to correct for those. 22_23 and 23_24 also had identical columns, so no need to check.
```

1.Informal and Transport binaries.
1.1Matching the NIC columns:
```{r}
NIC17 <- analysis17_18$b5pt1q5_per_fv
NCO17 <- analysis17_18$b5pt2q6_per_fv

sum(is.na(NIC17) | NIC17 == "")

table(analysis17_18$b5pt1q3_per_fv, 
      is.na(analysis17_18$b5pt1q5_per_fv),
      useNA = "ifany")


nic_missing <- is.na(analysis17_18$b5pt1q5_per_fv) |
               trimws(analysis17_18$b5pt1q5_per_fv) == ""

tab <- table(
  analysis17_18$b5pt1q3_per_fv,
  nic_missing
)

tab <- as.data.frame.matrix(tab)

tab$status <- rownames(tab)
rownames(tab) <- NULL

tab <- tab[, c("status", "FALSE", "TRUE")]
names(tab) <- c("status", "NIC_present", "NIC_missing")

tab




length(unique(analysis17_18$MULT_per_fv))
```

1.2Checking which NIC values are present and whether there are empty rows:
```{r}
aggregate(data.frame(count = analysis17_18$b5pt2q5_per_fv), list(value = analysis17_18$b5pt2q5_per_fv), length)
```
#We find that all NIC values in data frames _____ are 6 digits long.




1.3Creating transport binary:
```{r}
library(dplyr)

vectorrr <- startsWith(NIC17,"491")
sum(vectorrr,na.rm=TRUE)

analysis17_18$transport <- startsWith(NIC, "491") | startsWith(NIC, "492")
sum(analysis17_18$transport,na.rm=TRUE)

analysis17_18$transport <- ifelse(
  NIC17 %in% c("49110", "49211","49212","49213","49219",
               "49222","49223","49224","49226","49229" ),
  1,
  0
)

analysis18_19$transport <- ifelse(
  NIC18 %in% c("4911", "49211", "49219", "492", "4921", "49212", "4922", "49213","49223","49224","49226","49229" ),
  1,
  0
)

analysis19_20$transport <- ifelse(
  NIC19 %in% c("4911", "49211", "49219", "492", "4921", "49212", "4922", "49213","49223","49224","49226","49229" ),
  1,
  0
)

analysis20_21$transport <- ifelse(
  NIC20 %in% c("4911", "49211", "49219", "492", "4921", "49212", "4922", "49213","49223","49224","49226","49229" ),
  1,
  0
)

analysis21_22$transport <- ifelse(
  NIC21 %in% c("4911", "49211", "49219", "492", "4921", "49212", "4922", "49213","49223","49224","49226","49229" ),
  1,
  0
)

analysis22_23$transport <- ifelse(
  NIC22 %in% c("4911", "49211", "49219", "492", "4921", "49212", "4922", "49213","49223","49224","49226","49229" ),
  1,
  0
)

analysis23_24$transport <- ifelse(
  NIC23 %in% c("4911", "49211", "49219", "492", "4921", "49212", "4922", "49213","49223","49224","49226","49229" ),
  1,
  0
)

analysis24$transport <- ifelse(
  NIC24 %in% c("4911", "49211", "49219", "492", "4921", "49212", "4922", "49213","49223","49224","49226","49229" ),
  1,
  0
)

analysis25$transport <- ifelse(
  NIC25 %in% c("491100", "492101","492103","492203","492204","492299" ),
  1,
  0
)

#Repeat for all years until 2025 when NIC codes changed

```



















Function for cleaning the dataset, so that the stata labels don't interfere with the data cleaning process.
```{r}
# ============================================================
# CLEAN PLFS CODE VARIABLES
# ============================================================

clean_code <- function(x) {

  # Remove Stata value labels while preserving underlying codes
  if (
    inherits(x, "haven_labelled") ||
    inherits(x, "labelled")
  ) {
    x <- haven::zap_labels(x)
  }

  # Prevent factors from being converted to their level numbers
  if (is.factor(x)) {
    x <- as.character(x)
  }

  # Clean character-formatted codes
  if (is.character(x)) {

    x <- trimws(x)

    x[
      x %in% c(
        "",
        "NA",
        "N.A.",
        "NaN",
        "NULL",
        "-",
        "--"
      )
    ] <- NA_character_

    # Convert values such as "31.0" to "31"
    x <- sub(
      "\\.0+$",
      "",
      x
    )
  }

  suppressWarnings(
    as.numeric(x)
  )
}


# ============================================================
# CLEAN NIC INDUSTRY CODES
#
# Retains whether the original code was recorded at:
#   4-digit level, e.g. 4921
#   5-digit level, e.g. 49213
#
# This is necessary because a four-digit 4921 code cannot
# distinguish buses from trams or metro rail.
# ============================================================

clean_nic_code <- function(x) {

  # Remove Stata value labels while preserving underlying codes
  if (
    inherits(x, "haven_labelled") ||
      inherits(x, "labelled")
  ) {
    x <- haven::zap_labels(x)
  }

  # Prevent factors being converted to factor-level numbers
  if (is.factor(x)) {
    x <- as.character(x)
  }

  x <- trimws(
    as.character(x)
  )

  x[
    x %in% c(
      "",
      "NA",
      "N.A.",
      "NaN",
      "NULL",
      "-",
      "--"
    )
  ] <- NA_character_

  # Remove decimal suffixes created during CSV import
  x <- sub(
    "\\.0+$",
    "",
    x
  )

  # Remove spaces and other non-numeric characters
  x <- gsub(
    "[^0-9]",
    "",
    x
  )

  # Blank strings created by cleaning become missing
  x[
    x == ""
  ] <- NA_character_

  # Only accept four- or five-digit NIC codes
  x[
    !is.na(x) &
      !(nchar(x) %in% c(4L, 5L))
  ] <- NA_character_

  x
}


# ============================================================
# CREATE NON-METRO TRANSPORT AND INFORMAL-EMPLOYMENT VARIABLES
#
# Main variables created:
#
# transport
#   1 = non-metro passenger land transport worker
#   0 = not a non-metro passenger transport worker
#  NA = insufficient NIC detail to determine status
#
# metro_worker
#   1 = NIC 49213
#
# tram_worker
#   1 = NIC 49212
#
# passenger_land_transport_broad
#   1 = any NIC beginning 4921 or 4922, including metro/tram
#
# informal_transport
#   1 = informal AND non-metro passenger transport
# ============================================================

make_informal_transport <- function(
    data,
    columns,
    dataset_name = "dataset"
) {

  # ----------------------------------------------------------
  # Check that the column mapping contains all required keys
  # ----------------------------------------------------------

  required_keys <- c(
    "pas",
    "nic",
    "enterprise",
    "enterprise_size",
    "social_security"
  )

  missing_keys <- setdiff(
    required_keys,
    names(columns)
  )

  if (length(missing_keys) > 0L) {

    stop(
      paste0(
        dataset_name,
        " column mapping is missing: ",
        paste(
          missing_keys,
          collapse = ", "
        )
      )
    )
  }


  mapped_columns <- unname(
    unlist(
      columns[required_keys]
    )
  )


  # ----------------------------------------------------------
  # Check that all mapped variables exist
  # ----------------------------------------------------------

  missing_columns <- setdiff(
    mapped_columns,
    names(data)
  )

  if (length(missing_columns) > 0L) {

    stop(
      paste0(
        dataset_name,
        " is missing the following variables: ",
        paste(
          missing_columns,
          collapse = ", "
        )
      )
    )
  }


  rows_before <- nrow(data)


  # ----------------------------------------------------------
  # Extract and clean the required variables
  # ----------------------------------------------------------

  pas <- clean_code(
    data[[columns[["pas"]]]]
  )

  nic_code_clean <- clean_nic_code(
    data[[columns[["nic"]]]]
  )

  enterprise <- clean_code(
    data[[columns[["enterprise"]]]]
  )

  enterprise_size <- clean_code(
    data[[columns[["enterprise_size"]]]]
  )

  social_security <- clean_code(
    data[[columns[["social_security"]]]]
  )


  # ----------------------------------------------------------
  # Construct four- and five-digit NIC versions
  # ----------------------------------------------------------

  nic_code_digits <- ifelse(
    is.na(nic_code_clean),
    NA_integer_,
    nchar(nic_code_clean)
  )


  nic4 <- ifelse(
    is.na(nic_code_clean),
    NA_character_,
    substr(
      nic_code_clean,
      1,
      4
    )
  )


  nic5 <- ifelse(
    nic_code_digits == 5L,
    nic_code_clean,
    NA_character_
  )


  # ----------------------------------------------------------
  # Employment status
  # ----------------------------------------------------------

  employed_status <- pas %in% c(
    11,
    12,
    21,
    31,
    41,
    51
  )


  # ----------------------------------------------------------
  # Broad passenger-land-transport indicator
  #
  # Includes:
  #   buses
  #   taxis
  #   private-hire vehicles
  #   other passenger transport
  #   tramways
  #   metro/underground/elevated railways
  # ----------------------------------------------------------

  passenger_land_transport_broad <- case_when(

    is.na(pas) ~
      NA_integer_,

    !employed_status ~
      0L,

    employed_status &
      is.na(nic4) ~
      NA_integer_,

    nic4 %in% c(
      "4921",
      "4922"
    ) ~
      1L,

    TRUE ~
      0L
  )


  # ----------------------------------------------------------
  # Metro worker
  #
  # NIC 49213:
  # Urban/suburban underground or elevated railways
  #
  # A four-digit 4921 code is ambiguous because it could be:
  #   49211 bus
  #   49212 tram
  #   49213 metro
  #   49219 other passenger transport
  # ----------------------------------------------------------

  metro_worker_status <- case_when(

    is.na(pas) ~
      NA_integer_,

    !employed_status ~
      0L,

    employed_status &
      is.na(nic_code_clean) ~
      NA_integer_,

    nic5 == "49213" ~
      1L,

    nic_code_digits == 4L &
      nic4 == "4921" ~
      NA_integer_,

    TRUE ~
      0L
  )


  # ----------------------------------------------------------
  # Tram worker
  #
  # NIC 49212:
  # Urban/suburban tramways
  # ----------------------------------------------------------

  tram_worker_status <- case_when(

    is.na(pas) ~
      NA_integer_,

    !employed_status ~
      0L,

    employed_status &
      is.na(nic_code_clean) ~
      NA_integer_,

    nic5 == "49212" ~
      1L,

    nic_code_digits == 4L &
      nic4 == "4921" ~
      NA_integer_,

    TRUE ~
      0L
  )


  # ----------------------------------------------------------
  # Non-metro passenger-transport worker
  #
  # Included five-digit NIC codes:
  #
  # 49211 = urban/suburban passenger buses
  # 49219 = other urban/suburban passenger transport
  # 49221 = long-distance buses
  # 49222 = charter/coach services
  # 49223 = private cars with drivers
  # 49224 = taxi operation
  # 49225 = school/employee buses
  # 49226 = human/animal-drawn passenger vehicles
  # 49229 = other non-urban passenger transport
  #
  # Excluded:
  #
  # 49212 = tramways
  # 49213 = metro/underground/elevated railway
  # ----------------------------------------------------------

  non_metro_transport_codes <- c(
    "49211",
    "49219",
    "49221",
    "49222",
    "49223",
    "49224",
    "49225",
    "49226",
    "49229"
  )


  transport_status <- case_when(

    # PAS is unavailable
    is.na(pas) ~
      NA_integer_,


    # Known non-employed observation
    !employed_status ~
      0L,


    # Employed but NIC is unavailable or malformed
    employed_status &
      is.na(nic_code_clean) ~
      NA_integer_,


    # Exact five-digit non-metro transport code
    nic5 %in%
      non_metro_transport_codes ~
      1L,


    # Explicitly exclude tram and metro workers
    nic5 %in% c(
      "49212",
      "49213"
    ) ~
      0L,


    # Four-digit 4922 is safe to include because its
    # subclasses are non-metro passenger land transport
    nic_code_digits == 4L &
      nic4 == "4922" ~
      1L,


    # Four-digit 4921 cannot distinguish bus from metro/tram
    nic_code_digits == 4L &
      nic4 == "4921" ~
      NA_integer_,


    # Employed in another known industry
    TRUE ~
      0L
  )


  # ----------------------------------------------------------
  # Informal employment
  # ----------------------------------------------------------

  informal_status <- case_when(

    # PAS cannot be identified
    is.na(pas) ~
      NA_integer_,


    # --------------------------------------------------------
    # Wage or casual employees
    # --------------------------------------------------------

    pas %in% c(
      31,
      41,
      51
    ) &
      social_security == 8 ~
      1L,


    pas %in% c(
      31,
      41,
      51
    ) &
      social_security %in% 1:7 ~
      0L,


    # Employee but social-security status is unavailable
    pas %in% c(
      31,
      41,
      51
    ) ~
      NA_integer_,


    # --------------------------------------------------------
    # Unpaid household helpers
    # --------------------------------------------------------

    pas == 21 ~
      1L,


    # --------------------------------------------------------
    # Own-account workers and employers
    # --------------------------------------------------------

    pas %in% c(
      11,
      12
    ) &
      enterprise %in% c(
        1,
        2,
        3,
        4
      ) &
      enterprise_size %in% c(
        1,
        2
      ) ~
      1L,


    # Known enterprise type outside the informal definition
    pas %in% c(
      11,
      12
    ) &
      !is.na(enterprise) &
      !(enterprise %in% c(
        1,
        2,
        3,
        4
      )) ~
      0L,


    # Known enterprise size outside the informal definition
    pas %in% c(
      11,
      12
    ) &
      !is.na(enterprise_size) &
      !(enterprise_size %in% c(
        1,
        2
      )) ~
      0L,


    # Self-employed but insufficient information to classify
    pas %in% c(
      11,
      12
    ) ~
      NA_integer_,


    # Known non-employed status
    TRUE ~
      0L
  )


  # ----------------------------------------------------------
  # Informal non-metro passenger transport
  # ----------------------------------------------------------

  informal_transport_status <- case_when(

    # Definitely not a non-metro transport worker
    transport_status == 0L ~
      0L,

    # Definitely not informal
    informal_status == 0L ~
      0L,

    # Informal non-metro passenger transport worker
    transport_status == 1L &
      informal_status == 1L ~
      1L,

    # At least one component cannot be classified
    TRUE ~
      NA_integer_
  )


  # ----------------------------------------------------------
  # Add main outcome variables
  # ----------------------------------------------------------

  data$transport <-
    transport_status

  data$informal <-
    informal_status

  data$informal_transport <-
    informal_transport_status


  # ----------------------------------------------------------
  # Add separate transport-category audit variables
  # ----------------------------------------------------------

  data$passenger_land_transport_broad <-
    passenger_land_transport_broad

  data$metro_worker <-
    metro_worker_status

  data$tram_worker <-
    tram_worker_status


  # ----------------------------------------------------------
  # Unknown-classification audit variables
  # ----------------------------------------------------------

  data$transport_unknown <- as.integer(
    is.na(
      transport_status
    )
  )

  data$informal_unknown <- as.integer(
    is.na(
      informal_status
    )
  )

  data$informal_transport_unknown <- as.integer(
    is.na(
      informal_transport_status
    )
  )

  data$metro_worker_unknown <- as.integer(
    is.na(
      metro_worker_status
    )
  )


  # Specifically identifies ambiguous four-digit 4921 codes
  data$transport_scope_unknown <- as.integer(
    employed_status &
      nic_code_digits == 4L &
      nic4 == "4921"
  )

  data$transport_scope_unknown[
    is.na(
      data$transport_scope_unknown
    )
  ] <- 0L


  # ----------------------------------------------------------
  # Retain cleaned NIC variables for auditing
  # ----------------------------------------------------------

  data$nic_code_clean <-
    nic_code_clean

  data$nic_code_digits <-
    nic_code_digits

  data$nic4_clean <-
    nic4

  data$nic5_clean <-
    nic5


  # ----------------------------------------------------------
  # Confirm no respondent rows were added or removed
  # ----------------------------------------------------------

  if (nrow(data) != rows_before) {

    stop(
      paste(
        dataset_name,
        "row count changed while creating outcomes."
      )
    )
  }


  # ----------------------------------------------------------
  # Confirm binary variables contain only 0, 1 or NA
  # ----------------------------------------------------------

  binary_variables <- c(
    "transport",
    "informal",
    "informal_transport",
    "passenger_land_transport_broad",
    "metro_worker",
    "tram_worker",
    "transport_unknown",
    "informal_unknown",
    "informal_transport_unknown",
    "metro_worker_unknown",
    "transport_scope_unknown"
  )


  for (variable_name in binary_variables) {

    invalid_values <- unique(
      data[[variable_name]][
        !is.na(data[[variable_name]]) &
          !(data[[variable_name]] %in% c(
            0L,
            1L
          ))
      ]
    )

    if (length(invalid_values) > 0L) {

      stop(
        paste(
          dataset_name,
          "-",
          variable_name,
          "contains values other than 0, 1 or NA."
        )
      )
    }
  }


  data
}
```

Creating analysis dataset: Finding informal transport workers
```{r}
# ============================================================
# 3. MAP PLFS VARIABLES USED TO IDENTIFY
#    TRANSPORT AND INFORMAL EMPLOYMENT
# ============================================================

library(dplyr)
library(purrr)
library(tibble)


# ------------------------------------------------------------
# 2017-18
# ------------------------------------------------------------

columns17_18 <- c(
  pas = "b5pt1q3_per_fv",
  nic = "b5pt1q5_per_fv",
  enterprise = "b5pt1q9_per_fv",
  enterprise_size = "b5pt1q10_per_fv",
  social_security = "b5pt1q13_per_fv"
)


# ------------------------------------------------------------
# 2018-19
# ------------------------------------------------------------

columns18_19 <- c(
  pas = "b5pt1q3_per_fv",
  nic = "b5pt1q5_per_fv",
  enterprise = "b5pt1q9_per_fv",
  enterprise_size = "b5pt1q10_per_fv",
  social_security = "b5pt1q13_per_fv"
)


# ------------------------------------------------------------
# 2019-20
# ------------------------------------------------------------

columns19_20 <- c(
  pas = "b5pt1q3_per_fv",
  nic = "b5pt1q5_per_fv",
  enterprise = "b5pt1q9_per_fv",
  enterprise_size = "b5pt1q10_per_fv",
  social_security = "b5pt1q13_per_fv"
)


# ------------------------------------------------------------
# 2020-21
# ------------------------------------------------------------

columns20_21 <- c(
  pas = "b5pt1q3_perv1",
  nic = "b5pt1q5_perv1",
  enterprise = "b5pt1q9_perv1",
  enterprise_size = "b5pt1q10_perv1",
  social_security = "b5pt1q13_perv1"
)


# ------------------------------------------------------------
# July 2021-June 2022
#
# This section is used only if data21_22_fv has been imported.
# ------------------------------------------------------------

columns21_22 <- c(
  pas = "b5pt1q3_perv1",
  nic = "b5pt1q5_perv1",
  enterprise = "b5pt1q9_perv1",
  enterprise_size = "b5pt1q10_perv1",
  social_security = "b5pt1q13_perv1"
)


# ------------------------------------------------------------
# Calendar year 2022
# ------------------------------------------------------------

columns22 <- c(
  pas = "b5pt1q3_cperv1",
  nic = "b5pt1q5_cperv1",
  enterprise = "b5pt1q9_cperv1",
  enterprise_size = "b5pt1q10_cperv1",
  social_security = "b5pt1q13_cperv1"
)


# ------------------------------------------------------------
# Calendar year 2023
# ------------------------------------------------------------

columns23 <- c(
  pas = "b5pt1q3_cperv1",
  nic = "b5pt1q5_cperv1",
  enterprise = "b5pt1q9_cperv1",
  enterprise_size = "b5pt1q10_cperv1",
  social_security = "b5pt1q13_cperv1"
)


# ------------------------------------------------------------
# Calendar year 2024
# ------------------------------------------------------------

columns24 <- c(
  pas = "Principal_Status_Code",
  nic = "Principal_Industry_Code",
  enterprise = "Principal_Enterprise_Type",
  enterprise_size = "Principal_Workers_Count",
  social_security = "Principal_Social_Security"
)


# ------------------------------------------------------------
# Calendar year 2025
# ------------------------------------------------------------

columns25 <- c(
  pas = "pas",
  nic = "ind_pas",
  enterprise = "etyp_pas",
  enterprise_size = "wrkr_pas",
  social_security = "ssec_pas"
)


# ============================================================
# 4. CREATE A REGISTRY OF DATASETS AND VARIABLE MAPPINGS
# ============================================================

outcome_registry <- list(
  `2017-18` = list(
    data = data17_18_fv,
    columns = columns17_18,
    output_name = "analysis17_18"
  ),

  `2018-19` = list(
    data = data18_19_fv,
    columns = columns18_19,
    output_name = "analysis18_19"
  ),

  `2019-20` = list(
    data = data19_20_fv,
    columns = columns19_20,
    output_name = "analysis19_20"
  ),

  `2020-21` = list(
    data = data20_21_fv,
    columns = columns20_21,
    output_name = "analysis20_21"
  ),

  `2022` = list(
    data = data22,
    columns = columns22,
    output_name = "analysis22"
  ),

  `2023` = list(
    data = data23,
    columns = columns23,
    output_name = "analysis23"
  ),

  `2024` = list(
    data = data24,
    columns = columns24,
    output_name = "analysis24"
  ),

  `2025` = list(
    data = data25,
    columns = columns25,
    output_name = "analysis25"
  )
)


# Add the July 2021-June 2022 file if it has been imported
if (exists("data21_22_fv")) {

  outcome_registry <- append(
    outcome_registry,
    list(
      `2021-22` = list(
        data = data21_22_fv,
        columns = columns21_22,
        output_name = "analysis21_22"
      )
    ),
    after = 4
  )
}


# ============================================================
# 5. AUDIT THE VARIABLE MAPPINGS
# ============================================================

mapping_audit <- imap_dfr(
  outcome_registry,
  function(item, dataset_name) {

    tibble(
      dataset = dataset_name,
      variable_role = names(item$columns),
      source_column = unname(item$columns),
      column_exists = unname(item$columns) %in%
        names(item$data)
    )
  }
)


print(
  mapping_audit,
  n = Inf
)


missing_mapped_columns <- mapping_audit %>%
  filter(
    !column_exists
  )


if (nrow(missing_mapped_columns) > 0L) {

  print(
    missing_mapped_columns,
    n = Inf
  )

  stop(
    paste(
      "At least one mapped PLFS variable does not exist.",
      "Correct the source-column name before continuing."
    )
  )
}


# ============================================================
# 6. CREATE EACH ANNUAL ANALYSIS DATASET
# ============================================================

analysis_results <- imap(
  outcome_registry,
  function(item, dataset_name) {

    make_informal_transport(
      data = item$data,
      columns = item$columns,
      dataset_name = dataset_name
    )
  }
)


# Save the resulting datasets under their intended object names
walk2(
  analysis_results,
  outcome_registry,
  function(result, item) {

    assign(
      x = item$output_name,
      value = result,
      envir = .GlobalEnv
    )
  }
)


# ============================================================
# 7. AUDIT THE CREATED OUTCOMES
# ============================================================

analysis_outcome_audit <- imap_dfr(
  analysis_results,
  function(data, dataset_name) {

    tibble(
      dataset = dataset_name,

      observations = nrow(data),

      transport_workers = sum(
        data$transport == 1L,
        na.rm = TRUE
      ),

      informal_workers = sum(
        data$informal == 1L,
        na.rm = TRUE
      ),

      informal_transport_workers = sum(
        data$informal_transport == 1L,
        na.rm = TRUE
      ),

      transport_unknown = sum(
        data$transport_unknown == 1L,
        na.rm = TRUE
      ),

      informal_unknown = sum(
        data$informal_unknown == 1L,
        na.rm = TRUE
      ),

      informal_transport_unknown = sum(
        data$informal_transport_unknown == 1L,
        na.rm = TRUE
      )
    )
  }
) %>%
  mutate(
    transport_rate =
      transport_workers / observations,

    informal_rate =
      informal_workers / observations,

    informal_transport_rate =
      informal_transport_workers / observations,

    transport_unknown_rate =
      transport_unknown / observations,

    informal_unknown_rate =
      informal_unknown / observations,

    informal_transport_unknown_rate =
      informal_transport_unknown / observations
  )


print(
  analysis_outcome_audit,
  n = Inf
)


# ============================================================
# 8. STRICT CHECKS
# ============================================================

if (any(analysis_outcome_audit$observations == 0L)) {

  stop(
    "At least one annual analysis dataset has zero observations."
  )
}


if (any(
  analysis_outcome_audit$transport_unknown_rate > 0.10,
  na.rm = TRUE
)) {

  warning(
    paste(
      "At least one year has more than 10% unknown",
      "transport classifications."
    )
  )
}


if (any(
  analysis_outcome_audit$informal_unknown_rate > 0.10,
  na.rm = TRUE
)) {

  warning(
    paste(
      "At least one year has more than 10% unknown",
      "informality classifications."
    )
  )
}


message(
  "Outcome mappings and annual classifications completed."
)
```



Function for adding state identifiers for diff-in-diff
```{r}
# ============================================================
# 8. IMPORT AND ATTACH PANEL-SPECIFIC DISTRICT INFORMATION
#
# Matching hierarchy:
#
# 1. Match the state/district code against the correct PLFS
#    panel lookup.
#
# 2. For unmatched P4 codes only, search for the exact same
#    state and district code in the official 2025 lookup.
#
# 3. If still unmatched, retain the panel-specific code and
#    classify the district name as unresolved.
#
# No cross-year harmonised district identifier is created here.
# ============================================================

library(readxl)
library(dplyr)
library(purrr)
library(tibble)
library(here)
library(haven)


# ============================================================
# 8.1 AUTOMATICALLY IMPORT A DISTRICT LOOKUP
#
# This detects the row containing the actual column headers.
# It therefore works whether the workbook has title rows above
# the table or begins directly with the column headings.
# ============================================================

read_plfs_district_lookup <- function(
    path,
    lookup_name
) {

  if (!file.exists(path)) {

    stop(
      paste(
        lookup_name,
        "was not found at:",
        path
      )
    )
  }


  preview <- read_xlsx(
    path,
    col_names = FALSE,
    n_max = 25,
    .name_repair = "minimal"
  )


  preview_text <- apply(
    preview,
    1,
    function(row) {

      paste(
        tolower(
          trimws(
            as.character(row)
          )
        ),
        collapse = " | "
      )
    }
  )


  # The true header row should mention both state and district
  possible_header_rows <- which(
    grepl("state", preview_text) &
      grepl("district", preview_text)
  )


  if (length(possible_header_rows) == 0L) {

    print(preview)

    stop(
      paste(
        "Could not identify the header row in",
        lookup_name
      )
    )
  }


  header_row <- possible_header_rows[1]


  lookup_data <- read_xlsx(
    path,
    skip = header_row - 1L
  )


  # Remove completely empty columns
  non_empty_columns <- vapply(
    lookup_data,
    function(x) {

      any(
        !is.na(x) &
          trimws(
            as.character(x)
          ) != ""
      )
    },
    logical(1)
  )


  lookup_data <- lookup_data[
    ,
    non_empty_columns,
    drop = FALSE
  ]


  if (ncol(lookup_data) < 4L) {

    print(
      names(lookup_data)
    )

    stop(
      paste(
        lookup_name,
        "contains fewer than four usable columns."
      )
    )
  }


  message(
    lookup_name,
    ": header detected on row ",
    header_row,
    "; imported ",
    nrow(lookup_data),
    " rows."
  )


  lookup_data
}


# ============================================================
# 8.2 IMPORT THE FIVE LOOKUP WORKBOOKS
# ============================================================

panel1 <- read_plfs_district_lookup(
  path = here(
    "Data",
    "Crosswalks",
    "PLFS_District_Codes",
    "17-19.xlsx"
  ),
  lookup_name = "Panel 1 lookup"
)


panel2 <- read_plfs_district_lookup(
  path = here(
    "Data",
    "Crosswalks",
    "PLFS_District_Codes",
    "19-21.xlsx"
  ),
  lookup_name = "Panel 2 lookup"
)


panel3 <- read_plfs_district_lookup(
  path = here(
    "Data",
    "Crosswalks",
    "PLFS_District_Codes",
    "21-23.xlsx"
  ),
  lookup_name = "Panel 3 lookup"
)


panel4 <- read_plfs_district_lookup(
  path = here(
    "Data",
    "Crosswalks",
    "PLFS_District_Codes",
    "23-24.xlsx"
  ),
  lookup_name = "Panel 4 lookup"
)


panel2025 <- read_plfs_district_lookup(
  path = here(
    "Data",
    "Crosswalks",
    "PLFS_District_Codes",
    "2025.xlsx"
  ),
  lookup_name = "2025 lookup"
)


# ============================================================
# 8.3 CLEAN ONE DISTRICT LOOKUP
#
# The function searches the imported column names rather than
# assuming that the relevant variables are always columns 1–4.
# ============================================================

prepare_district_lookup <- function(
    lookup_data,
    panel_code,
    lookup_name
) {

  normalised_column_names <- tolower(
    gsub(
      "[^a-z0-9]+",
      " ",
      names(lookup_data)
    )
  )


  state_code_column <- which(
    grepl("state", normalised_column_names) &
      grepl("code", normalised_column_names)
  )


  district_code_column <- which(
    grepl("district", normalised_column_names) &
      grepl("code", normalised_column_names)
  )


  state_name_column <- which(
    grepl("state", normalised_column_names) &
      !grepl("code", normalised_column_names)
  )


  district_name_column <- which(
    grepl("district", normalised_column_names) &
      !grepl("code", normalised_column_names)
  )


  # If the names cannot be identified, fall back to the first
  # four columns in the expected order.
  if (
    length(state_code_column) == 0L ||
      length(state_name_column) == 0L ||
      length(district_code_column) == 0L ||
      length(district_name_column) == 0L
  ) {

    warning(
      paste(
        lookup_name,
        "column names could not be identified automatically;",
        "using the first four columns."
      )
    )

    selected_lookup <- lookup_data[, 1:4]

  } else {

    selected_lookup <- lookup_data[
      ,
      c(
        state_code_column[1],
        state_name_column[1],
        district_code_column[1],
        district_name_column[1]
      )
    ]
  }


  names(selected_lookup) <- c(
    "state_code_source",
    "State_Name_Panel",
    "district_code_source",
    "District_Name_Panel"
  )


  lookup_clean <- selected_lookup %>%
    mutate(
      state_code_numeric = clean_code(
        state_code_source
      ),

      district_code_numeric = clean_code(
        district_code_source
      )
    ) %>%

    # Remove blank rows, notes and footer text
    filter(
      !is.na(state_code_numeric),
      !is.na(district_code_numeric)
    )


  # ----------------------------------------------------------
  # Check that codes are whole numbers
  # ----------------------------------------------------------

  invalid_state_integer <-
    !is.na(lookup_clean$state_code_numeric) &
    abs(
      lookup_clean$state_code_numeric -
        round(lookup_clean$state_code_numeric)
    ) > 1e-8


  invalid_district_integer <-
    !is.na(lookup_clean$district_code_numeric) &
    abs(
      lookup_clean$district_code_numeric -
        round(lookup_clean$district_code_numeric)
    ) > 1e-8


  if (any(invalid_state_integer)) {

    stop(
      paste(
        lookup_name,
        "contains non-integer state codes."
      )
    )
  }


  if (any(invalid_district_integer)) {

    stop(
      paste(
        lookup_name,
        "contains non-integer district codes."
      )
    )
  }


  lookup_clean <- lookup_clean %>%
    transmute(
      panel_code = panel_code,

      state_code_std = sprintf(
        "%02d",
        as.integer(
          round(state_code_numeric)
        )
      ),

      district_code_std = sprintf(
        "%03d",
        as.integer(
          round(district_code_numeric)
        )
      ),

      State_Name_Panel = trimws(
        gsub(
          "\\s+",
          " ",
          as.character(State_Name_Panel)
        )
      ),

      District_Name_Panel = trimws(
        gsub(
          "\\s+",
          " ",
          as.character(District_Name_Panel)
        )
      )
    )


  # ----------------------------------------------------------
  # Check for missing names
  # ----------------------------------------------------------

  missing_names <- lookup_clean %>%
    filter(
      is.na(State_Name_Panel) |
        State_Name_Panel == "" |
        is.na(District_Name_Panel) |
        District_Name_Panel == ""
    )


  if (nrow(missing_names) > 0L) {

    print(
      missing_names,
      n = Inf
    )

    stop(
      paste(
        lookup_name,
        "contains missing state or district names."
      )
    )
  }


  # ----------------------------------------------------------
  # Each code must occur once within the lookup
  # ----------------------------------------------------------

  duplicate_keys <- lookup_clean %>%
    count(
      panel_code,
      state_code_std,
      district_code_std,
      name = "lookup_rows"
    ) %>%

    filter(
      lookup_rows > 1L
    )


  if (nrow(duplicate_keys) > 0L) {

    print(
      duplicate_keys,
      n = Inf
    )

    stop(
      paste(
        lookup_name,
        "contains duplicated state-district code keys."
      )
    )
  }


  lookup_clean
}


# ============================================================
# 8.4 PREPARE AND COMBINE THE LOOKUPS
# ============================================================

district_lookup_p1 <- prepare_district_lookup(
  lookup_data = panel1,
  panel_code = "P1",
  lookup_name = "Panel 1 lookup"
)


district_lookup_p2 <- prepare_district_lookup(
  lookup_data = panel2,
  panel_code = "P2",
  lookup_name = "Panel 2 lookup"
)


district_lookup_p3 <- prepare_district_lookup(
  lookup_data = panel3,
  panel_code = "P3",
  lookup_name = "Panel 3 lookup"
)


district_lookup_p4 <- prepare_district_lookup(
  lookup_data = panel4,
  panel_code = "P4",
  lookup_name = "Panel 4 lookup"
)


district_lookup_2025 <- prepare_district_lookup(
  lookup_data = panel2025,
  panel_code = "2025",
  lookup_name = "2025 lookup"
)


district_lookup_all <- bind_rows(
  district_lookup_p1,
  district_lookup_p2,
  district_lookup_p3,
  district_lookup_p4,
  district_lookup_2025
)


# 2025 lookup without panel_code, used only as an exact-code
# fallback for unmatched P4 observations
district_lookup_2025_backfill <- district_lookup_2025 %>%
  select(
    state_code_std,
    district_code_std,

    State_Name_2025_Backfill =
      State_Name_Panel,

    District_Name_2025_Backfill =
      District_Name_Panel
  )


duplicate_combined_keys <- district_lookup_all %>%
  count(
    panel_code,
    state_code_std,
    district_code_std,
    name = "lookup_rows"
  ) %>%

  filter(
    lookup_rows > 1L
  )


if (nrow(duplicate_combined_keys) > 0L) {

  print(
    duplicate_combined_keys,
    n = Inf
  )

  stop(
    "The combined district lookup contains duplicated keys."
  )
}


district_lookup_audit <- district_lookup_all %>%
  group_by(
    panel_code
  ) %>%

  summarise(
    states = n_distinct(
      state_code_std
    ),

    district_codes = n(),

    .groups = "drop"
  )


print(
  district_lookup_audit,
  n = Inf
)


# ============================================================
# 8.5 CLEAN THE PANEL VARIABLE IN CALENDAR-YEAR 2023
# ============================================================

clean_person_panel_code <- function(
    x,
    dataset_name = "dataset"
) {

  if (
    inherits(x, "haven_labelled") ||
      inherits(x, "labelled")
  ) {

    labelled_value <- tryCatch(
      as.character(
        haven::as_factor(x)
      ),

      error = function(e) {
        rep(
          NA_character_,
          length(x)
        )
      }
    )

    raw_value <- as.character(
      haven::zap_labels(x)
    )

  } else {

    labelled_value <- as.character(x)
    raw_value <- as.character(x)
  }


  labelled_value <- toupper(
    trimws(labelled_value)
  )

  raw_value <- toupper(
    trimws(raw_value)
  )


  panel_text <- ifelse(
    !is.na(labelled_value) &
      labelled_value != "",

    labelled_value,

    raw_value
  )


  panel_clean <- case_when(

    panel_text %in% c(
      "3",
      "P3",
      "PANEL 3",
      "PANEL3"
    ) ~
      "P3",

    panel_text %in% c(
      "4",
      "P4",
      "PANEL 4",
      "PANEL4"
    ) ~
      "P4",

    grepl(
      "PANEL[[:space:]]*3",
      panel_text
    ) ~
      "P3",

    grepl(
      "PANEL[[:space:]]*4",
      panel_text
    ) ~
      "P4",

    TRUE ~
      NA_character_
  )


  invalid_values <-
    !is.na(panel_text) &
    panel_text != "" &
    is.na(panel_clean)


  if (any(invalid_values)) {

    cat(
      "\nUnrecognised panel values in ",
      dataset_name,
      ":\n",
      sep = ""
    )

    print(
      unique(
        panel_text[
          invalid_values
        ]
      )
    )

    stop(
      paste(
        dataset_name,
        "contains unrecognised panel values."
      )
    )
  }


  panel_clean
}


# ============================================================
# 8.6 ATTACH DISTRICT INFORMATION TO ONE DATASET
# ============================================================

add_official_district_names <- function(
    data,
    state_column,
    district_column,
    lookup_data,
    lookup_2025_backfill,
    dataset_name = "dataset",
    fixed_panel = NULL,
    panel_column = NULL
) {

  required_columns <- c(
    state_column,
    district_column
  )


  if (!is.null(panel_column)) {

    required_columns <- c(
      required_columns,
      panel_column
    )
  }


  missing_columns <- setdiff(
    required_columns,
    names(data)
  )


  if (length(missing_columns) > 0L) {

    stop(
      paste0(
        dataset_name,
        " is missing: ",
        paste(
          missing_columns,
          collapse = ", "
        )
      )
    )
  }


  if (
    is.null(fixed_panel) &&
      is.null(panel_column)
  ) {

    stop(
      paste(
        dataset_name,
        "requires either fixed_panel or panel_column."
      )
    )
  }


  if (
    !is.null(fixed_panel) &&
      !is.null(panel_column)
  ) {

    stop(
      paste(
        dataset_name,
        "cannot use fixed_panel and panel_column simultaneously."
      )
    )
  }


  rows_before <- nrow(data)


  # ----------------------------------------------------------
  # Remove all geographic variables from previous runs
  # ----------------------------------------------------------

  geographic_variable_pattern <- paste0(
    "^(",
    paste(
      c(
        "panel_code",
        "state_code_std",
        "district_code_std",
        "district_id_raw",
        "District_ID_Panel",
        "State_Name",
        "District_Name",
        "State_Name_Panel",
        "District_Name_Panel",
        "State_Name_2025_Backfill",
        "District_Name_2025_Backfill",
        "State_Name_2025_Candidate",
        "District_Name_2025_Candidate",
        "State_Name_Working",
        "District_Name_Working",
        "State_Name_Key",
        "District_Name_Key",
        "District_ID",
        "District_Name_Source",
        "district_lookup_status",
        "district_lookup_matched",
        "district_panel_lookup_matched",
        "district_name_available"
      ),
      collapse = "|"
    ),
    ")(\\.x|\\.y)?$"
  )


  data <- data %>%
    select(
      -matches(
        geographic_variable_pattern
      )
    )


  # ----------------------------------------------------------
  # Clean the person-file codes
  # ----------------------------------------------------------

  state_code <- clean_code(
    data[[state_column]]
  )

  district_code <- clean_code(
    data[[district_column]]
  )


  invalid_state_integer <-
    !is.na(state_code) &
    abs(
      state_code -
        round(state_code)
    ) > 1e-8


  invalid_district_integer <-
    !is.na(district_code) &
    abs(
      district_code -
        round(district_code)
    ) > 1e-8


  if (any(invalid_state_integer)) {

    stop(
      paste(
        dataset_name,
        "contains non-integer state codes."
      )
    )
  }


  if (any(invalid_district_integer)) {

    stop(
      paste(
        dataset_name,
        "contains non-integer district codes."
      )
    )
  }


  state_code <- as.integer(
    round(state_code)
  )

  district_code <- as.integer(
    round(district_code)
  )


  # ----------------------------------------------------------
  # Determine each person's PLFS panel
  # ----------------------------------------------------------

  if (!is.null(fixed_panel)) {

    person_panel <- rep(
      fixed_panel,
      nrow(data)
    )

  } else {

    person_panel <- clean_person_panel_code(
      data[[panel_column]],
      dataset_name = dataset_name
    )
  }


  # ----------------------------------------------------------
  # Construct panel-specific identifiers
  # ----------------------------------------------------------

  data <- data %>%
    mutate(
      .row_order_geography = row_number(),

      panel_code = person_panel,

      state_code_std = if_else(
        is.na(state_code),
        NA_character_,
        sprintf(
          "%02d",
          state_code
        )
      ),

      district_code_std = if_else(
        is.na(district_code),
        NA_character_,
        sprintf(
          "%03d",
          district_code
        )
      ),

      district_id_raw = if_else(
        is.na(state_code_std) |
          is.na(district_code_std),

        NA_character_,

        paste0(
          state_code_std,
          "_",
          district_code_std
        )
      ),

      District_ID_Panel = if_else(
        is.na(panel_code) |
          is.na(state_code_std) |
          is.na(district_code_std),

        NA_character_,

        paste(
          panel_code,
          state_code_std,
          district_code_std,
          sep = "_"
        )
      )
    )


  missing_geography <- data %>%
    filter(
      is.na(panel_code) |
        is.na(state_code_std) |
        is.na(district_code_std)
    )


  if (nrow(missing_geography) > 0L) {

    print(
      missing_geography %>%
        count(
          panel_code,
          state_code_std,
          district_code_std,
          name = "person_observations"
        ),
      n = Inf
    )

    stop(
      paste(
        dataset_name,
        "contains missing panel, state or district codes."
      )
    )
  }


  # ----------------------------------------------------------
  # First join: correct panel-specific lookup
  # ----------------------------------------------------------

  data <- data %>%
    left_join(
      lookup_data,
      by = c(
        "panel_code",
        "state_code_std",
        "district_code_std"
      )
    )


  if (nrow(data) != rows_before) {

    stop(
      paste(
        dataset_name,
        "row count changed during the panel lookup join."
      )
    )
  }


  # ----------------------------------------------------------
  # Second join: exact-code 2025 fallback
  #
  # This fallback is only used when:
  #
  #   panel_code == P4
  #   no Panel 4 name was found
  #   the exact same state/district code exists in 2025
  # ----------------------------------------------------------

  data <- data %>%
    left_join(
      lookup_2025_backfill,
      by = c(
        "state_code_std",
        "district_code_std"
      )
    )


  if (nrow(data) != rows_before) {

    stop(
      paste(
        dataset_name,
        "row count changed during the 2025 backfill join."
      )
    )
  }


  # ----------------------------------------------------------
  # Classify the source and reliability of the district name
  # ----------------------------------------------------------

  data <- data %>%
    mutate(
      district_lookup_status = case_when(

        !is.na(State_Name_Panel) &
          !is.na(District_Name_Panel) ~
          "matched_panel_lookup",

        panel_code == "P4" &
          is.na(District_Name_Panel) &
          !is.na(State_Name_2025_Backfill) &
          !is.na(District_Name_2025_Backfill) ~
          "backfilled_from_2025_exact_code",

        TRUE ~
          "unresolved"
      ),


      State_Name = case_when(

        district_lookup_status ==
          "matched_panel_lookup" ~
          State_Name_Panel,

        district_lookup_status ==
          "backfilled_from_2025_exact_code" ~
          State_Name_2025_Backfill,

        TRUE ~
          NA_character_
      ),


      District_Name = case_when(

        district_lookup_status ==
          "matched_panel_lookup" ~
          District_Name_Panel,

        district_lookup_status ==
          "backfilled_from_2025_exact_code" ~
          District_Name_2025_Backfill,

        TRUE ~
          NA_character_
      ),


      District_Name_Source = case_when(

        district_lookup_status ==
          "matched_panel_lookup" ~
          paste0(
            panel_code,
            " panel lookup"
          ),

        district_lookup_status ==
          "backfilled_from_2025_exact_code" ~
          "2025 exact-code backfill",

        TRUE ~
          "unresolved"
      ),


      # Direct panel match only
      district_panel_lookup_matched =
        as.integer(
          district_lookup_status ==
            "matched_panel_lookup"
        ),


      # A usable name is available either directly or by backfill
      district_lookup_matched =
        as.integer(
          district_lookup_status !=
            "unresolved"
        ),


      district_name_available =
        as.integer(
          !is.na(State_Name) &
            !is.na(District_Name)
        )
    )


  # ----------------------------------------------------------
  # Unmatched codes outside P4 indicate a genuine error
  # ----------------------------------------------------------

  unexpected_unresolved <- data %>%
    filter(
      district_lookup_status == "unresolved",
      panel_code != "P4"
    ) %>%

    count(
      panel_code,
      state_code_std,
      district_code_std,
      District_ID_Panel,
      name = "person_observations"
    )


  if (nrow(unexpected_unresolved) > 0L) {

    print(
      unexpected_unresolved,
      n = Inf
    )

    stop(
      paste(
        dataset_name,
        "contains unresolved district codes outside Panel 4."
      )
    )
  }


  # ----------------------------------------------------------
  # P4 unresolved codes are retained and reported
  # ----------------------------------------------------------

  unresolved_p4 <- data %>%
    filter(
      district_lookup_status == "unresolved",
      panel_code == "P4"
    ) %>%

    count(
      panel_code,
      state_code_std,
      district_code_std,
      District_ID_Panel,
      name = "person_observations"
    ) %>%

    arrange(
      state_code_std,
      district_code_std
    )


  if (nrow(unresolved_p4) > 0L) {

    cat(
      "\nUnresolved Panel 4 district codes in ",
      dataset_name,
      ":\n",
      sep = ""
    )

    print(
      unresolved_p4,
      n = Inf
    )

    warning(
      paste(
        dataset_name,
        "contains unresolved Panel 4 district codes.",
        "They were retained with missing district names."
      )
    )
  }


  data <- data %>%
    arrange(
      .row_order_geography
    ) %>%

    select(
      -.row_order_geography
    )


  if (nrow(data) != rows_before) {

    stop(
      paste(
        dataset_name,
        "row count changed while attaching district information."
      )
    )
  }


  data
}


# ============================================================
# 8.7 MAP EACH ANALYSIS DATASET TO ITS PANEL
# ============================================================

district_registry <- list(
  `2017-18` = list(
    object_name = "analysis17_18",
    state_column = "state_per_fv",
    district_column = "b1q4_per_fv",
    fixed_panel = "P1",
    panel_column = NULL
  ),

  `2018-19` = list(
    object_name = "analysis18_19",
    state_column = "state_per_fv",
    district_column = "b1q4_per_fv",
    fixed_panel = "P1",
    panel_column = NULL
  ),

  `2019-20` = list(
    object_name = "analysis19_20",
    state_column = "state_per_fv",
    district_column = "b1q4_per_fv",
    fixed_panel = "P2",
    panel_column = NULL
  ),

  `2020-21` = list(
    object_name = "analysis20_21",
    state_column = "state_perv1",
    district_column = "distcode_perv1",
    fixed_panel = "P2",
    panel_column = NULL
  ),

  `2022` = list(
    object_name = "analysis22",
    state_column = "state_cperv1",
    district_column = "distcode_cperv1",
    fixed_panel = "P3",
    panel_column = NULL
  ),

  # Calendar-year 2023 contains both P3 and P4 observations
  `2023` = list(
    object_name = "analysis23",
    state_column = "state_cperv1",
    district_column = "distcode_cperv1",
    fixed_panel = NULL,
    panel_column = "panel_cperv1"
  ),

  `2024` = list(
    object_name = "analysis24",
    state_column = "State_UT_Code",
    district_column = "District_Code",
    fixed_panel = "P4",
    panel_column = NULL
  ),

  `2025` = list(
    object_name = "analysis25",
    state_column = "st",
    district_column = "dc",
    fixed_panel = "2025",
    panel_column = NULL
  )
)


if (exists(
  "analysis21_22",
  envir = .GlobalEnv
)) {

  district_registry <- append(
    district_registry,

    list(
      `2021-22` = list(
        object_name = "analysis21_22",
        state_column = "state_perv1",
        district_column = "distcode_perv1",
        fixed_panel = "P3",
        panel_column = NULL
      )
    ),

    after = 4L
  )
}


# ============================================================
# 8.8 CHECK DATASETS AND SOURCE VARIABLES
# ============================================================

district_mapping_audit <- imap_dfr(
  district_registry,
  function(item, dataset_name) {

    object_exists <- exists(
      item$object_name,
      envir = .GlobalEnv
    )


    if (object_exists) {

      current_data <- get(
        item$object_name,
        envir = .GlobalEnv
      )

      state_column_exists <-
        item$state_column %in%
        names(current_data)

      district_column_exists <-
        item$district_column %in%
        names(current_data)

      panel_column_exists <-
        is.null(item$panel_column) ||
        item$panel_column %in%
          names(current_data)

    } else {

      state_column_exists <- FALSE
      district_column_exists <- FALSE
      panel_column_exists <- FALSE
    }


    tibble(
      dataset = dataset_name,
      object_name = item$object_name,
      object_exists = object_exists,
      state_column = item$state_column,
      state_column_exists = state_column_exists,
      district_column = item$district_column,
      district_column_exists = district_column_exists,

      fixed_panel = ifelse(
        is.null(item$fixed_panel),
        NA_character_,
        item$fixed_panel
      ),

      panel_column = ifelse(
        is.null(item$panel_column),
        NA_character_,
        item$panel_column
      ),

      panel_column_exists = panel_column_exists
    )
  }
)


print(
  district_mapping_audit,
  n = Inf
)


invalid_mappings <- district_mapping_audit %>%
  filter(
    !object_exists |
      !state_column_exists |
      !district_column_exists |
      !panel_column_exists
  )


if (nrow(invalid_mappings) > 0L) {

  print(
    invalid_mappings,
    n = Inf
  )

  stop(
    "At least one dataset or geographic source variable is missing."
  )
}


# ============================================================
# 8.9 APPLY THE LOOKUPS ONE DATASET AT A TIME
#
# gc() is run between years to reduce memory pressure.
# ============================================================

for (dataset_name in names(district_registry)) {

  item <- district_registry[[dataset_name]]


  message(
    "Processing district information for ",
    dataset_name,
    "..."
  )


  current_data <- get(
    item$object_name,
    envir = .GlobalEnv
  )


  updated_data <- add_official_district_names(
    data = current_data,
    state_column = item$state_column,
    district_column = item$district_column,
    lookup_data = district_lookup_all,
    lookup_2025_backfill =
      district_lookup_2025_backfill,
    dataset_name = dataset_name,
    fixed_panel = item$fixed_panel,
    panel_column = item$panel_column
  )


  assign(
    item$object_name,
    updated_data,
    envir = .GlobalEnv
  )


  rm(
    current_data,
    updated_data
  )

  gc()


  message(
    "Finished ",
    dataset_name
  )
}


# ============================================================
# 8.10 AUDIT THE RESULTS
# ============================================================

district_id_audit <- imap_dfr(
  district_registry,
  function(item, dataset_name) {

    current_data <- get(
      item$object_name,
      envir = .GlobalEnv
    )


    current_data %>%
      group_by(
        panel_code
      ) %>%

      summarise(
        dataset = dataset_name,

        observations = n(),

        unique_states = n_distinct(
          state_code_std
        ),

        unique_panel_districts = n_distinct(
          District_ID_Panel
        ),

        matched_panel_observations = sum(
          district_lookup_status ==
            "matched_panel_lookup"
        ),

        backfilled_observations = sum(
          district_lookup_status ==
            "backfilled_from_2025_exact_code"
        ),

        unresolved_observations = sum(
          district_lookup_status ==
            "unresolved"
        ),

        .groups = "drop"
      ) %>%

      select(
        dataset,
        everything()
      )
  }
)


print(
  district_id_audit,
  n = Inf
)


# ============================================================
# 8.11 CREATE A UNIQUE DISTRICT INVENTORY
# ============================================================

district_inventory_by_year <- imap_dfr(
  district_registry,
  function(item, dataset_name) {

    current_data <- get(
      item$object_name,
      envir = .GlobalEnv
    )


    current_data %>%
      count(
        panel_code,
        state_code_std,
        district_code_std,
        District_ID_Panel,
        State_Name,
        District_Name,
        District_Name_Source,
        district_lookup_status,
        name = "person_observations"
      ) %>%

      mutate(
        dataset = dataset_name,
        .before = 1
      )
  }
) %>%

  arrange(
    factor(
      dataset,
      levels = names(
        district_registry
      )
    ),
    panel_code,
    state_code_std,
    district_code_std
  )


# Separate table containing only unresolved districts
unresolved_district_inventory <- district_inventory_by_year %>%
  filter(
    district_lookup_status ==
      "unresolved"
  )


print(
  unresolved_district_inventory,
  n = Inf
)


# ============================================================
# 8.12 FINAL CHECKS
# ============================================================

missing_panel_ids <- imap_int(
  district_registry,
  function(item, dataset_name) {

    current_data <- get(
      item$object_name,
      envir = .GlobalEnv
    )

    sum(
      is.na(
        current_data$District_ID_Panel
      )
    )
  }
)


if (any(missing_panel_ids > 0L)) {

  stop(
    "At least one observation has a missing panel-specific district ID."
  )
}


total_unresolved <- sum(
  district_id_audit$unresolved_observations
)


if (total_unresolved > 0L) {

  warning(
    paste(
      total_unresolved,
      "person observations have unresolved district names.",
      "Their District_ID_Panel values were retained."
    )
  )
}


message(
  paste(
    "Panel-specific district information attached.",
    "Unmatched P4 codes were backfilled from the exact 2025",
    "state-district code where possible."
  )
)
```

Qurter and month insertion:
```{r}
# ============================================================
# Validate a scalar calendar year
# ============================================================

validate_calendar_year <- function(x, argument_name = "calendar_year") {

  if (length(x) != 1L) {
    stop(
      argument_name,
      " must contain exactly one value."
    )
  }

  year <- suppressWarnings(
    as.integer(x)
  )

  if (
    is.na(year) ||
    year < 1900L ||
    year > 2100L
  ) {
    stop(
      argument_name,
      " must be a valid four-digit year."
    )
  }

  year
}


# ============================================================
# Strictly extract PLFS quarter codes Q1-Q8
# ============================================================

get_plfs_quarter_number <- function(
    x,
    strict = TRUE
) {

  is_labelled <- (
    inherits(x, "haven_labelled") ||
    inherits(x, "labelled")
  )

  # Use labels where available, but retain unlabelled values
  if (is_labelled) {

    text_value <- as.character(
      haven::as_factor(
        x,
        levels = "default"
      )
    )

    raw_value <- haven::zap_labels(x)

  } else {

    text_value <- as.character(x)
    raw_value <- x
  }


  text_value <- toupper(
    trimws(text_value)
  )


  quarter_number <- rep(
    NA_integer_,
    length(text_value)
  )


  # Accept exact values such as Q1, Q2, ..., Q8
  q_format <- grepl(
    "^Q\\s*[1-8]$",
    text_value
  )

  quarter_number[q_format] <- as.integer(
    sub(
      "^Q\\s*",
      "",
      text_value[q_format]
    )
  )


  # Also accept exact numeric strings 1, 2, ..., 8
  numeric_format <- grepl(
    "^[1-8]$",
    text_value
  )

  quarter_number[numeric_format] <- as.integer(
    text_value[numeric_format]
  )


  # For labelled numeric vectors, use the underlying value
  # when the displayed label could not be parsed
  raw_number <- suppressWarnings(
    as.integer(
      as.character(raw_value)
    )
  )

  use_raw_number <- (
    is.na(quarter_number) &
    raw_number %in% 1:8
  )

  quarter_number[use_raw_number] <- raw_number[use_raw_number]


  # Detect nonmissing values that could not be parsed
  invalid_value <- (
    !is.na(text_value) &
    nzchar(text_value) &
    is.na(quarter_number)
  )


  if (any(invalid_value)) {

    examples <- paste(
      utils::head(
        unique(text_value[invalid_value]),
        5L
      ),
      collapse = ", "
    )

    message <- paste0(
      "Unrecognised PLFS quarter values: ",
      examples
    )

    if (strict) {
      stop(message)
    } else {
      warning(
        message,
        call. = FALSE
      )
    }
  }


  quarter_number
}


# ============================================================
# Extract month numbers 1-12
#
# Accepts:
# 1, 01, M1, M01, January, Jan, etc.
# ============================================================

get_plfs_month_number <- function(
    x,
    strict = TRUE
) {

  is_labelled <- (
    inherits(x, "haven_labelled") ||
    inherits(x, "labelled")
  )

  if (is_labelled) {

    text_value <- as.character(
      haven::as_factor(
        x,
        levels = "default"
      )
    )

    raw_value <- haven::zap_labels(x)

  } else {

    text_value <- as.character(x)
    raw_value <- x
  }


  text_value <- toupper(
    trimws(text_value)
  )


  month_number <- rep(
    NA_integer_,
    length(text_value)
  )


  # Numeric month: 1, 01, ..., 12
  numeric_format <- grepl(
    "^(0?[1-9]|1[0-2])$",
    text_value
  )

  month_number[numeric_format] <- as.integer(
    text_value[numeric_format]
  )


  # M1, M01, ..., M12
  m_format <- grepl(
    "^M\\s*(0?[1-9]|1[0-2])$",
    text_value
  )

  month_number[m_format] <- as.integer(
    sub(
      "^M\\s*",
      "",
      text_value[m_format]
    )
  )


  # Full and abbreviated English month names
  month_lookup <- c(
    setNames(1:12, toupper(month.name)),
    setNames(1:12, toupper(month.abb))
  )

  named_month <- match(
    text_value,
    names(month_lookup)
  )

  use_named_month <- (
    is.na(month_number) &
    !is.na(named_month)
  )

  month_number[use_named_month] <- unname(
    month_lookup[named_month[use_named_month]]
  )


  # Underlying numeric fallback
  raw_number <- suppressWarnings(
    as.integer(
      as.character(raw_value)
    )
  )

  use_raw_number <- (
    is.na(month_number) &
    raw_number %in% 1:12
  )

  month_number[use_raw_number] <- raw_number[use_raw_number]


  invalid_value <- (
    !is.na(text_value) &
    nzchar(text_value) &
    is.na(month_number)
  )


  if (any(invalid_value)) {

    examples <- paste(
      utils::head(
        unique(text_value[invalid_value]),
        5L
      ),
      collapse = ", "
    )

    message <- paste0(
      "Unrecognised PLFS month values: ",
      examples
    )

    if (strict) {
      stop(message)
    } else {
      warning(
        message,
        call. = FALSE
      )
    }
  }


  month_number
}


# ============================================================
# Legacy Q1-Q8 to calendar quarter
# ============================================================

convert_legacy_quarter <- function(q_number) {

  calendar_quarter <- rep(
    NA_integer_,
    length(q_number)
  )

  calendar_quarter[q_number %in% c(1L, 5L)] <- 3L
  calendar_quarter[q_number %in% c(2L, 6L)] <- 4L
  calendar_quarter[q_number %in% c(3L, 7L)] <- 1L
  calendar_quarter[q_number %in% c(4L, 8L)] <- 2L

  calendar_quarter
}


# ============================================================
# Add common timestamp variables
# ============================================================

finalise_plfs_time <- function(
    data,
    calendar_year,
    calendar_quarter,
    survey_month = NULL,
    plfs_quarter_code = NULL
) {

  number_of_rows <- nrow(data)


  calendar_year <- rep_len(
    as.integer(calendar_year),
    number_of_rows
  )

  calendar_quarter <- rep_len(
    as.integer(calendar_quarter),
    number_of_rows
  )


  if (is.null(survey_month)) {

    survey_month <- rep(
      NA_integer_,
      number_of_rows
    )

  } else {

    survey_month <- rep_len(
      as.integer(survey_month),
      number_of_rows
    )
  }


  if (is.null(plfs_quarter_code)) {

    plfs_quarter_code <- rep(
      NA_integer_,
      number_of_rows
    )

  } else {

    plfs_quarter_code <- rep_len(
      as.integer(plfs_quarter_code),
      number_of_rows
    )
  }


  invalid_quarter <- (
    !is.na(calendar_quarter) &
    !(calendar_quarter %in% 1:4)
  )

  if (any(invalid_quarter)) {
    stop(
      "Calendar quarter must be between 1 and 4."
    )
  }


  invalid_month <- (
    !is.na(survey_month) &
    !(survey_month %in% 1:12)
  )

  if (any(invalid_month)) {
    stop(
      "Survey month must be between 1 and 12."
    )
  }


  # Check that month and quarter agree
  quarter_from_month <- rep(
    NA_integer_,
    number_of_rows
  )

  has_month <- !is.na(survey_month)

  quarter_from_month[has_month] <- (
    (survey_month[has_month] - 1L) %/% 3L
  ) + 1L


  inconsistent_time <- (
    !is.na(quarter_from_month) &
    !is.na(calendar_quarter) &
    quarter_from_month != calendar_quarter
  )

  if (any(inconsistent_time)) {
    stop(
      "Survey month and calendar quarter disagree."
    )
  }


  data$calendar_year <- calendar_year
  data$calendar_quarter <- calendar_quarter
  data$survey_month <- survey_month
  data$plfs_quarter_code <- plfs_quarter_code


  # First month of the quarter.
  # This is not necessarily the actual survey month.
  data$quarter_start_month <- rep(
    NA_integer_,
    number_of_rows
  )

  valid_quarter <- (
    !is.na(calendar_year) &
    !is.na(calendar_quarter)
  )

  data$quarter_start_month[valid_quarter] <- (
    calendar_quarter[valid_quarter] - 1L
  ) * 3L + 1L


  data$year_quarter <- rep(
    NA_character_,
    number_of_rows
  )

  data$year_quarter[valid_quarter] <- sprintf(
    "%04d_Q%d",
    calendar_year[valid_quarter],
    calendar_quarter[valid_quarter]
  )


  # Continuous quarterly index
  data$quarter_id <- rep(
    NA_integer_,
    number_of_rows
  )

  data$quarter_id[valid_quarter] <- as.integer(
    calendar_year[valid_quarter] * 4L +
      calendar_quarter[valid_quarter] -
      1L
  )


  data$quarter_start_date <- rep(
    as.Date(NA),
    number_of_rows
  )

  data$quarter_start_date[valid_quarter] <- as.Date(
    sprintf(
      "%04d-%02d-01",
      calendar_year[valid_quarter],
      data$quarter_start_month[valid_quarter]
    )
  )


  valid_month <- (
    !is.na(calendar_year) &
    !is.na(survey_month)
  )


  data$year_month <- rep(
    NA_character_,
    number_of_rows
  )

  data$year_month[valid_month] <- sprintf(
    "%04d_%02d",
    calendar_year[valid_month],
    survey_month[valid_month]
  )


  data$month_id <- rep(
    NA_integer_,
    number_of_rows
  )

  data$month_id[valid_month] <- as.integer(
    calendar_year[valid_month] * 12L +
      survey_month[valid_month] -
      1L
  )


  data$month_start_date <- rep(
    as.Date(NA),
    number_of_rows
  )

  data$month_start_date[valid_month] <- as.Date(
    sprintf(
      "%04d-%02d-01",
      calendar_year[valid_month],
      survey_month[valid_month]
    )
  )


  data
}


# ============================================================
# July-June annual PLFS files
# ============================================================

add_annual_quarter <- function(
    data,
    quarter_column,
    start_year
) {

  if (!(quarter_column %in% names(data))) {
    stop(
      "Quarter column not found: ",
      quarter_column
    )
  }

  start_year <- validate_calendar_year(
    start_year,
    "start_year"
  )


  plfs_quarter_code <- get_plfs_quarter_number(
    data[[quarter_column]]
  )


  calendar_quarter <- convert_legacy_quarter(
    plfs_quarter_code
  )


  calendar_year <- rep(
    NA_integer_,
    length(plfs_quarter_code)
  )


  first_half <- plfs_quarter_code %in% c(
    1L, 2L, 5L, 6L
  )

  second_half <- plfs_quarter_code %in% c(
    3L, 4L, 7L, 8L
  )


  calendar_year[first_half] <- start_year
  calendar_year[second_half] <- start_year + 1L


  finalise_plfs_time(
    data = data,
    calendar_year = calendar_year,
    calendar_quarter = calendar_quarter,
    plfs_quarter_code = plfs_quarter_code
  )
}


# ============================================================
# Legacy calendar-year PLFS files
# ============================================================

add_calendar_year_quarter <- function(
    data,
    quarter_column,
    calendar_year
) {

  if (!(quarter_column %in% names(data))) {
    stop(
      "Quarter column not found: ",
      quarter_column
    )
  }

  calendar_year <- validate_calendar_year(
    calendar_year
  )


  plfs_quarter_code <- get_plfs_quarter_number(
    data[[quarter_column]]
  )


  calendar_quarter <- convert_legacy_quarter(
    plfs_quarter_code
  )


  finalise_plfs_time(
    data = data,
    calendar_year = calendar_year,
    calendar_quarter = calendar_quarter,
    plfs_quarter_code = plfs_quarter_code
  )
}


# ============================================================
# Monthly calendar-year data
#
# Month is the source of truth.
# Quarter is derived from month.
# ============================================================

add_calendar_month <- function(
    data,
    month_column,
    calendar_year,
    quarter_column = NULL
) {

  if (!(month_column %in% names(data))) {
    stop(
      "Month column not found: ",
      month_column
    )
  }

  calendar_year <- validate_calendar_year(
    calendar_year
  )


  survey_month <- get_plfs_month_number(
    data[[month_column]]
  )


  calendar_quarter <- rep(
    NA_integer_,
    length(survey_month)
  )

  valid_month <- !is.na(survey_month)

  calendar_quarter[valid_month] <- (
    (survey_month[valid_month] - 1L) %/% 3L
  ) + 1L


  plfs_quarter_code <- rep(
    NA_integer_,
    length(survey_month)
  )


  # Optionally validate the supplied direct quarter code
  if (!is.null(quarter_column)) {

    if (!(quarter_column %in% names(data))) {
      stop(
        "Quarter column not found: ",
        quarter_column
      )
    }

    plfs_quarter_code <- get_plfs_quarter_number(
      data[[quarter_column]]
    )


    invalid_direct_quarter <- (
      !is.na(plfs_quarter_code) &
      !(plfs_quarter_code %in% 1:4)
    )

    if (any(invalid_direct_quarter)) {
      stop(
        "Monthly-format quarter codes must be Q1-Q4."
      )
    }


    mismatch <- (
      !is.na(plfs_quarter_code) &
      !is.na(calendar_quarter) &
      plfs_quarter_code != calendar_quarter
    )

    if (any(mismatch)) {
      stop(
        "The supplied month and quarter columns disagree."
      )
    }
  }


  finalise_plfs_time(
    data = data,
    calendar_year = calendar_year,
    calendar_quarter = calendar_quarter,
    survey_month = survey_month,
    plfs_quarter_code = plfs_quarter_code
  )
}


# ============================================================
# 2025 wrapper
# ============================================================

add_2025_time <- function(
    data,
    month_column = "month",
    quarter_column = "qtr"
) {

  add_calendar_month(
    data = data,
    month_column = month_column,
    calendar_year = 2025L,
    quarter_column = quarter_column
  )
}


```



Functions for quarter and year identifiers, so they can be in chronological order.
```{r}
# ============================================================
# Extract the quarter number from values such as Q1, Q2, etc.
# ============================================================

get_plfs_quarter_number <- function(x) {

  # If the quarter variable is Stata-labelled,
  # convert it to the displayed label first
  if (
    inherits(x, "haven_labelled") ||
    inherits(x, "labelled")
  ) {
    x <- haven::as_factor(
      x,
      levels = "labels"
    )
  }

  x <- trimws(as.character(x))

  # Extract the numeric part:
  # Q1 -> 1, Q7 -> 7
  suppressWarnings(
    as.integer(
      gsub("[^0-9]", "", x)
    )
  )
}


# ============================================================
# Legacy PLFS quarter conversion
#
# Q1/Q5 = July-September      = calendar quarter 3
# Q2/Q6 = October-December   = calendar quarter 4
# Q3/Q7 = January-March      = calendar quarter 1
# Q4/Q8 = April-June         = calendar quarter 2
# ============================================================

convert_legacy_quarter <- function(q_number) {

  quarter <- rep(
    NA_integer_,
    length(q_number)
  )

  quarter[q_number %in% c(1, 5)] <- 3L
  quarter[q_number %in% c(2, 6)] <- 4L
  quarter[q_number %in% c(3, 7)] <- 1L
  quarter[q_number %in% c(4, 8)] <- 2L

  quarter
}

add_annual_quarter <- function(
    data,
    quarter_column,
    start_year
) {

  # Check that the quarter column exists
  if (!(quarter_column %in% names(data))) {
    stop(
      paste(
        "Quarter column not found:",
        quarter_column
      )
    )
  }


  # Read the original PLFS quarter code
  plfs_quarter_number <- get_plfs_quarter_number(
    data[[quarter_column]]
  )


  # Convert Q1-Q8 into calendar quarters 1-4
  data$quarter <- convert_legacy_quarter(
    plfs_quarter_number
  )


  # July-December belong to the starting year.
  # January-June belong to the following year.
  data$year <- ifelse(
    plfs_quarter_number %in% c(3, 4, 7, 8),
    start_year + 1L,
    start_year
  )


  # Convert year back to integer because ifelse can alter type
  data$year <- as.integer(data$year)


  # Create a readable period label
  data$year_quarter <- ifelse(
    is.na(data$year) | is.na(data$quarter),
    NA_character_,
    paste0(
      data$year,
      "_Q",
      data$quarter
    )
  )


  # Numeric quarterly index:
  # useful for sorting observations chronologically
  data$period_index <- ifelse(
    is.na(data$year) | is.na(data$quarter),
    NA_integer_,
    data$year * 4L + data$quarter
  )


  # Retain the original PLFS quarter number for auditing
  data$plfs_quarter_code <- plfs_quarter_number


  data
}

add_calendar_year_quarter <- function(
    data,
    quarter_column,
    calendar_year
) {

  if (!(quarter_column %in% names(data))) {
    stop(
      paste(
        "Quarter column not found:",
        quarter_column
      )
    )
  }


  plfs_quarter_number <- get_plfs_quarter_number(
    data[[quarter_column]]
  )


  # The year is fixed because this is a calendar-year file
  data$year <- as.integer(calendar_year)


  # Convert the rolling Q1-Q8 code to calendar quarter 1-4
  data$quarter <- convert_legacy_quarter(
    plfs_quarter_number
  )


  data$year_quarter <- ifelse(
    is.na(data$quarter),
    NA_character_,
    paste0(
      data$year,
      "_Q",
      data$quarter
    )
  )


  data$period_index <- ifelse(
    is.na(data$quarter),
    NA_integer_,
    data$year * 4L + data$quarter
  )


  data$plfs_quarter_code <- plfs_quarter_number


  data
}

add_2025_quarter <- function(
    data,
    quarter_column = "qtr"
) {

  if (!(quarter_column %in% names(data))) {
    stop(
      paste(
        "Quarter column not found:",
        quarter_column
      )
    )
  }


  quarter_number <- get_plfs_quarter_number(
    data[[quarter_column]]
  )


  # In 2025, Q1-Q4 correspond directly to calendar quarters
  valid_quarter <- quarter_number %in% 1:4

  data$year <- 2025L

  data$quarter <- ifelse(
    valid_quarter,
    quarter_number,
    NA_integer_
  )

  data$quarter <- as.integer(
    data$quarter
  )


  data$year_quarter <- ifelse(
    is.na(data$quarter),
    NA_character_,
    paste0(
      "2025_Q",
      data$quarter
    )
  )


  data$period_index <- ifelse(
    is.na(data$quarter),
    NA_integer_,
    data$year * 4L + data$quarter
  )


  data$plfs_quarter_code <- quarter_number


  data
}
```

Applying quarter-year functions.
```{r}

# 2017-18: Q1-Q4, July 2017-June 2018
analysis17_18 <- add_annual_quarter(
  data = analysis17_18,
  quarter_column = "quarter_per_fv",
  start_year = 2017
)


# 2018-19: Q5-Q8, July 2018-June 2019
analysis18_19 <- add_annual_quarter(
  data = analysis18_19,
  quarter_column = "quarter_per_fv",
  start_year = 2018
)


# 2019-20: Q1-Q4, July 2019-June 2020
analysis19_20 <- add_annual_quarter(
  data = analysis19_20,
  quarter_column = "quarter_per_fv",
  start_year = 2019
)


# 2020-21: Q5-Q8, July 2020-June 2021
analysis20_21 <- add_annual_quarter(
  data = analysis20_21,
  quarter_column = "qtr_perv1",
  start_year = 2020
)


# Calendar year 2022
analysis22 <- add_calendar_year_quarter(
  data = analysis22,
  quarter_column = "qtr_cperv1",
  calendar_year = 2022
)


# Calendar year 2023
analysis23 <- add_calendar_year_quarter(
  data = analysis23,
  quarter_column = "qtr_cperv1",
  calendar_year = 2023
)


# Calendar year 2024
analysis24 <- add_calendar_year_quarter(
  data = analysis24,
  quarter_column = "Quarter",
  calendar_year = 2024
)


# Calendar year 2025
analysis25 <- add_2025_quarter(
  data = analysis25,
  quarter_column = "qtr"
)
```


Combining all the above into one CSV file so that I don't have to load the whole thing every single time:
```{r}

```



Metro binary and urban binary
```{r}
# ============================================================
# Nationwide district-quarter metro and urban indicators
# ============================================================

library(dplyr)
library(tidyr)
library(purrr)
library(readxl)
library(stringr)
library(tibble)
library(here)



# ============================================================
# 1. Add the PLFS urban/rural indicator
#
# PLFS sector code:
# 1 = rural
# 2 = urban
# ============================================================

add_urban_flag <- function(
    data,
    sector_column
) {

  if (!(sector_column %in% names(data))) {
    stop(
      paste(
        "Sector column not found:",
        sector_column
      )
    )
  }

  sector_code <- clean_code(
    data[[sector_column]]
  )

  data$sector_code <- as.integer(
    sector_code
  )

  data$urban <- case_when(
    sector_code == 2 ~ 1L,
    sector_code == 1 ~ 0L,
    TRUE ~ NA_integer_
  )

  data$rural <- case_when(
    sector_code == 1 ~ 1L,
    sector_code == 2 ~ 0L,
    TRUE ~ NA_integer_
  )

  data
}


analysis17_18 <- add_urban_flag(
  analysis17_18,
  sector_column = "b1q3_per_fv"
)

analysis18_19 <- add_urban_flag(
  analysis18_19,
  sector_column = "b1q3_per_fv"
)

analysis19_20 <- add_urban_flag(
  analysis19_20,
  sector_column = "b1q3_per_fv"
)

analysis20_21 <- add_urban_flag(
  analysis20_21,
  sector_column = "b1q3_perv1"
)

analysis22 <- add_urban_flag(
  analysis22,
  sector_column = "b1q3_cperv1"
)

analysis23 <- add_urban_flag(
  analysis23,
  sector_column = "b1q3_cperv1"
)

analysis24 <- add_urban_flag(
  analysis24,
  sector_column = "Sector"
)

analysis25 <- add_urban_flag(
  analysis25,
  sector_column = "sec"
)


# ============================================================
# 2. Download the official MoSPI district-code workbooks
#
# This section needs an internet connection on the first run.
# The files are then retained locally.
# ============================================================

district_crosswalk_folder <- here(
  "Data",
  "Crosswalks",
  "PLFS_District_Codes"
)

dir.create(
  district_crosswalk_folder,
  recursive = TRUE,
  showWarnings = FALSE
)


district_file_information <- tribble(
  ~panel,   ~file_name,                  ~download_url,

  "panel1", "plfs_district_panel1.xlsx",
  "https://microdata.gov.in/NADA/index.php/catalog/204/download/2662",

  "panel2", "plfs_district_panel2.xlsx",
  "https://microdata.gov.in/NADA/index.php/catalog/217/download/2665",

  "panel3", "plfs_district_panel3.xlsx",
  "https://microdata.gov.in/NADA/index.php/catalog/210/download/2667",

  "panel4", "plfs_district_panel4.xlsx",
  "https://microdata.gov.in/NADA/index.php/catalog/213/download/2668",

  "y2025",  "plfs_district_2025.xlsx",
  "https://microdata.gov.in/NADA/index.php/catalog/284/download/5920"
) %>%
  mutate(
    local_path = file.path(
      district_crosswalk_folder,
      file_name
    )
  )


download_government_file <- function(
    download_url,
    local_path
) {

  needs_download <-
    !file.exists(local_path) ||
    is.na(file.info(local_path)$size) ||
    file.info(local_path)$size < 5000

  if (needs_download) {

    message(
      "Downloading official MoSPI file: ",
      basename(local_path)
    )

    tryCatch(
      {
        download.file(
          url = download_url,
          destfile = local_path,
          mode = "wb",
          quiet = FALSE
        )
      },
      error = function(e) {
        stop(
          paste0(
            "Could not download:\n",
            download_url,
            "\n\nDownload it manually and save it as:\n",
            local_path,
            "\n\nOriginal error:\n",
            e$message
          )
        )
      }
    )
  }

  local_path
}


district_file_information$local_path <- map2_chr(
  district_file_information$download_url,
  district_file_information$local_path,
  download_government_file
)


# ============================================================
# 3. Functions for reading the official district workbooks
# ============================================================

clean_workbook_names <- function(x) {

  x <- as.character(x)

  x[is.na(x)] <- ""

  x <- tolower(
    trimws(x)
  )

  x <- str_replace_all(
    x,
    "[^a-z0-9]+",
    "_"
  )

  x <- str_replace_all(
    x,
    "^_+|_+$",
    ""
  )

  blank_names <- which(x == "")

  if (length(blank_names) > 0) {
    x[blank_names] <- paste0(
      "unnamed_",
      seq_along(blank_names)
    )
  }

  make.unique(
    x,
    sep = "_"
  )
}


# ============================================================
# Identify columns inside the MoSPI district-code workbooks
# ============================================================

find_workbook_column <- function(
    column_names,
    patterns,
    description
) {

  for (pattern in patterns) {

    matches <- grep(
      pattern,
      column_names,
      value = TRUE,
      ignore.case = TRUE
    )

    if (length(matches) > 0) {
      return(matches[1])
    }
  }

  stop(
    paste0(
      "Could not identify the ",
      description,
      " column.\nAvailable columns: ",
      paste(column_names, collapse = ", ")
    )
  )
}


# ============================================================
# Combine one or two workbook rows into possible headers
# ============================================================

make_header_candidate <- function(
    raw_sheet,
    start_row,
    number_of_rows = 1
) {

  end_row <- min(
    start_row + number_of_rows - 1,
    nrow(raw_sheet)
  )

  header_rows <- raw_sheet[
    start_row:end_row,
    ,
    drop = FALSE
  ]

  combined_headers <- vapply(
    seq_len(ncol(header_rows)),
    function(column_number) {

      values <- as.character(
        header_rows[[column_number]]
      )

      values <- values[
        !is.na(values) &
          trimws(values) != ""
      ]

      paste(
        values,
        collapse = " "
      )
    },
    character(1)
  )

  combined_headers
}


# ============================================================
# Score whether a row looks like the real table header
# ============================================================

score_header_candidate <- function(header_values) {

  header_text <- tolower(
    paste(
      header_values,
      collapse = " | "
    )
  )

  has_state_code <- grepl(
    "state.*code|state.?ut.*code|st.*code",
    header_text
  )

  has_district_code <- grepl(
    "district.*code|dist.*code",
    header_text
  )

  has_district_name <- grepl(
    "district.*name|name.*district|name of district",
    header_text
  )

  score <-
    5L * as.integer(has_state_code) +
    5L * as.integer(has_district_code) +
    2L * as.integer(has_district_name)

  list(
    score = score,
    has_state_code = has_state_code,
    has_district_code = has_district_code
  )
}


# ============================================================
# Find the actual header row
# ============================================================

find_district_header <- function(
    raw_sheet,
    maximum_rows = 40
) {

  rows_to_check <- seq_len(
    min(
      maximum_rows,
      nrow(raw_sheet)
    )
  )

  candidates <- list()
  candidate_number <- 1L


  for (header_depth in c(1L, 2L)) {

    for (header_row in rows_to_check) {

      if (
        header_row + header_depth - 1L >
          nrow(raw_sheet)
      ) {
        next
      }

      header_values <- make_header_candidate(
        raw_sheet = raw_sheet,
        start_row = header_row,
        number_of_rows = header_depth
      )

      header_score <- score_header_candidate(
        header_values
      )

      candidates[[candidate_number]] <- list(
        header_row = header_row,
        header_depth = header_depth,
        header_values = header_values,
        score = header_score$score,
        valid =
          header_score$has_state_code &&
          header_score$has_district_code
      )

      candidate_number <- candidate_number + 1L
    }
  }


  valid_candidates <- candidates[
    vapply(
      candidates,
      function(x) x$valid,
      logical(1)
    )
  ]


  if (length(valid_candidates) == 0) {

    preview <- raw_sheet[
      seq_len(min(10, nrow(raw_sheet))),
      ,
      drop = FALSE
    ]

    print(preview)

    stop(
      paste(
        "Could not find the actual district-table header.",
        "The first ten workbook rows were printed above."
      )
    )
  }


  candidate_scores <- vapply(
    valid_candidates,
    function(x) x$score,
    integer(1)
  )

  valid_candidates[[
    which.max(candidate_scores)
  ]]
}


extract_first_integer <- function(x) {

  suppressWarnings(
    as.integer(
      str_extract(
        as.character(x),
        "[0-9]+"
      )
    )
  )
}


# ============================================================
# Read one sheet from a MoSPI district-code workbook
# ============================================================

read_district_sheet <- function(
    file_path,
    sheet_name,
    panel_name
) {

  raw_sheet <- suppressMessages(
    readxl::read_excel(
      path = file_path,
      sheet = sheet_name,
      col_names = FALSE,
      col_types = "text"
    )
  )

  raw_sheet <- as.data.frame(
    raw_sheet,
    stringsAsFactors = FALSE
  )


  if (
    nrow(raw_sheet) == 0 ||
    ncol(raw_sheet) == 0
  ) {
    return(
      tibble::tibble()
    )
  }


  # Find the genuine table header rather than the title
  header_information <- find_district_header(
    raw_sheet
  )

  header_row <- header_information$header_row
  header_depth <- header_information$header_depth

  header_values <- header_information$header_values

  first_data_row <-
    header_row +
    header_depth


  if (first_data_row > nrow(raw_sheet)) {
    return(
      tibble::tibble()
    )
  }


  body <- raw_sheet[
    first_data_row:nrow(raw_sheet),
    ,
    drop = FALSE
  ]


  names(body) <- clean_workbook_names(
    header_values
  )


  # Remove completely empty columns
  nonempty_columns <- vapply(
    body,
    function(x) {
      any(
        !is.na(x) &
          trimws(as.character(x)) != ""
      )
    },
    logical(1)
  )

  body <- body[
    ,
    nonempty_columns,
    drop = FALSE
  ]


  # Identify the three required columns
  state_code_column <- find_workbook_column(
    names(body),
    patterns = c(
      "state_ut.*code",
      "state.*code",
      "^st.*code",
      "^state_ut$"
    ),
    description = "state code"
  )


  district_code_column <- find_workbook_column(
    names(body),
    patterns = c(
      "district.*code",
      "^dist.*code"
    ),
    description = "district code"
  )


  district_name_column <- find_workbook_column(
    names(body),
    patterns = c(
      "district.*name",
      "name.*district",
      "name_of_district",
      "^district_name$",
      "^district$"
    ),
    description = "district name"
  )


  # Replace blank state entries with NA before filling downward
  body[[state_code_column]] <- as.character(
    body[[state_code_column]]
  )

  body[[state_code_column]][
    trimws(body[[state_code_column]]) == ""
  ] <- NA_character_


  # Workbooks often display the state code only on the first
  # district row for each state
  body <- tidyr::fill(
    body,
    dplyr::all_of(state_code_column),
    .direction = "down"
  )


  result <- tibble::tibble(
    panel = panel_name,

    state_code = extract_first_integer(
      body[[state_code_column]]
    ),

    district_code = extract_first_integer(
      body[[district_code_column]]
    ),

    district_name = stringr::str_squish(
      as.character(
        body[[district_name_column]]
      )
    )
  ) %>%
    dplyr::filter(
      !is.na(state_code),
      !is.na(district_code),
      !is.na(district_name),
      district_name != "",
      !stringr::str_detect(
        tolower(district_name),
        "^total"
      )
    )


  result
}


read_district_workbook <- function(
    file_path,
    panel_name
) {

  sheet_names <- excel_sheets(
    file_path
  )

  result <- map_dfr(
    sheet_names,
    ~ read_district_sheet(
      file_path = file_path,
      sheet_name = .x,
      panel_name = panel_name
    )
  )

  if (nrow(result) == 0) {
    stop(
      paste(
        "No district codes could be read from:",
        file_path
      )
    )
  }

  result
}


official_district_lookup <- map2_dfr(
  district_file_information$local_path,
  district_file_information$panel,
  read_district_workbook
) %>%
  distinct(
    panel,
    state_code,
    district_code,
    .keep_all = TRUE
  )


# ============================================================
# 4. Standardise district names
# ============================================================

normalise_district_name <- function(x) {

  x <- iconv(
    as.character(x),
    from = "",
    to = "ASCII//TRANSLIT"
  )

  x <- tolower(x)

  x <- str_replace_all(
    x,
    "&",
    " and "
  )

  x <- str_replace_all(
    x,
    "[^a-z0-9]+",
    " "
  )

  x <- str_squish(x)


  case_when(
    x %in% c(
      "gurgaon",
      "gurugram"
    ) ~ "gurugram",

    x %in% c(
      "gautam budh nagar",
      "gautam buddh nagar",
      "gautam buddha nagar"
    ) ~ "gautam buddha nagar",

    x %in% c(
      "bangalore urban",
      "bengaluru urban",
      "bangalore"
    ) ~ "bengaluru urban",

    x %in% c(
      "ranga reddy",
      "rangareddy"
    ) ~ "rangareddy",

    x %in% c(
      "medchal malkajgiri",
      "medchal malkajgiri district"
    ) ~ "medchal malkajgiri",

    x %in% c(
      "north twenty four parganas",
      "north 24 parganas",
      "24 parganas north",
      "north 24 paraganas"
    ) ~ "north 24 parganas",

    x %in% c(
      "south twenty four parganas",
      "south 24 parganas",
      "24 parganas south",
      "south 24 paraganas"
    ) ~ "south 24 parganas",

    x %in% c(
      "calcutta",
      "kolkata"
    ) ~ "kolkata",

    x %in% c(
      "kancheepuram",
      "kanchipuram",
      "kanchipuram district"
    ) ~ "kanchipuram",

    x %in% c(
      "ahmadabad",
      "ahmedabad"
    ) ~ "ahmedabad",

    x %in% c(
      "bombay suburban",
      "mumbai suburban"
    ) ~ "mumbai suburban",

    x %in% c(
      "bombay city",
      "mumbai city"
    ) ~ "mumbai city",

    x %in% c(
      "kanpur",
      "kanpur nagar"
    ) ~ "kanpur nagar",

    x %in% c(
      "ahmadnagar",
      "ahmednagar",
      "ahilyanagar"
    ) ~ "ahilyanagar",

    TRUE ~ x
  )
}


official_district_lookup <- official_district_lookup %>%
  mutate(
    district_name_key = normalise_district_name(
      district_name
    )
  )


# ============================================================
# 5. Attach official district names to one PLFS dataset
# ============================================================

attach_official_district_name <- function(
    data,
    panel
) {

  if (length(panel) == 1) {
    panel <- rep(
      panel,
      nrow(data)
    )
  }

  if (length(panel) != nrow(data)) {
    stop(
      "The panel vector must have one value per observation."
    )
  }

  data$.original_row_order <- seq_len(
    nrow(data)
  )

  data$.district_panel <- as.character(
    panel
  )

  data$.state_code_join <- suppressWarnings(
    as.integer(
      data$state_code_std
    )
  )

  data$.district_code_join <- suppressWarnings(
    as.integer(
      data$district_code_std
    )
  )


  data <- data %>%
    select(
      -any_of(
        c(
          "district_name",
          "district_name_key"
        )
      )
    ) %>%
    left_join(
      official_district_lookup %>%
        select(
          panel,
          state_code,
          district_code,
          district_name,
          district_name_key
        ),

      by = c(
        ".district_panel" = "panel",
        ".state_code_join" = "state_code",
        ".district_code_join" = "district_code"
      )
    ) %>%
    arrange(
      .original_row_order
    ) %>%
    select(
      -.original_row_order,
      -.state_code_join,
      -.district_code_join,
      -.district_panel
    )

  data
}


# Panel assignments
analysis17_18 <- attach_official_district_name(
  analysis17_18,
  panel = "panel1"
)

analysis18_19 <- attach_official_district_name(
  analysis18_19,
  panel = "panel1"
)

analysis19_20 <- attach_official_district_name(
  analysis19_20,
  panel = "panel2"
)

analysis20_21 <- attach_official_district_name(
  analysis20_21,
  panel = "panel2"
)

analysis22 <- attach_official_district_name(
  analysis22,
  panel = "panel3"
)


# Calendar 2023 combines the second half of Panel 3
# and the first half of Panel 4.
analysis23_panel <- ifelse(
  analysis23$quarter %in% c(1L, 2L),
  "panel3",
  "panel4"
)

analysis23 <- attach_official_district_name(
  analysis23,
  panel = analysis23_panel
)


analysis24 <- attach_official_district_name(
  analysis24,
  panel = "panel4"
)

analysis25 <- attach_official_district_name(
  analysis25,
  panel = "y2025"
)


# ============================================================
# 6. Official source registry
#
# Source URLs are retained in the resulting datasets so each
# positive metro classification can be audited.
# ============================================================

metro_sources <- tribble(
  ~source_id, ~metro_source_url,

  "DMRC",
  "https://www.delhimetrorail.com/pages/en/introduction",

  "JMRC",
  "https://transport.rajasthan.gov.in/content/transportportal/en/metro.html",

  "BMRCL",
  "https://english.bmrc.co.in/",

  "CMRL",
  "https://chennaimetrorail.org/",

  "KMRL",
  "https://kochimetro.org/",

  "KOLKATA_METRO",
  "https://mtp.indianrailways.gov.in/",

  "MMRDA",
  "https://mmrda.maharashtra.gov.in/en/projects/transport/metro-line-1/overview",

  "HMRL",
  "https://hmrl.co.in/",

  "PIB_LUCKNOW",
  "https://www.pib.gov.in/newsite/PrintRelease.aspx?lang=2&reg=48&relid=170511",

  "PIB_BAHADURGARH",
  "https://www.pib.gov.in/Pressreleaseshare.aspx?PRID=1536394",

  "PIB_AHMEDABAD",
  "https://www.pib.gov.in/PressReleasePage.aspx?PRID=1567375",

  "PIB_NAGPUR",
  "https://www.pib.gov.in/PressReleasePage.aspx?PRID=1568013",

  "PIB_KANPUR",
  "https://www.pib.gov.in/PressReleasePage.aspx?PRID=1844647",

  "PIB_PUNE",
  "https://www.pib.gov.in/PressReleasePage.aspx?PRID=1803340",

  "CIDCO",
  "https://cidco.maharashtra.gov.in/",

  "PIB_AGRA_HOWRAH",
  "https://www.pib.gov.in/PressReleasePage.aspx?PRID=2011999",

  "PIB_GANDHINAGAR",
  "https://www.pib.gov.in/PressReleasePage.aspx?PRID=2055368",

  "MMRCL",
  "https://mmrcl.com/",

  "PIB_INDORE",
  "https://www.pib.gov.in/PressReleasePage.aspx?PRID=2132584",

  "PIB_BHOPAL",
  "https://www.pib.gov.in/PressReleaseIframePage.aspx?PRID=2207030&lang=2&reg=48"
)


# ============================================================
# 7. District-level metro-opening crosswalk
#
# pre_sample = TRUE means that the district already had an
# operational metro before the first PLFS quarter, 2017 Q3.
#
# manual_review = 1 flags cases affected by changing district
# boundaries or metro lines crossing district boundaries.
# ============================================================

metro_district_crosswalk <- tribble(
  ~state_code, ~district_name,          ~metro_system,
  ~opening_year, ~opening_quarter, ~pre_sample,
  ~source_id, ~manual_review, ~metro_note,

  6L, "Gurugram", "Delhi Metro / Rapid Metro",
  2010L, 2L, TRUE,
  "DMRC", 0L,
  "Operational before the PLFS sample.",

  6L, "Faridabad", "Delhi Metro",
  2015L, 3L, TRUE,
  "DMRC", 0L,
  "Operational before the PLFS sample.",

  6L, "Jhajjar", "Delhi Metro: Bahadurgarh extension",
  2018L, 2L, FALSE,
  "PIB_BAHADURGARH", 0L,
  "Bahadurgarh is within Jhajjar district.",


  9L, "Ghaziabad", "Delhi Metro",
  2011L, 3L, TRUE,
  "DMRC", 0L,
  "Vaishali metro was operational before the PLFS sample.",

  9L, "Gautam Buddha Nagar", "Delhi Metro / Noida Metro",
  2009L, 4L, TRUE,
  "DMRC", 0L,
  "Noida had metro service before the PLFS sample.",

  9L, "Lucknow", "Lucknow Metro",
  2017L, 3L, FALSE,
  "PIB_LUCKNOW", 0L,
  "Priority corridor inaugurated in September 2017.",

  9L, "Kanpur Nagar", "Kanpur Metro",
  2021L, 4L, FALSE,
  "PIB_KANPUR", 0L,
  "Commercial operation began in December 2021.",

  9L, "Agra", "Agra Metro",
  2024L, 1L, FALSE,
  "PIB_AGRA_HOWRAH", 0L,
  "First operational stretch inaugurated in March 2024.",


  8L, "Jaipur", "Jaipur Metro",
  2015L, 2L, TRUE,
  "JMRC", 0L,
  "Operational before the PLFS sample.",


  24L, "Ahmedabad", "Ahmedabad Metro",
  2019L, 1L, FALSE,
  "PIB_AHMEDABAD", 0L,
  "First operational section inaugurated in March 2019.",

  24L, "Gandhinagar", "Ahmedabad-Gandhinagar Metro",
  2024L, 3L, FALSE,
  "PIB_GANDHINAGAR", 0L,
  "Phase II inaugurated in September 2024.",


  27L, "Mumbai Suburban", "Mumbai Metro",
  2014L, 2L, TRUE,
  "MMRDA", 0L,
  "Mumbai Metro Line 1 was operational before the sample.",

  27L, "Mumbai City", "Mumbai Metro Line 3",
  2025L, 2L, FALSE,
  "MMRCL", 1L,
  paste(
    "Southward extension reached Mumbai City district",
    "during 2025 Q2; review against final station boundaries."
  ),

  27L, "Thane", "Navi Mumbai Metro",
  2023L, 4L, FALSE,
  "CIDCO", 1L,
  paste(
    "Navi Mumbai Metro crosses the Thane-Raigad boundary;",
    "retain for boundary audit."
  ),

  27L, "Raigad", "Navi Mumbai Metro",
  2023L, 4L, FALSE,
  "CIDCO", 1L,
  paste(
    "Navi Mumbai Metro crosses the Thane-Raigad boundary;",
    "retain for boundary audit."
  ),

  27L, "Nagpur", "Nagpur Metro",
  2019L, 1L, FALSE,
  "PIB_NAGPUR", 0L,
  "First operational section inaugurated in March 2019.",

  27L, "Pune", "Pune Metro",
  2022L, 1L, FALSE,
  "PIB_PUNE", 0L,
  "First operational sections inaugurated in March 2022.",


  29L, "Bengaluru Urban", "Namma Metro",
  2011L, 4L, TRUE,
  "BMRCL", 0L,
  "Operational before the PLFS sample.",


  36L, "Hyderabad", "Hyderabad Metro",
  2017L, 4L, FALSE,
  "HMRL", 0L,
  "First operational corridors opened in November 2017.",

  36L, "Rangareddy", "Hyderabad Metro",
  2017L, 4L, FALSE,
  "HMRL", 1L,
  "First corridor crosses Hyderabad metropolitan district boundaries.",

  36L, "Medchal Malkajgiri", "Hyderabad Metro",
  2017L, 4L, FALSE,
  "HMRL", 1L,
  "First corridor crosses Hyderabad metropolitan district boundaries.",


  33L, "Chennai", "Chennai Metro",
  2015L, 2L, TRUE,
  "CMRL", 0L,
  "Operational before the PLFS sample.",

  33L, "Kanchipuram", "Chennai Metro",
  2015L, 2L, TRUE,
  "CMRL", 1L,
  paste(
    "Review because Chennai district boundaries expanded",
    "during the study period."
  ),


  32L, "Ernakulam", "Kochi Metro",
  2017L, 2L, TRUE,
  "KMRL", 0L,
  "Kochi Metro opened immediately before the first PLFS quarter.",


  19L, "Kolkata", "Kolkata Metro",
  1984L, 4L, TRUE,
  "KOLKATA_METRO", 0L,
  "Operational before the PLFS sample.",

  19L, "North 24 Parganas", "Kolkata Metro",
  1995L, 3L, TRUE,
  "KOLKATA_METRO", 1L,
  "Dum Dum and later extensions serve this district.",

  19L, "Howrah", "Kolkata Metro: East-West line",
  2024L, 1L, FALSE,
  "PIB_AGRA_HOWRAH", 0L,
  "Howrah Maidan-Esplanade section opened in March 2024.",


  23L, "Indore", "Indore Metro",
  2025L, 2L, FALSE,
  "PIB_INDORE", 0L,
  "Passenger services inaugurated in May 2025.",

  23L, "Bhopal", "Bhopal Metro",
  2025L, 4L, FALSE,
  "PIB_BHOPAL", 0L,
  "Priority corridor opened in December 2025."
) %>%
  mutate(
    district_name_key = normalise_district_name(
      district_name
    ),

    metro_pre_sample = as.integer(
      pre_sample
    ),

    metro_start_period_index = case_when(
      pre_sample ~ 0L,

      TRUE ~
        opening_year * 4L +
        opening_quarter
    ),

    metro_manual_review = as.integer(
      manual_review
    )
  ) %>%
  left_join(
    metro_sources,
    by = "source_id"
  ) %>%
  transmute(
    state_code,
    district_name_key,

    metro_system,

    metro_opening_year =
      as.integer(opening_year),

    metro_opening_quarter =
      as.integer(opening_quarter),

    metro_start_period_index =
      as.integer(metro_start_period_index),

    metro_pre_sample,

    metro_source_id = source_id,

    metro_source_url,

    metro_manual_review,

    metro_note
  )


# ============================================================
# 8. Delhi state-level rule
#
# Every PLFS district in Delhi is treated as having metro
# infrastructure throughout the study period.
#
# The historical opening quarter differs by district, but every
# Delhi district was already served before the PLFS sample.
# ============================================================

metro_state_crosswalk <- tibble(
  state_code = 7L,

  metro_system =
    "Delhi Metro",

  metro_opening_year =
    NA_integer_,

  metro_opening_quarter =
    NA_integer_,

  metro_start_period_index =
    0L,

  metro_pre_sample =
    1L,

  metro_source_id =
    "DMRC",

  metro_manual_review =
    1L,

  metro_note = paste(
    "All Delhi districts are coded as pre-sample metro districts.",
    "District-specific historical opening dates vary."
  )
) %>%
  left_join(
    metro_sources,
    by = c(
      "metro_source_id" = "source_id"
    )
  )


# ============================================================
# 9. Add metro indicators to one dataset
# ============================================================

add_metro_flag <- function(data) {

  required_columns <- c(
    "state_code_std",
    "district_name_key",
    "year",
    "quarter",
    "period_index"
  )

  missing_columns <- setdiff(
    required_columns,
    names(data)
  )

  if (length(missing_columns) > 0) {
    stop(
      paste(
        "Missing columns:",
        paste(missing_columns, collapse = ", ")
      )
    )
  }


  # Make the function safe to rerun
  data <- data %>%
    select(
      -any_of(
        c(
          "has_metro",
          "metro_opened_this_quarter",
          "metro_urban",
          "metro_system",
          "metro_opening_year",
          "metro_opening_quarter",
          "metro_start_period_index",
          "metro_pre_sample",
          "metro_source_id",
          "metro_source_url",
          "metro_manual_review",
          "metro_note"
        )
      )
    ) %>%
    mutate(
      .state_code_num = suppressWarnings(
        as.integer(state_code_std)
      )
    )


  # Exact state-district matches
  data <- data %>%
    left_join(
      metro_district_crosswalk,

      by = c(
        ".state_code_num" = "state_code",
        "district_name_key" = "district_name_key"
      )
    )


  # State-wide rules, currently used for Delhi
  state_lookup <- metro_state_crosswalk %>%
    rename(
      metro_system_state =
        metro_system,

      metro_opening_year_state =
        metro_opening_year,

      metro_opening_quarter_state =
        metro_opening_quarter,

      metro_start_period_index_state =
        metro_start_period_index,

      metro_pre_sample_state =
        metro_pre_sample,

      metro_source_id_state =
        metro_source_id,

      metro_source_url_state =
        metro_source_url,

      metro_manual_review_state =
        metro_manual_review,

      metro_note_state =
        metro_note
    )


  data <- data %>%
    left_join(
      state_lookup,
      by = c(
        ".state_code_num" = "state_code"
      )
    ) %>%
    mutate(
      metro_system = coalesce(
        metro_system,
        metro_system_state
      ),

      metro_opening_year = coalesce(
        metro_opening_year,
        metro_opening_year_state
      ),

      metro_opening_quarter = coalesce(
        metro_opening_quarter,
        metro_opening_quarter_state
      ),

      metro_start_period_index = coalesce(
        metro_start_period_index,
        metro_start_period_index_state
      ),

      metro_pre_sample = coalesce(
        metro_pre_sample,
        metro_pre_sample_state
      ),

      metro_source_id = coalesce(
        metro_source_id,
        metro_source_id_state
      ),

      metro_source_url = coalesce(
        metro_source_url,
        metro_source_url_state
      ),

      metro_manual_review = coalesce(
        metro_manual_review,
        metro_manual_review_state
      ),

      metro_note = coalesce(
        metro_note,
        metro_note_state
      ),


      # Main district-quarter treatment indicator
      has_metro = as.integer(
        !is.na(metro_start_period_index) &
          period_index >=
          metro_start_period_index
      ),


      # Identifies the opening quarter itself
      metro_opened_this_quarter = as.integer(
        !is.na(metro_start_period_index) &
          metro_start_period_index > 0L &
          period_index ==
          metro_start_period_index
      ),


      # Recommended exposure indicator:
      # metro district AND urban PLFS respondent
      metro_urban = case_when(
        is.na(urban) ~ NA_integer_,

        has_metro == 1L &
          urban == 1L ~ 1L,

        TRUE ~ 0L
      ),


      # Zero for districts with no verified metro match
      metro_pre_sample = coalesce(
        metro_pre_sample,
        0L
      ),

      metro_manual_review = coalesce(
        metro_manual_review,
        0L
      )
    ) %>%
    select(
      -.state_code_num,

      -metro_system_state,
      -metro_opening_year_state,
      -metro_opening_quarter_state,
      -metro_start_period_index_state,
      -metro_pre_sample_state,
      -metro_source_id_state,
      -metro_source_url_state,
      -metro_manual_review_state,
      -metro_note_state
    )


  data
}


# ============================================================
# 10. Apply the metro function to every analysis dataset
# ============================================================

analysis17_18 <- add_metro_flag(
  analysis17_18
)

analysis18_19 <- add_metro_flag(
  analysis18_19
)

analysis19_20 <- add_metro_flag(
  analysis19_20
)

analysis20_21 <- add_metro_flag(
  analysis20_21
)

analysis22 <- add_metro_flag(
  analysis22
)

analysis23 <- add_metro_flag(
  analysis23
)

analysis24 <- add_metro_flag(
  analysis24
)

analysis25 <- add_metro_flag(
  analysis25
)


# ============================================================
# 11. Validate the resulting columns
# ============================================================

validate_metro_flags <- function(
    data,
    dataset_name
) {

  valid_urban_values <- all(
    is.na(data$urban) |
      data$urban %in% c(0L, 1L)
  )

  valid_metro_values <- all(
    data$has_metro %in% c(0L, 1L)
  )

  valid_opening_values <- all(
    data$metro_opened_this_quarter %in%
      c(0L, 1L)
  )

  valid_metro_urban <- all(
    is.na(data$metro_urban) |
      data$metro_urban ==
      as.integer(
        data$has_metro == 1L &
          data$urban == 1L
      )
  )

  if (!valid_urban_values) {
    stop(
      dataset_name,
      ": urban contains invalid values."
    )
  }

  if (!valid_metro_values) {
    stop(
      dataset_name,
      ": has_metro contains invalid values."
    )
  }

  if (!valid_opening_values) {
    stop(
      dataset_name,
      ": metro_opened_this_quarter contains invalid values."
    )
  }

  if (!valid_metro_urban) {
    stop(
      dataset_name,
      ": metro_urban was not constructed correctly."
    )
  }

  message(
    dataset_name,
    ": metro and urban checks passed."
  )
}


validate_metro_flags(
  analysis17_18,
  "2017-18"
)

validate_metro_flags(
  analysis18_19,
  "2018-19"
)

validate_metro_flags(
  analysis19_20,
  "2019-20"
)

validate_metro_flags(
  analysis20_21,
  "2020-21"
)

validate_metro_flags(
  analysis22,
  "2022"
)

validate_metro_flags(
  analysis23,
  "2023"
)

validate_metro_flags(
  analysis24,
  "2024"
)

validate_metro_flags(
  analysis25,
  "2025"
)


# ============================================================
# 12. Combine a small district-quarter validation table
# ============================================================

# ============================================================
# Create a standardised district-quarter validation table
# without combining all original PLFS variables
# ============================================================

prepare_metro_validation <- function(
    data,
    dataset_name
) {

  data %>%
    transmute(
      source_dataset = dataset_name,

      state_code_std = as.character(
        state_code_std
      ),

      district_name = as.character(
        district_name
      ),

      district_name_key = as.character(
        district_name_key
      ),

      year = as.integer(
        year
      ),

      quarter = as.integer(
        quarter
      ),

      year_quarter = as.character(
        year_quarter
      ),

      period_index = as.integer(
        period_index
      ),

      urban = as.integer(
        urban
      ),

      rural = as.integer(
        rural
      ),

      has_metro = as.integer(
        has_metro
      ),

      metro_opened_this_quarter = as.integer(
        metro_opened_this_quarter
      ),

      metro_urban = as.integer(
        metro_urban
      ),

      metro_system = as.character(
        metro_system
      ),

      metro_opening_year = as.integer(
        metro_opening_year
      ),

      metro_opening_quarter = as.integer(
        metro_opening_quarter
      ),

      metro_manual_review = as.integer(
        metro_manual_review
      ),

      metro_source_url = as.character(
        metro_source_url
      )
    )
}


metro_validation_data <- bind_rows(
  prepare_metro_validation(
    analysis17_18,
    "2017-18"
  ),

  prepare_metro_validation(
    analysis18_19,
    "2018-19"
  ),

  prepare_metro_validation(
    analysis19_20,
    "2019-20"
  ),

  prepare_metro_validation(
    analysis20_21,
    "2020-21"
  ),

  prepare_metro_validation(
    analysis22,
    "2022"
  ),

  prepare_metro_validation(
    analysis23,
    "2023"
  ),

  prepare_metro_validation(
    analysis24,
    "2024"
  ),

  prepare_metro_validation(
    analysis25,
    "2025"
  )
) %>%
  distinct() %>%
  arrange(
    period_index,
    state_code_std,
    district_name
  )


# Opening events occurring inside the PLFS period
metro_opening_events <- metro_validation_data %>%
  filter(
    metro_opened_this_quarter == 1L
  ) %>%
  distinct(
    district_name,
    year_quarter,
    metro_system,
    metro_source_url,
    metro_manual_review
  ) %>%
  arrange(
    year_quarter,
    district_name
  )

print(
  metro_opening_events
)


# All districts ever coded as metro districts
positive_metro_districts <- metro_validation_data %>%
  filter(
    has_metro == 1L
  ) %>%
  distinct(
    state_code_std,
    district_name,
    metro_system,
    metro_opening_year,
    metro_opening_quarter,
    metro_manual_review
  ) %>%
  arrange(
    state_code_std,
    district_name
  )

print(
  positive_metro_districts
)


# Cases that should be reviewed before the final dissertation
metro_boundary_review <- positive_metro_districts %>%
  filter(
    metro_manual_review == 1L
  )

print(
  metro_boundary_review
)


# ============================================================
# 13. Check for impossible treatment reversals
#
# Once a district receives has_metro = 1, it should never
# return to zero in a later quarter.
# ============================================================

metro_reversal_check <- metro_validation_data %>%
  group_by(
    state_code_std,
    district_name_key
  ) %>%
  arrange(
    period_index,
    .by_group = TRUE
  ) %>%
  summarise(
    treatment_reversal = any(
      diff(has_metro) < 0,
      na.rm = TRUE
    ),

    .groups = "drop"
  ) %>%
  filter(
    treatment_reversal
  )

if (nrow(metro_reversal_check) > 0) {

  print(
    metro_reversal_check
  )

  warning(
    "At least one district changes from has_metro = 1 back to 0."
  )

} else {

  message(
    "No metro treatment reversals detected."
  )
}
```




