---
title: "Cleaning the Dataset"
---
Import the data.
```{r}
#for reading dta
library(haven)
#for easier file navigation
library(here)

data17_18_fv <- read_dta(here("Data", "17-18", "hh_per_fv_2017-18.dta"))
#data17_18_rv <- read_dta(here("Data", "17-18", "hh_per_rv_2017-18.dta"))

data18_19_fv <- read_dta(here("Data","18-19","PerV1_2018-19.dta"))
#data18_19_rv <- read_dta(here("Data","18-19","PerRV_2018-19.dta"))

data19_20_fv <- read_dta(here("Data","19-20","PERFV_2019-20.dta"))
#data19_20_rv <- read_dta(here("Data","19-20","PERRV_2019-20.dta"))

data20_21_fv <- read_dta(here("Data","20-21","perv1.dta"))
#data20_21_rv <- read_dta(here("Data","20-21","perrv.dta"))

data22 <- read_dta(here("Data","22","cperv1.dta"))

data23 <- read_dta(here("Data","23","cperv1.dta"))

data24 <- read.csv(here("Data","24","cperv1.csv"))

data25 <- read_dta(here("Data","25","cperv12025.dta"))


```

Function for cleaning the dataset, so that the stata labels don't interfere with the data cleaning process.

```{r}
clean_code <- function(x) {
  
  # Remove the special Stata-labelled class
  if (inherits(x, "haven_labelled") ||
      inherits(x, "labelled")) {
    x <- haven::zap_labels(x)
  }
  
  # Convert factors to character first
  if (is.factor(x)) {
    x <- as.character(x)
  }
  
  suppressWarnings(as.numeric(x))
}

make_informal_transport <- function(data, columns) {
  
  # Check that all specified columns exist
  missing_columns <- setdiff(
    unname(columns),
    names(data)
  )
  
  if (length(missing_columns) > 0) {
    stop(
      paste(
        "The following columns are missing:",
        paste(missing_columns, collapse = ", ")
      )
    )
  }
  
  
  # Extract and clean required variables
  pas <- clean_code(
    data[[columns[["pas"]]]]
  )
  
  nic <- clean_code(
    data[[columns[["nic"]]]]
  )
  
  enterprise <- clean_code(
    data[[columns[["enterprise"]]]]
  )
  
  enterprise_size <- clean_code(
    data[[columns[["enterprise_size"]]]]
  )
  
  social_security <- clean_code(
    data[[columns[["social_security"]]]]
  )
  
  
  # Convert NIC to five-digit character codes
  nic5 <- ifelse(
    is.na(nic),
    NA_character_,
    sprintf("%05d", nic)
  )
  
  
  # Passenger land transport
  passenger_transport <-
    substr(nic5, 1, 4) %in% c("4921", "4922")
  
  
  # Informal wage/casual employees
  informal_employee <-
    pas %in% c(31, 41, 51) &
    social_security == 8
  
  
  # Informal self-employed workers
  informal_self_employed <-
    pas == 21 |
    (
      pas %in% c(11, 12) &
        enterprise %in% c(1, 2, 3, 4) &
        enterprise_size %in% c(1, 2)
    )
  # Informal employment in any industry
  informal_status <-
    informal_employee |
    informal_self_employed


  # Flag employed observations whose informality is unknown
  informal_unknown <-
    (
      pas %in% c(31, 41, 51) &
      !(social_security %in% 1:8)
    ) |
    (
      pas %in% c(11, 12) &
      (
        is.na(enterprise) |
        is.na(enterprise_size)
      )
    )

  informal_unknown[is.na(informal_unknown)] <- FALSE


  # Convert missing transport/informal results to FALSE
  passenger_transport[is.na(passenger_transport)] <- FALSE
  informal_status[is.na(informal_status)] <- FALSE


  # Add the three binary outcome variables
  data$transport <- as.integer(
    passenger_transport
  )

  data$informal <- as.integer(
    informal_status
  )

  data$informal_transport <- as.integer(
    passenger_transport &
    informal_status
  )


  # Retain an audit flag for uncertain classifications
  data$informal_unknown <- as.integer(
    informal_unknown
  )


  # Return every original observation
  data
}
```

Identifying which columns to clean/look for to identify transport and informal workers.
```{r}
columns17 <- c(
  pas = "b5pt1q3_per_fv",
  nic = "b5pt1q5_per_fv",
  enterprise = "b5pt1q9_per_fv",
  enterprise_size = "b5pt1q10_per_fv",
  social_security = "b5pt1q13_per_fv"
)


columns18_19 <- c(
  pas = "b5pt1q3_per_fv",
  nic = "b5pt1q5_per_fv",
  enterprise = "b5pt1q9_per_fv",
  enterprise_size = "b5pt1q10_per_fv",
  social_security = "b5pt1q13_per_fv"
)


columns19_20 <- c(
  pas = "b5pt1q3_per_fv",
  nic = "b5pt1q5_per_fv",
  enterprise = "b5pt1q9_per_fv",
  enterprise_size = "b5pt1q10_per_fv",
  social_security = "b5pt1q13_per_fv"
)


columns20_21 <- c(
  pas = "b5pt1q3_perv1",
  nic = "b5pt1q5_perv1",
  enterprise = "b5pt1q9_perv1",
  enterprise_size = "b5pt1q10_perv1",
  social_security = "b5pt1q13_perv1"
)


columns22 <- c(
  pas = "b5pt1q3_cperv1",
  nic = "b5pt1q5_cperv1",
  enterprise = "b5pt1q9_cperv1",
  enterprise_size = "b5pt1q10_cperv1",
  social_security = "b5pt1q13_cperv1"
)


columns23 <- c(
  pas = "b5pt1q3_cperv1",
  nic = "b5pt1q5_cperv1",
  enterprise = "b5pt1q9_cperv1",
  enterprise_size = "b5pt1q10_cperv1",
  social_security = "b5pt1q13_cperv1"
)


columns24 <- c(
  pas = "Principal_Status_Code",
  nic = "Principal_Industry_Code",
  enterprise = "Principal_Enterprise_Type",
  enterprise_size = "Principal_Workers_Count",
  social_security = "Principal_Social_Security"
)


columns25 <- c(
  pas = "pas",
  nic = "ind_pas",
  enterprise = "etyp_pas",
  enterprise_size = "wrkr_pas",
  social_security = "ssec_pas"
)

```

Function for adding state identifiers for diff-in-diff
```{r}
add_district_id <- function(
    data,
    state_column,
    district_column
) {

  # Confirm the two columns exist
  missing_columns <- setdiff(
    c(state_column, district_column),
    names(data)
  )

  if (length(missing_columns) > 0) {
    stop(
      paste(
        "Missing columns:",
        paste(missing_columns, collapse = ", ")
      )
    )
  }


  state_code <- clean_code(
    data[[state_column]]
  )

  district_code <- clean_code(
    data[[district_column]]
  )


  # Standardised versions of the original codes
  data$state_code_std <- ifelse(
    is.na(state_code),
    NA_character_,
    sprintf("%02d", state_code)
  )

  data$district_code_std <- ifelse(
    is.na(district_code),
    NA_character_,
    sprintf("%03d", district_code)
  )


  # Raw state–district identifier
  data$district_id_raw <- ifelse(
    is.na(state_code) | is.na(district_code),
    NA_character_,
    paste0(
      data$state_code_std,
      "_",
      data$district_code_std
    )
  )


  data
}
```

Applying the functions to new data frames for diff-in-diff.
```{r}
#Create data frames for informal transport for each year
analysis17_18 <- make_informal_transport(
  data17_18_fv,
  columns17
)

analysis18_19 <- make_informal_transport(
  data18_19_fv,
  columns18_19
)

analysis19_20 <- make_informal_transport(
  data19_20_fv,
  columns19_20
)

analysis20_21 <- make_informal_transport(
  data20_21_fv,
  columns20_21
)

analysis22 <- make_informal_transport(
  data22,
  columns22
)

analysis23 <- make_informal_transport(
  data23,
  columns23
)

analysis24 <- make_informal_transport(
  data24,
  columns24
)

analysis25 <- make_informal_transport(
  data25,
  columns25
)

#Adding unique district identifiers
analysis17_18 <- add_district_id(
  analysis17_18,
  state_column = "state_per_fv",
  district_column = "b1q4_per_fv"
)

analysis18_19 <- add_district_id(
  analysis18_19,
  state_column = "state_per_fv",
  district_column = "b1q4_per_fv"
)

analysis19_20 <- add_district_id(
  analysis19_20,
  state_column = "state_per_fv",
  district_column = "b1q4_per_fv"
)

analysis20_21 <- add_district_id(
  analysis20_21,
  state_column = "state_perv1",
  district_column = "distcode_perv1"
)

analysis22 <- add_district_id(
  analysis22,
  state_column = "state_cperv1",
  district_column = "distcode_cperv1"
)

analysis23 <- add_district_id(
  analysis23,
  state_column = "state_cperv1",
  district_column = "distcode_cperv1"
)

analysis24 <- add_district_id(
  analysis24,
  state_column = "State_UT_Code",
  district_column = "District_Code"
)

analysis25 <- add_district_id(
  analysis25,
  state_column = "st",
  district_column = "dc"
)
```

Functions for quarter and year identifiers, so they can be in chronological order.
```{r}
# ============================================================
# Extract the quarter number from values such as Q1, Q2, etc.
# ============================================================

get_plfs_quarter_number <- function(x) {

  # If the quarter variable is Stata-labelled,
  # convert it to the displayed label first
  if (
    inherits(x, "haven_labelled") ||
    inherits(x, "labelled")
  ) {
    x <- haven::as_factor(
      x,
      levels = "labels"
    )
  }

  x <- trimws(as.character(x))

  # Extract the numeric part:
  # Q1 -> 1, Q7 -> 7
  suppressWarnings(
    as.integer(
      gsub("[^0-9]", "", x)
    )
  )
}


# ============================================================
# Legacy PLFS quarter conversion
#
# Q1/Q5 = July-September      = calendar quarter 3
# Q2/Q6 = October-December   = calendar quarter 4
# Q3/Q7 = January-March      = calendar quarter 1
# Q4/Q8 = April-June         = calendar quarter 2
# ============================================================

convert_legacy_quarter <- function(q_number) {

  quarter <- rep(
    NA_integer_,
    length(q_number)
  )

  quarter[q_number %in% c(1, 5)] <- 3L
  quarter[q_number %in% c(2, 6)] <- 4L
  quarter[q_number %in% c(3, 7)] <- 1L
  quarter[q_number %in% c(4, 8)] <- 2L

  quarter
}

add_annual_quarter <- function(
    data,
    quarter_column,
    start_year
) {

  # Check that the quarter column exists
  if (!(quarter_column %in% names(data))) {
    stop(
      paste(
        "Quarter column not found:",
        quarter_column
      )
    )
  }


  # Read the original PLFS quarter code
  plfs_quarter_number <- get_plfs_quarter_number(
    data[[quarter_column]]
  )


  # Convert Q1-Q8 into calendar quarters 1-4
  data$quarter <- convert_legacy_quarter(
    plfs_quarter_number
  )


  # July-December belong to the starting year.
  # January-June belong to the following year.
  data$year <- ifelse(
    plfs_quarter_number %in% c(3, 4, 7, 8),
    start_year + 1L,
    start_year
  )


  # Convert year back to integer because ifelse can alter type
  data$year <- as.integer(data$year)


  # Create a readable period label
  data$year_quarter <- ifelse(
    is.na(data$year) | is.na(data$quarter),
    NA_character_,
    paste0(
      data$year,
      "_Q",
      data$quarter
    )
  )


  # Numeric quarterly index:
  # useful for sorting observations chronologically
  data$period_index <- ifelse(
    is.na(data$year) | is.na(data$quarter),
    NA_integer_,
    data$year * 4L + data$quarter
  )


  # Retain the original PLFS quarter number for auditing
  data$plfs_quarter_code <- plfs_quarter_number


  data
}

add_calendar_year_quarter <- function(
    data,
    quarter_column,
    calendar_year
) {

  if (!(quarter_column %in% names(data))) {
    stop(
      paste(
        "Quarter column not found:",
        quarter_column
      )
    )
  }


  plfs_quarter_number <- get_plfs_quarter_number(
    data[[quarter_column]]
  )


  # The year is fixed because this is a calendar-year file
  data$year <- as.integer(calendar_year)


  # Convert the rolling Q1-Q8 code to calendar quarter 1-4
  data$quarter <- convert_legacy_quarter(
    plfs_quarter_number
  )


  data$year_quarter <- ifelse(
    is.na(data$quarter),
    NA_character_,
    paste0(
      data$year,
      "_Q",
      data$quarter
    )
  )


  data$period_index <- ifelse(
    is.na(data$quarter),
    NA_integer_,
    data$year * 4L + data$quarter
  )


  data$plfs_quarter_code <- plfs_quarter_number


  data
}

add_2025_quarter <- function(
    data,
    quarter_column = "qtr"
) {

  if (!(quarter_column %in% names(data))) {
    stop(
      paste(
        "Quarter column not found:",
        quarter_column
      )
    )
  }


  quarter_number <- get_plfs_quarter_number(
    data[[quarter_column]]
  )


  # In 2025, Q1-Q4 correspond directly to calendar quarters
  valid_quarter <- quarter_number %in% 1:4

  data$year <- 2025L

  data$quarter <- ifelse(
    valid_quarter,
    quarter_number,
    NA_integer_
  )

  data$quarter <- as.integer(
    data$quarter
  )


  data$year_quarter <- ifelse(
    is.na(data$quarter),
    NA_character_,
    paste0(
      "2025_Q",
      data$quarter
    )
  )


  data$period_index <- ifelse(
    is.na(data$quarter),
    NA_integer_,
    data$year * 4L + data$quarter
  )


  data$plfs_quarter_code <- quarter_number


  data
}
```

Applying quarter-year functions.
```{r}

# 2017-18: Q1-Q4, July 2017-June 2018
analysis17_18 <- add_annual_quarter(
  data = analysis17_18,
  quarter_column = "quarter_per_fv",
  start_year = 2017
)


# 2018-19: Q5-Q8, July 2018-June 2019
analysis18_19 <- add_annual_quarter(
  data = analysis18_19,
  quarter_column = "quarter_per_fv",
  start_year = 2018
)


# 2019-20: Q1-Q4, July 2019-June 2020
analysis19_20 <- add_annual_quarter(
  data = analysis19_20,
  quarter_column = "quarter_per_fv",
  start_year = 2019
)


# 2020-21: Q5-Q8, July 2020-June 2021
analysis20_21 <- add_annual_quarter(
  data = analysis20_21,
  quarter_column = "qtr_perv1",
  start_year = 2020
)


# Calendar year 2022
analysis22 <- add_calendar_year_quarter(
  data = analysis22,
  quarter_column = "qtr_cperv1",
  calendar_year = 2022
)


# Calendar year 2023
analysis23 <- add_calendar_year_quarter(
  data = analysis23,
  quarter_column = "qtr_cperv1",
  calendar_year = 2023
)


# Calendar year 2024
analysis24 <- add_calendar_year_quarter(
  data = analysis24,
  quarter_column = "Quarter",
  calendar_year = 2024
)


# Calendar year 2025
analysis25 <- add_2025_quarter(
  data = analysis25,
  quarter_column = "qtr"
)
```

Metro binary and urban binary
```{r}
# ============================================================
# Nationwide district-quarter metro and urban indicators
# ============================================================

library(dplyr)
library(tidyr)
library(purrr)
library(readxl)
library(stringr)
library(tibble)
library(here)



# ============================================================
# 1. Add the PLFS urban/rural indicator
#
# PLFS sector code:
# 1 = rural
# 2 = urban
# ============================================================

add_urban_flag <- function(
    data,
    sector_column
) {

  if (!(sector_column %in% names(data))) {
    stop(
      paste(
        "Sector column not found:",
        sector_column
      )
    )
  }

  sector_code <- clean_code(
    data[[sector_column]]
  )

  data$sector_code <- as.integer(
    sector_code
  )

  data$urban <- case_when(
    sector_code == 2 ~ 1L,
    sector_code == 1 ~ 0L,
    TRUE ~ NA_integer_
  )

  data$rural <- case_when(
    sector_code == 1 ~ 1L,
    sector_code == 2 ~ 0L,
    TRUE ~ NA_integer_
  )

  data
}


analysis17_18 <- add_urban_flag(
  analysis17_18,
  sector_column = "b1q3_per_fv"
)

analysis18_19 <- add_urban_flag(
  analysis18_19,
  sector_column = "b1q3_per_fv"
)

analysis19_20 <- add_urban_flag(
  analysis19_20,
  sector_column = "b1q3_per_fv"
)

analysis20_21 <- add_urban_flag(
  analysis20_21,
  sector_column = "b1q3_perv1"
)

analysis22 <- add_urban_flag(
  analysis22,
  sector_column = "b1q3_cperv1"
)

analysis23 <- add_urban_flag(
  analysis23,
  sector_column = "b1q3_cperv1"
)

analysis24 <- add_urban_flag(
  analysis24,
  sector_column = "Sector"
)

analysis25 <- add_urban_flag(
  analysis25,
  sector_column = "sec"
)


# ============================================================
# 2. Download the official MoSPI district-code workbooks
#
# This section needs an internet connection on the first run.
# The files are then retained locally.
# ============================================================

district_crosswalk_folder <- here(
  "Data",
  "Crosswalks",
  "PLFS_District_Codes"
)

dir.create(
  district_crosswalk_folder,
  recursive = TRUE,
  showWarnings = FALSE
)


district_file_information <- tribble(
  ~panel,   ~file_name,                  ~download_url,

  "panel1", "plfs_district_panel1.xlsx",
  "https://microdata.gov.in/NADA/index.php/catalog/204/download/2662",

  "panel2", "plfs_district_panel2.xlsx",
  "https://microdata.gov.in/NADA/index.php/catalog/217/download/2665",

  "panel3", "plfs_district_panel3.xlsx",
  "https://microdata.gov.in/NADA/index.php/catalog/210/download/2667",

  "panel4", "plfs_district_panel4.xlsx",
  "https://microdata.gov.in/NADA/index.php/catalog/213/download/2668",

  "y2025",  "plfs_district_2025.xlsx",
  "https://microdata.gov.in/NADA/index.php/catalog/284/download/5920"
) %>%
  mutate(
    local_path = file.path(
      district_crosswalk_folder,
      file_name
    )
  )


download_government_file <- function(
    download_url,
    local_path
) {

  needs_download <-
    !file.exists(local_path) ||
    is.na(file.info(local_path)$size) ||
    file.info(local_path)$size < 5000

  if (needs_download) {

    message(
      "Downloading official MoSPI file: ",
      basename(local_path)
    )

    tryCatch(
      {
        download.file(
          url = download_url,
          destfile = local_path,
          mode = "wb",
          quiet = FALSE
        )
      },
      error = function(e) {
        stop(
          paste0(
            "Could not download:\n",
            download_url,
            "\n\nDownload it manually and save it as:\n",
            local_path,
            "\n\nOriginal error:\n",
            e$message
          )
        )
      }
    )
  }

  local_path
}


district_file_information$local_path <- map2_chr(
  district_file_information$download_url,
  district_file_information$local_path,
  download_government_file
)


# ============================================================
# 3. Functions for reading the official district workbooks
# ============================================================

clean_workbook_names <- function(x) {

  x <- as.character(x)

  x[is.na(x)] <- ""

  x <- tolower(
    trimws(x)
  )

  x <- str_replace_all(
    x,
    "[^a-z0-9]+",
    "_"
  )

  x <- str_replace_all(
    x,
    "^_+|_+$",
    ""
  )

  blank_names <- which(x == "")

  if (length(blank_names) > 0) {
    x[blank_names] <- paste0(
      "unnamed_",
      seq_along(blank_names)
    )
  }

  make.unique(
    x,
    sep = "_"
  )
}


# ============================================================
# Identify columns inside the MoSPI district-code workbooks
# ============================================================

find_workbook_column <- function(
    column_names,
    patterns,
    description
) {

  for (pattern in patterns) {

    matches <- grep(
      pattern,
      column_names,
      value = TRUE,
      ignore.case = TRUE
    )

    if (length(matches) > 0) {
      return(matches[1])
    }
  }

  stop(
    paste0(
      "Could not identify the ",
      description,
      " column.\nAvailable columns: ",
      paste(column_names, collapse = ", ")
    )
  )
}


# ============================================================
# Combine one or two workbook rows into possible headers
# ============================================================

make_header_candidate <- function(
    raw_sheet,
    start_row,
    number_of_rows = 1
) {

  end_row <- min(
    start_row + number_of_rows - 1,
    nrow(raw_sheet)
  )

  header_rows <- raw_sheet[
    start_row:end_row,
    ,
    drop = FALSE
  ]

  combined_headers <- vapply(
    seq_len(ncol(header_rows)),
    function(column_number) {

      values <- as.character(
        header_rows[[column_number]]
      )

      values <- values[
        !is.na(values) &
          trimws(values) != ""
      ]

      paste(
        values,
        collapse = " "
      )
    },
    character(1)
  )

  combined_headers
}


# ============================================================
# Score whether a row looks like the real table header
# ============================================================

score_header_candidate <- function(header_values) {

  header_text <- tolower(
    paste(
      header_values,
      collapse = " | "
    )
  )

  has_state_code <- grepl(
    "state.*code|state.?ut.*code|st.*code",
    header_text
  )

  has_district_code <- grepl(
    "district.*code|dist.*code",
    header_text
  )

  has_district_name <- grepl(
    "district.*name|name.*district|name of district",
    header_text
  )

  score <-
    5L * as.integer(has_state_code) +
    5L * as.integer(has_district_code) +
    2L * as.integer(has_district_name)

  list(
    score = score,
    has_state_code = has_state_code,
    has_district_code = has_district_code
  )
}


# ============================================================
# Find the actual header row
# ============================================================

find_district_header <- function(
    raw_sheet,
    maximum_rows = 40
) {

  rows_to_check <- seq_len(
    min(
      maximum_rows,
      nrow(raw_sheet)
    )
  )

  candidates <- list()
  candidate_number <- 1L


  for (header_depth in c(1L, 2L)) {

    for (header_row in rows_to_check) {

      if (
        header_row + header_depth - 1L >
          nrow(raw_sheet)
      ) {
        next
      }

      header_values <- make_header_candidate(
        raw_sheet = raw_sheet,
        start_row = header_row,
        number_of_rows = header_depth
      )

      header_score <- score_header_candidate(
        header_values
      )

      candidates[[candidate_number]] <- list(
        header_row = header_row,
        header_depth = header_depth,
        header_values = header_values,
        score = header_score$score,
        valid =
          header_score$has_state_code &&
          header_score$has_district_code
      )

      candidate_number <- candidate_number + 1L
    }
  }


  valid_candidates <- candidates[
    vapply(
      candidates,
      function(x) x$valid,
      logical(1)
    )
  ]


  if (length(valid_candidates) == 0) {

    preview <- raw_sheet[
      seq_len(min(10, nrow(raw_sheet))),
      ,
      drop = FALSE
    ]

    print(preview)

    stop(
      paste(
        "Could not find the actual district-table header.",
        "The first ten workbook rows were printed above."
      )
    )
  }


  candidate_scores <- vapply(
    valid_candidates,
    function(x) x$score,
    integer(1)
  )

  valid_candidates[[
    which.max(candidate_scores)
  ]]
}


extract_first_integer <- function(x) {

  suppressWarnings(
    as.integer(
      str_extract(
        as.character(x),
        "[0-9]+"
      )
    )
  )
}


# ============================================================
# Read one sheet from a MoSPI district-code workbook
# ============================================================

read_district_sheet <- function(
    file_path,
    sheet_name,
    panel_name
) {

  raw_sheet <- suppressMessages(
    readxl::read_excel(
      path = file_path,
      sheet = sheet_name,
      col_names = FALSE,
      col_types = "text"
    )
  )

  raw_sheet <- as.data.frame(
    raw_sheet,
    stringsAsFactors = FALSE
  )


  if (
    nrow(raw_sheet) == 0 ||
    ncol(raw_sheet) == 0
  ) {
    return(
      tibble::tibble()
    )
  }


  # Find the genuine table header rather than the title
  header_information <- find_district_header(
    raw_sheet
  )

  header_row <- header_information$header_row
  header_depth <- header_information$header_depth

  header_values <- header_information$header_values

  first_data_row <-
    header_row +
    header_depth


  if (first_data_row > nrow(raw_sheet)) {
    return(
      tibble::tibble()
    )
  }


  body <- raw_sheet[
    first_data_row:nrow(raw_sheet),
    ,
    drop = FALSE
  ]


  names(body) <- clean_workbook_names(
    header_values
  )


  # Remove completely empty columns
  nonempty_columns <- vapply(
    body,
    function(x) {
      any(
        !is.na(x) &
          trimws(as.character(x)) != ""
      )
    },
    logical(1)
  )

  body <- body[
    ,
    nonempty_columns,
    drop = FALSE
  ]


  # Identify the three required columns
  state_code_column <- find_workbook_column(
    names(body),
    patterns = c(
      "state_ut.*code",
      "state.*code",
      "^st.*code",
      "^state_ut$"
    ),
    description = "state code"
  )


  district_code_column <- find_workbook_column(
    names(body),
    patterns = c(
      "district.*code",
      "^dist.*code"
    ),
    description = "district code"
  )


  district_name_column <- find_workbook_column(
    names(body),
    patterns = c(
      "district.*name",
      "name.*district",
      "name_of_district",
      "^district_name$",
      "^district$"
    ),
    description = "district name"
  )


  # Replace blank state entries with NA before filling downward
  body[[state_code_column]] <- as.character(
    body[[state_code_column]]
  )

  body[[state_code_column]][
    trimws(body[[state_code_column]]) == ""
  ] <- NA_character_


  # Workbooks often display the state code only on the first
  # district row for each state
  body <- tidyr::fill(
    body,
    dplyr::all_of(state_code_column),
    .direction = "down"
  )


  result <- tibble::tibble(
    panel = panel_name,

    state_code = extract_first_integer(
      body[[state_code_column]]
    ),

    district_code = extract_first_integer(
      body[[district_code_column]]
    ),

    district_name = stringr::str_squish(
      as.character(
        body[[district_name_column]]
      )
    )
  ) %>%
    dplyr::filter(
      !is.na(state_code),
      !is.na(district_code),
      !is.na(district_name),
      district_name != "",
      !stringr::str_detect(
        tolower(district_name),
        "^total"
      )
    )


  result
}


read_district_workbook <- function(
    file_path,
    panel_name
) {

  sheet_names <- excel_sheets(
    file_path
  )

  result <- map_dfr(
    sheet_names,
    ~ read_district_sheet(
      file_path = file_path,
      sheet_name = .x,
      panel_name = panel_name
    )
  )

  if (nrow(result) == 0) {
    stop(
      paste(
        "No district codes could be read from:",
        file_path
      )
    )
  }

  result
}


official_district_lookup <- map2_dfr(
  district_file_information$local_path,
  district_file_information$panel,
  read_district_workbook
) %>%
  distinct(
    panel,
    state_code,
    district_code,
    .keep_all = TRUE
  )


# ============================================================
# 4. Standardise district names
# ============================================================

normalise_district_name <- function(x) {

  x <- iconv(
    as.character(x),
    from = "",
    to = "ASCII//TRANSLIT"
  )

  x <- tolower(x)

  x <- str_replace_all(
    x,
    "&",
    " and "
  )

  x <- str_replace_all(
    x,
    "[^a-z0-9]+",
    " "
  )

  x <- str_squish(x)


  case_when(
    x %in% c(
      "gurgaon",
      "gurugram"
    ) ~ "gurugram",

    x %in% c(
      "gautam budh nagar",
      "gautam buddh nagar",
      "gautam buddha nagar"
    ) ~ "gautam buddha nagar",

    x %in% c(
      "bangalore urban",
      "bengaluru urban",
      "bangalore"
    ) ~ "bengaluru urban",

    x %in% c(
      "ranga reddy",
      "rangareddy"
    ) ~ "rangareddy",

    x %in% c(
      "medchal malkajgiri",
      "medchal malkajgiri district"
    ) ~ "medchal malkajgiri",

    x %in% c(
      "north twenty four parganas",
      "north 24 parganas",
      "24 parganas north",
      "north 24 paraganas"
    ) ~ "north 24 parganas",

    x %in% c(
      "south twenty four parganas",
      "south 24 parganas",
      "24 parganas south",
      "south 24 paraganas"
    ) ~ "south 24 parganas",

    x %in% c(
      "calcutta",
      "kolkata"
    ) ~ "kolkata",

    x %in% c(
      "kancheepuram",
      "kanchipuram",
      "kanchipuram district"
    ) ~ "kanchipuram",

    x %in% c(
      "ahmadabad",
      "ahmedabad"
    ) ~ "ahmedabad",

    x %in% c(
      "bombay suburban",
      "mumbai suburban"
    ) ~ "mumbai suburban",

    x %in% c(
      "bombay city",
      "mumbai city"
    ) ~ "mumbai city",

    x %in% c(
      "kanpur",
      "kanpur nagar"
    ) ~ "kanpur nagar",

    x %in% c(
      "ahmadnagar",
      "ahmednagar",
      "ahilyanagar"
    ) ~ "ahilyanagar",

    TRUE ~ x
  )
}


official_district_lookup <- official_district_lookup %>%
  mutate(
    district_name_key = normalise_district_name(
      district_name
    )
  )


# ============================================================
# 5. Attach official district names to one PLFS dataset
# ============================================================

attach_official_district_name <- function(
    data,
    panel
) {

  if (length(panel) == 1) {
    panel <- rep(
      panel,
      nrow(data)
    )
  }

  if (length(panel) != nrow(data)) {
    stop(
      "The panel vector must have one value per observation."
    )
  }

  data$.original_row_order <- seq_len(
    nrow(data)
  )

  data$.district_panel <- as.character(
    panel
  )

  data$.state_code_join <- suppressWarnings(
    as.integer(
      data$state_code_std
    )
  )

  data$.district_code_join <- suppressWarnings(
    as.integer(
      data$district_code_std
    )
  )


  data <- data %>%
    select(
      -any_of(
        c(
          "district_name",
          "district_name_key"
        )
      )
    ) %>%
    left_join(
      official_district_lookup %>%
        select(
          panel,
          state_code,
          district_code,
          district_name,
          district_name_key
        ),

      by = c(
        ".district_panel" = "panel",
        ".state_code_join" = "state_code",
        ".district_code_join" = "district_code"
      )
    ) %>%
    arrange(
      .original_row_order
    ) %>%
    select(
      -.original_row_order,
      -.state_code_join,
      -.district_code_join,
      -.district_panel
    )

  data
}


# Panel assignments
analysis17_18 <- attach_official_district_name(
  analysis17_18,
  panel = "panel1"
)

analysis18_19 <- attach_official_district_name(
  analysis18_19,
  panel = "panel1"
)

analysis19_20 <- attach_official_district_name(
  analysis19_20,
  panel = "panel2"
)

analysis20_21 <- attach_official_district_name(
  analysis20_21,
  panel = "panel2"
)

analysis22 <- attach_official_district_name(
  analysis22,
  panel = "panel3"
)


# Calendar 2023 combines the second half of Panel 3
# and the first half of Panel 4.
analysis23_panel <- ifelse(
  analysis23$quarter %in% c(1L, 2L),
  "panel3",
  "panel4"
)

analysis23 <- attach_official_district_name(
  analysis23,
  panel = analysis23_panel
)


analysis24 <- attach_official_district_name(
  analysis24,
  panel = "panel4"
)

analysis25 <- attach_official_district_name(
  analysis25,
  panel = "y2025"
)


# ============================================================
# 6. Official source registry
#
# Source URLs are retained in the resulting datasets so each
# positive metro classification can be audited.
# ============================================================

metro_sources <- tribble(
  ~source_id, ~metro_source_url,

  "DMRC",
  "https://www.delhimetrorail.com/pages/en/introduction",

  "JMRC",
  "https://transport.rajasthan.gov.in/content/transportportal/en/metro.html",

  "BMRCL",
  "https://english.bmrc.co.in/",

  "CMRL",
  "https://chennaimetrorail.org/",

  "KMRL",
  "https://kochimetro.org/",

  "KOLKATA_METRO",
  "https://mtp.indianrailways.gov.in/",

  "MMRDA",
  "https://mmrda.maharashtra.gov.in/en/projects/transport/metro-line-1/overview",

  "HMRL",
  "https://hmrl.co.in/",

  "PIB_LUCKNOW",
  "https://www.pib.gov.in/newsite/PrintRelease.aspx?lang=2&reg=48&relid=170511",

  "PIB_BAHADURGARH",
  "https://www.pib.gov.in/Pressreleaseshare.aspx?PRID=1536394",

  "PIB_AHMEDABAD",
  "https://www.pib.gov.in/PressReleasePage.aspx?PRID=1567375",

  "PIB_NAGPUR",
  "https://www.pib.gov.in/PressReleasePage.aspx?PRID=1568013",

  "PIB_KANPUR",
  "https://www.pib.gov.in/PressReleasePage.aspx?PRID=1844647",

  "PIB_PUNE",
  "https://www.pib.gov.in/PressReleasePage.aspx?PRID=1803340",

  "CIDCO",
  "https://cidco.maharashtra.gov.in/",

  "PIB_AGRA_HOWRAH",
  "https://www.pib.gov.in/PressReleasePage.aspx?PRID=2011999",

  "PIB_GANDHINAGAR",
  "https://www.pib.gov.in/PressReleasePage.aspx?PRID=2055368",

  "MMRCL",
  "https://mmrcl.com/",

  "PIB_INDORE",
  "https://www.pib.gov.in/PressReleasePage.aspx?PRID=2132584",

  "PIB_BHOPAL",
  "https://www.pib.gov.in/PressReleaseIframePage.aspx?PRID=2207030&lang=2&reg=48"
)


# ============================================================
# 7. District-level metro-opening crosswalk
#
# pre_sample = TRUE means that the district already had an
# operational metro before the first PLFS quarter, 2017 Q3.
#
# manual_review = 1 flags cases affected by changing district
# boundaries or metro lines crossing district boundaries.
# ============================================================

metro_district_crosswalk <- tribble(
  ~state_code, ~district_name,          ~metro_system,
  ~opening_year, ~opening_quarter, ~pre_sample,
  ~source_id, ~manual_review, ~metro_note,

  6L, "Gurugram", "Delhi Metro / Rapid Metro",
  2010L, 2L, TRUE,
  "DMRC", 0L,
  "Operational before the PLFS sample.",

  6L, "Faridabad", "Delhi Metro",
  2015L, 3L, TRUE,
  "DMRC", 0L,
  "Operational before the PLFS sample.",

  6L, "Jhajjar", "Delhi Metro: Bahadurgarh extension",
  2018L, 2L, FALSE,
  "PIB_BAHADURGARH", 0L,
  "Bahadurgarh is within Jhajjar district.",


  9L, "Ghaziabad", "Delhi Metro",
  2011L, 3L, TRUE,
  "DMRC", 0L,
  "Vaishali metro was operational before the PLFS sample.",

  9L, "Gautam Buddha Nagar", "Delhi Metro / Noida Metro",
  2009L, 4L, TRUE,
  "DMRC", 0L,
  "Noida had metro service before the PLFS sample.",

  9L, "Lucknow", "Lucknow Metro",
  2017L, 3L, FALSE,
  "PIB_LUCKNOW", 0L,
  "Priority corridor inaugurated in September 2017.",

  9L, "Kanpur Nagar", "Kanpur Metro",
  2021L, 4L, FALSE,
  "PIB_KANPUR", 0L,
  "Commercial operation began in December 2021.",

  9L, "Agra", "Agra Metro",
  2024L, 1L, FALSE,
  "PIB_AGRA_HOWRAH", 0L,
  "First operational stretch inaugurated in March 2024.",


  8L, "Jaipur", "Jaipur Metro",
  2015L, 2L, TRUE,
  "JMRC", 0L,
  "Operational before the PLFS sample.",


  24L, "Ahmedabad", "Ahmedabad Metro",
  2019L, 1L, FALSE,
  "PIB_AHMEDABAD", 0L,
  "First operational section inaugurated in March 2019.",

  24L, "Gandhinagar", "Ahmedabad-Gandhinagar Metro",
  2024L, 3L, FALSE,
  "PIB_GANDHINAGAR", 0L,
  "Phase II inaugurated in September 2024.",


  27L, "Mumbai Suburban", "Mumbai Metro",
  2014L, 2L, TRUE,
  "MMRDA", 0L,
  "Mumbai Metro Line 1 was operational before the sample.",

  27L, "Mumbai City", "Mumbai Metro Line 3",
  2025L, 2L, FALSE,
  "MMRCL", 1L,
  paste(
    "Southward extension reached Mumbai City district",
    "during 2025 Q2; review against final station boundaries."
  ),

  27L, "Thane", "Navi Mumbai Metro",
  2023L, 4L, FALSE,
  "CIDCO", 1L,
  paste(
    "Navi Mumbai Metro crosses the Thane-Raigad boundary;",
    "retain for boundary audit."
  ),

  27L, "Raigad", "Navi Mumbai Metro",
  2023L, 4L, FALSE,
  "CIDCO", 1L,
  paste(
    "Navi Mumbai Metro crosses the Thane-Raigad boundary;",
    "retain for boundary audit."
  ),

  27L, "Nagpur", "Nagpur Metro",
  2019L, 1L, FALSE,
  "PIB_NAGPUR", 0L,
  "First operational section inaugurated in March 2019.",

  27L, "Pune", "Pune Metro",
  2022L, 1L, FALSE,
  "PIB_PUNE", 0L,
  "First operational sections inaugurated in March 2022.",


  29L, "Bengaluru Urban", "Namma Metro",
  2011L, 4L, TRUE,
  "BMRCL", 0L,
  "Operational before the PLFS sample.",


  36L, "Hyderabad", "Hyderabad Metro",
  2017L, 4L, FALSE,
  "HMRL", 0L,
  "First operational corridors opened in November 2017.",

  36L, "Rangareddy", "Hyderabad Metro",
  2017L, 4L, FALSE,
  "HMRL", 1L,
  "First corridor crosses Hyderabad metropolitan district boundaries.",

  36L, "Medchal Malkajgiri", "Hyderabad Metro",
  2017L, 4L, FALSE,
  "HMRL", 1L,
  "First corridor crosses Hyderabad metropolitan district boundaries.",


  33L, "Chennai", "Chennai Metro",
  2015L, 2L, TRUE,
  "CMRL", 0L,
  "Operational before the PLFS sample.",

  33L, "Kanchipuram", "Chennai Metro",
  2015L, 2L, TRUE,
  "CMRL", 1L,
  paste(
    "Review because Chennai district boundaries expanded",
    "during the study period."
  ),


  32L, "Ernakulam", "Kochi Metro",
  2017L, 2L, TRUE,
  "KMRL", 0L,
  "Kochi Metro opened immediately before the first PLFS quarter.",


  19L, "Kolkata", "Kolkata Metro",
  1984L, 4L, TRUE,
  "KOLKATA_METRO", 0L,
  "Operational before the PLFS sample.",

  19L, "North 24 Parganas", "Kolkata Metro",
  1995L, 3L, TRUE,
  "KOLKATA_METRO", 1L,
  "Dum Dum and later extensions serve this district.",

  19L, "Howrah", "Kolkata Metro: East-West line",
  2024L, 1L, FALSE,
  "PIB_AGRA_HOWRAH", 0L,
  "Howrah Maidan-Esplanade section opened in March 2024.",


  23L, "Indore", "Indore Metro",
  2025L, 2L, FALSE,
  "PIB_INDORE", 0L,
  "Passenger services inaugurated in May 2025.",

  23L, "Bhopal", "Bhopal Metro",
  2025L, 4L, FALSE,
  "PIB_BHOPAL", 0L,
  "Priority corridor opened in December 2025."
) %>%
  mutate(
    district_name_key = normalise_district_name(
      district_name
    ),

    metro_pre_sample = as.integer(
      pre_sample
    ),

    metro_start_period_index = case_when(
      pre_sample ~ 0L,

      TRUE ~
        opening_year * 4L +
        opening_quarter
    ),

    metro_manual_review = as.integer(
      manual_review
    )
  ) %>%
  left_join(
    metro_sources,
    by = "source_id"
  ) %>%
  transmute(
    state_code,
    district_name_key,

    metro_system,

    metro_opening_year =
      as.integer(opening_year),

    metro_opening_quarter =
      as.integer(opening_quarter),

    metro_start_period_index =
      as.integer(metro_start_period_index),

    metro_pre_sample,

    metro_source_id = source_id,

    metro_source_url,

    metro_manual_review,

    metro_note
  )


# ============================================================
# 8. Delhi state-level rule
#
# Every PLFS district in Delhi is treated as having metro
# infrastructure throughout the study period.
#
# The historical opening quarter differs by district, but every
# Delhi district was already served before the PLFS sample.
# ============================================================

metro_state_crosswalk <- tibble(
  state_code = 7L,

  metro_system =
    "Delhi Metro",

  metro_opening_year =
    NA_integer_,

  metro_opening_quarter =
    NA_integer_,

  metro_start_period_index =
    0L,

  metro_pre_sample =
    1L,

  metro_source_id =
    "DMRC",

  metro_manual_review =
    1L,

  metro_note = paste(
    "All Delhi districts are coded as pre-sample metro districts.",
    "District-specific historical opening dates vary."
  )
) %>%
  left_join(
    metro_sources,
    by = c(
      "metro_source_id" = "source_id"
    )
  )


# ============================================================
# 9. Add metro indicators to one dataset
# ============================================================

add_metro_flag <- function(data) {

  required_columns <- c(
    "state_code_std",
    "district_name_key",
    "year",
    "quarter",
    "period_index"
  )

  missing_columns <- setdiff(
    required_columns,
    names(data)
  )

  if (length(missing_columns) > 0) {
    stop(
      paste(
        "Missing columns:",
        paste(missing_columns, collapse = ", ")
      )
    )
  }


  # Make the function safe to rerun
  data <- data %>%
    select(
      -any_of(
        c(
          "has_metro",
          "metro_opened_this_quarter",
          "metro_urban",
          "metro_system",
          "metro_opening_year",
          "metro_opening_quarter",
          "metro_start_period_index",
          "metro_pre_sample",
          "metro_source_id",
          "metro_source_url",
          "metro_manual_review",
          "metro_note"
        )
      )
    ) %>%
    mutate(
      .state_code_num = suppressWarnings(
        as.integer(state_code_std)
      )
    )


  # Exact state-district matches
  data <- data %>%
    left_join(
      metro_district_crosswalk,

      by = c(
        ".state_code_num" = "state_code",
        "district_name_key" = "district_name_key"
      )
    )


  # State-wide rules, currently used for Delhi
  state_lookup <- metro_state_crosswalk %>%
    rename(
      metro_system_state =
        metro_system,

      metro_opening_year_state =
        metro_opening_year,

      metro_opening_quarter_state =
        metro_opening_quarter,

      metro_start_period_index_state =
        metro_start_period_index,

      metro_pre_sample_state =
        metro_pre_sample,

      metro_source_id_state =
        metro_source_id,

      metro_source_url_state =
        metro_source_url,

      metro_manual_review_state =
        metro_manual_review,

      metro_note_state =
        metro_note
    )


  data <- data %>%
    left_join(
      state_lookup,
      by = c(
        ".state_code_num" = "state_code"
      )
    ) %>%
    mutate(
      metro_system = coalesce(
        metro_system,
        metro_system_state
      ),

      metro_opening_year = coalesce(
        metro_opening_year,
        metro_opening_year_state
      ),

      metro_opening_quarter = coalesce(
        metro_opening_quarter,
        metro_opening_quarter_state
      ),

      metro_start_period_index = coalesce(
        metro_start_period_index,
        metro_start_period_index_state
      ),

      metro_pre_sample = coalesce(
        metro_pre_sample,
        metro_pre_sample_state
      ),

      metro_source_id = coalesce(
        metro_source_id,
        metro_source_id_state
      ),

      metro_source_url = coalesce(
        metro_source_url,
        metro_source_url_state
      ),

      metro_manual_review = coalesce(
        metro_manual_review,
        metro_manual_review_state
      ),

      metro_note = coalesce(
        metro_note,
        metro_note_state
      ),


      # Main district-quarter treatment indicator
      has_metro = as.integer(
        !is.na(metro_start_period_index) &
          period_index >=
          metro_start_period_index
      ),


      # Identifies the opening quarter itself
      metro_opened_this_quarter = as.integer(
        !is.na(metro_start_period_index) &
          metro_start_period_index > 0L &
          period_index ==
          metro_start_period_index
      ),


      # Recommended exposure indicator:
      # metro district AND urban PLFS respondent
      metro_urban = case_when(
        is.na(urban) ~ NA_integer_,

        has_metro == 1L &
          urban == 1L ~ 1L,

        TRUE ~ 0L
      ),


      # Zero for districts with no verified metro match
      metro_pre_sample = coalesce(
        metro_pre_sample,
        0L
      ),

      metro_manual_review = coalesce(
        metro_manual_review,
        0L
      )
    ) %>%
    select(
      -.state_code_num,

      -metro_system_state,
      -metro_opening_year_state,
      -metro_opening_quarter_state,
      -metro_start_period_index_state,
      -metro_pre_sample_state,
      -metro_source_id_state,
      -metro_source_url_state,
      -metro_manual_review_state,
      -metro_note_state
    )


  data
}


# ============================================================
# 10. Apply the metro function to every analysis dataset
# ============================================================

analysis17_18 <- add_metro_flag(
  analysis17_18
)

analysis18_19 <- add_metro_flag(
  analysis18_19
)

analysis19_20 <- add_metro_flag(
  analysis19_20
)

analysis20_21 <- add_metro_flag(
  analysis20_21
)

analysis22 <- add_metro_flag(
  analysis22
)

analysis23 <- add_metro_flag(
  analysis23
)

analysis24 <- add_metro_flag(
  analysis24
)

analysis25 <- add_metro_flag(
  analysis25
)


# ============================================================
# 11. Validate the resulting columns
# ============================================================

validate_metro_flags <- function(
    data,
    dataset_name
) {

  valid_urban_values <- all(
    is.na(data$urban) |
      data$urban %in% c(0L, 1L)
  )

  valid_metro_values <- all(
    data$has_metro %in% c(0L, 1L)
  )

  valid_opening_values <- all(
    data$metro_opened_this_quarter %in%
      c(0L, 1L)
  )

  valid_metro_urban <- all(
    is.na(data$metro_urban) |
      data$metro_urban ==
      as.integer(
        data$has_metro == 1L &
          data$urban == 1L
      )
  )

  if (!valid_urban_values) {
    stop(
      dataset_name,
      ": urban contains invalid values."
    )
  }

  if (!valid_metro_values) {
    stop(
      dataset_name,
      ": has_metro contains invalid values."
    )
  }

  if (!valid_opening_values) {
    stop(
      dataset_name,
      ": metro_opened_this_quarter contains invalid values."
    )
  }

  if (!valid_metro_urban) {
    stop(
      dataset_name,
      ": metro_urban was not constructed correctly."
    )
  }

  message(
    dataset_name,
    ": metro and urban checks passed."
  )
}


validate_metro_flags(
  analysis17_18,
  "2017-18"
)

validate_metro_flags(
  analysis18_19,
  "2018-19"
)

validate_metro_flags(
  analysis19_20,
  "2019-20"
)

validate_metro_flags(
  analysis20_21,
  "2020-21"
)

validate_metro_flags(
  analysis22,
  "2022"
)

validate_metro_flags(
  analysis23,
  "2023"
)

validate_metro_flags(
  analysis24,
  "2024"
)

validate_metro_flags(
  analysis25,
  "2025"
)


# ============================================================
# 12. Combine a small district-quarter validation table
# ============================================================

# ============================================================
# Create a standardised district-quarter validation table
# without combining all original PLFS variables
# ============================================================

prepare_metro_validation <- function(
    data,
    dataset_name
) {

  data %>%
    transmute(
      source_dataset = dataset_name,

      state_code_std = as.character(
        state_code_std
      ),

      district_name = as.character(
        district_name
      ),

      district_name_key = as.character(
        district_name_key
      ),

      year = as.integer(
        year
      ),

      quarter = as.integer(
        quarter
      ),

      year_quarter = as.character(
        year_quarter
      ),

      period_index = as.integer(
        period_index
      ),

      urban = as.integer(
        urban
      ),

      rural = as.integer(
        rural
      ),

      has_metro = as.integer(
        has_metro
      ),

      metro_opened_this_quarter = as.integer(
        metro_opened_this_quarter
      ),

      metro_urban = as.integer(
        metro_urban
      ),

      metro_system = as.character(
        metro_system
      ),

      metro_opening_year = as.integer(
        metro_opening_year
      ),

      metro_opening_quarter = as.integer(
        metro_opening_quarter
      ),

      metro_manual_review = as.integer(
        metro_manual_review
      ),

      metro_source_url = as.character(
        metro_source_url
      )
    )
}


metro_validation_data <- bind_rows(
  prepare_metro_validation(
    analysis17_18,
    "2017-18"
  ),

  prepare_metro_validation(
    analysis18_19,
    "2018-19"
  ),

  prepare_metro_validation(
    analysis19_20,
    "2019-20"
  ),

  prepare_metro_validation(
    analysis20_21,
    "2020-21"
  ),

  prepare_metro_validation(
    analysis22,
    "2022"
  ),

  prepare_metro_validation(
    analysis23,
    "2023"
  ),

  prepare_metro_validation(
    analysis24,
    "2024"
  ),

  prepare_metro_validation(
    analysis25,
    "2025"
  )
) %>%
  distinct() %>%
  arrange(
    period_index,
    state_code_std,
    district_name
  )


# Opening events occurring inside the PLFS period
metro_opening_events <- metro_validation_data %>%
  filter(
    metro_opened_this_quarter == 1L
  ) %>%
  distinct(
    district_name,
    year_quarter,
    metro_system,
    metro_source_url,
    metro_manual_review
  ) %>%
  arrange(
    year_quarter,
    district_name
  )

print(
  metro_opening_events
)


# All districts ever coded as metro districts
positive_metro_districts <- metro_validation_data %>%
  filter(
    has_metro == 1L
  ) %>%
  distinct(
    state_code_std,
    district_name,
    metro_system,
    metro_opening_year,
    metro_opening_quarter,
    metro_manual_review
  ) %>%
  arrange(
    state_code_std,
    district_name
  )

print(
  positive_metro_districts
)


# Cases that should be reviewed before the final dissertation
metro_boundary_review <- positive_metro_districts %>%
  filter(
    metro_manual_review == 1L
  )

print(
  metro_boundary_review
)


# ============================================================
# 13. Check for impossible treatment reversals
#
# Once a district receives has_metro = 1, it should never
# return to zero in a later quarter.
# ============================================================

metro_reversal_check <- metro_validation_data %>%
  group_by(
    state_code_std,
    district_name_key
  ) %>%
  arrange(
    period_index,
    .by_group = TRUE
  ) %>%
  summarise(
    treatment_reversal = any(
      diff(has_metro) < 0,
      na.rm = TRUE
    ),

    .groups = "drop"
  ) %>%
  filter(
    treatment_reversal
  )

if (nrow(metro_reversal_check) > 0) {

  print(
    metro_reversal_check
  )

  warning(
    "At least one district changes from has_metro = 1 back to 0."
  )

} else {

  message(
    "No metro treatment reversals detected."
  )
}
```



Confirm Metro indicators added after time
```{r}
# ============================================================
# Combine all annual analysis datasets into one
# individual-level repeated-cross-section DiD dataset
# ============================================================

library(dplyr)
library(stringr)
library(tibble)


# ============================================================
# 1. Confirm that metro indicators were added after time
# ============================================================

required_final_variables <- c(
  "year",
  "quarter",
  "year_quarter",
  "period_index",
  "state_code_std",
  "district_code_std",
  "district_id_raw",
  "district_name",
  "district_name_key",
  "urban",
  "rural",
  "has_metro",
  "metro_urban",
  "metro_opened_this_quarter",
  "metro_start_period_index",
  "transport",
  "informal",
  "informal_transport",
  "informal_unknown"
)


check_final_variables <- function(
    data,
    dataset_name
) {

  missing_variables <- setdiff(
    required_final_variables,
    names(data)
  )

  if (length(missing_variables) > 0) {
    stop(
      paste0(
        dataset_name,
        " is missing: ",
        paste(
          missing_variables,
          collapse = ", "
        ),
        "\n\nMake sure the quarter/year functions are run before ",
        "attach_official_district_name() and add_metro_flag()."
      )
    )
  }

  message(
    dataset_name,
    ": required variables found."
  )
}


check_final_variables(
  analysis17_18,
  "2017-18"
)

check_final_variables(
  analysis18_19,
  "2018-19"
)

check_final_variables(
  analysis19_20,
  "2019-20"
)

check_final_variables(
  analysis20_21,
  "2020-21"
)

check_final_variables(
  analysis22,
  "2022"
)

check_final_variables(
  analysis23,
  "2023"
)

check_final_variables(
  analysis24,
  "2024"
)

check_final_variables(
  analysis25,
  "2025"
)
```
```{r}
# ============================================================
# 2. Find the first available variable from a list
# ============================================================

find_first_existing <- function(
    data,
    candidates,
    variable_description,
    dataset_name
) {

  existing_variables <- candidates[
    candidates %in% names(data)
  ]

  if (length(existing_variables) == 0) {
    stop(
      paste0(
        dataset_name,
        ": could not find ",
        variable_description,
        ".\nChecked: ",
        paste(
          candidates,
          collapse = ", "
        )
      )
    )
  }

  existing_variables[1]
}


# ============================================================
# Extract an optional variable or return missing values
# ============================================================

get_optional_character <- function(
    data,
    variable_name
) {

  if (variable_name %in% names(data)) {
    return(
      as.character(
        data[[variable_name]]
      )
    )
  }

  rep(
    NA_character_,
    nrow(data)
  )
}


get_optional_integer <- function(
    data,
    variable_name
) {

  if (variable_name %in% names(data)) {
    return(
      as.integer(
        clean_code(
          data[[variable_name]]
        )
      )
    )
  }

  rep(
    NA_integer_,
    nrow(data)
  )
}
```

```{r}
# ============================================================
# 3. Retain and rename only the DiD variables
# ============================================================

prepare_did_dataset <- function(
    data,
    dataset_name,
    age_candidates,
    education_candidates,
    weight_candidates
) {

  # Find the age variable
  age_variable <- find_first_existing(
    data = data,
    candidates = age_candidates,
    variable_description = "age",
    dataset_name = dataset_name
  )


  # Find the education variable
  education_variable <- find_first_existing(
    data = data,
    candidates = education_candidates,
    variable_description = "general education level",
    dataset_name = dataset_name
  )


  # Find the survey-weight variable
  weight_variable <- find_first_existing(
    data = data,
    candidates = weight_candidates,
    variable_description = "survey multiplier",
    dataset_name = dataset_name
  )


  # Create a stable district identifier using the official
  # state code and harmonised district name.
  #
  # district_id_raw is retained separately for auditing.
  stable_district_id <- ifelse(
    is.na(data$state_code_std) |
      is.na(data$district_name_key),

    as.character(
      data$district_id_raw
    ),

    paste0(
      as.character(
        data$state_code_std
      ),
      "_",
      str_replace_all(
        as.character(
          data$district_name_key
        ),
        " ",
        "_"
      )
    )
  )


  compact_data <- tibble(
    # --------------------------------------------------------
    # Source
    # --------------------------------------------------------

    source_dataset =
      dataset_name,


    # --------------------------------------------------------
    # Time
    # --------------------------------------------------------

    year = as.integer(
      data$year
    ),

    quarter = as.integer(
      data$quarter
    ),

    year_quarter = as.character(
      data$year_quarter
    ),

    period_index = as.integer(
      data$period_index
    ),


    # --------------------------------------------------------
    # Geography
    # --------------------------------------------------------

    state_code = as.character(
      data$state_code_std
    ),

    district_code_raw = as.character(
      data$district_code_std
    ),

    district_id_raw = as.character(
      data$district_id_raw
    ),

    district_name = as.character(
      data$district_name
    ),

    district_name_key = as.character(
      data$district_name_key
    ),

    district_id = as.character(
      stable_district_id
    ),


    # --------------------------------------------------------
    # Controls
    # --------------------------------------------------------

    age = as.numeric(
      clean_code(
        data[[age_variable]]
      )
    ),

    education_level = as.numeric(
      clean_code(
        data[[education_variable]]
      )
    ),


    # --------------------------------------------------------
    # PLFS survey weight
    # --------------------------------------------------------

    survey_weight = as.numeric(
      clean_code(
        data[[weight_variable]]
      )
    ),


    # --------------------------------------------------------
    # Urban/rural indicators
    # --------------------------------------------------------

    urban = as.integer(
      data$urban
    ),

    rural = as.integer(
      data$rural
    ),


    # --------------------------------------------------------
    # Metro treatment variables
    # --------------------------------------------------------

    has_metro = as.integer(
      data$has_metro
    ),

    metro_urban = as.integer(
      data$metro_urban
    ),

    metro_opened_this_quarter = as.integer(
      data$metro_opened_this_quarter
    ),

    metro_start_period_index = as.integer(
      data$metro_start_period_index
    ),

    metro_pre_sample = get_optional_integer(
      data,
      "metro_pre_sample"
    ),

    metro_system = get_optional_character(
      data,
      "metro_system"
    ),

    metro_opening_year = get_optional_integer(
      data,
      "metro_opening_year"
    ),

    metro_opening_quarter = get_optional_integer(
      data,
      "metro_opening_quarter"
    ),

    metro_manual_review = get_optional_integer(
      data,
      "metro_manual_review"
    ),


    # --------------------------------------------------------
    # Outcomes
    # --------------------------------------------------------

    transport = as.integer(
      data$transport
    ),

    informal_transport = as.integer(
      data$informal_transport
    ),

    informal = as.integer(
      data$informal
    ),

    informal_unknown = as.integer(
      data$informal_unknown
    )
  )


  # Confirm that no observations were removed
  if (nrow(compact_data) != nrow(data)) {
    stop(
      paste0(
        dataset_name,
        ": row count changed during harmonisation."
      )
    )
  }


  compact_data
}
```

```{r}
# ============================================================
# 4. Create compact annual DiD datasets
# ============================================================

did17_18 <- prepare_did_dataset(
  data = analysis17_18,
  dataset_name = "2017-18",

  age_candidates = c(
    "b4q6_per_fv"
  ),

  education_candidates = c(
    "b4q8_per_fv"
  ),

  weight_candidates = c(
    "MULT_per_fv",
    "mult_per_fv"
  )
)


did18_19 <- prepare_did_dataset(
  data = analysis18_19,
  dataset_name = "2018-19",

  age_candidates = c(
    "b4q6_per_fv"
  ),

  education_candidates = c(
    "b4q8_per_fv"
  ),

  weight_candidates = c(
    "MULT_per_fv",
    "mult_per_fv"
  )
)


did19_20 <- prepare_did_dataset(
  data = analysis19_20,
  dataset_name = "2019-20",

  age_candidates = c(
    "b4q6_per_fv"
  ),

  education_candidates = c(
    "b4q8_per_fv"
  ),

  weight_candidates = c(
    "MULT_per_fv",
    "mult_per_fv"
  )
)


did20_21 <- prepare_did_dataset(
  data = analysis20_21,
  dataset_name = "2020-21",

  age_candidates = c(
    "b4q6_perv1"
  ),

  education_candidates = c(
    "b4q8_perv1"
  ),

  weight_candidates = c(
    "mult_perv1",
    "MULT_perv1"
  )
)


did22 <- prepare_did_dataset(
  data = analysis22,
  dataset_name = "2022",

  age_candidates = c(
    "b4q6_cperv1",
    "b4q6_perv1"
  ),

  education_candidates = c(
    "b4q8_cperv1",
    "b4q8_perv1"
  ),

  weight_candidates = c(
    "mult_cperv1",
    "mult_perv1"
  )
)


did23 <- prepare_did_dataset(
  data = analysis23,
  dataset_name = "2023",

  age_candidates = c(
    "b4q6_cperv1",
    "b4q6_perv1"
  ),

  education_candidates = c(
    "b4q8_cperv1",
    "b4q8_perv1"
  ),

  weight_candidates = c(
    "mult_cperv1",
    "mult_perv1"
  )
)


did24 <- prepare_did_dataset(
  data = analysis24,
  dataset_name = "2024",

  age_candidates = c(
    "Age"
  ),

  education_candidates = c(
    "General_Education_Level"
  ),

  weight_candidates = c(
    "Subsample_Multiplier"
  )
)


did25 <- prepare_did_dataset(
  data = analysis25,
  dataset_name = "2025",

  age_candidates = c(
    "age"
  ),

  education_candidates = c(
    "gedu_lvl"
  ),

  weight_candidates = c(
    "mult"
  )
)
```

```{r}
# ============================================================
# 5. Stack the annual datasets
# ============================================================

did_data <- bind_rows(
  did17_18,
  did18_19,
  did19_20,
  did20_21,
  did22,
  did23,
  did24,
  did25
) %>%
  arrange(
    period_index,
    district_id
  )
```

```{r}
# ============================================================
# 6. Add DiD sample flags
# ============================================================

first_sample_period <- min(
  did_data$period_index,
  na.rm = TRUE
)


did_data <- did_data %>%
  mutate(
    # Controls are available
    complete_controls = as.integer(
      !is.na(age) &
        !is.na(education_level)
    ),

    # Valid positive survey weight
    valid_survey_weight = as.integer(
      !is.na(survey_weight) &
        survey_weight > 0
    ),

    # Informality was classified with sufficient information
    informal_known = as.integer(
      informal_unknown == 0
    ),

    # Treatment status during the available sample
    metro_treatment_group = case_when(
      is.na(metro_start_period_index) ~
        "Never treated",

      metro_start_period_index == 0L |
        metro_start_period_index <
        first_sample_period ~
        "Metro before sample",

      TRUE ~
        "Metro opened during sample"
    ),

    # Quarters relative to metro opening.
    # Only meaningful for metros opening during the sample.
    event_time = case_when(
      metro_treatment_group ==
        "Metro opened during sample" ~

        period_index -
        metro_start_period_index,

      TRUE ~
        NA_integer_
    ),

    # Complete observation for a baseline model
    baseline_sample = as.integer(
      complete_controls == 1L &
        valid_survey_weight == 1L &
        !is.na(district_id) &
        !is.na(period_index)
    ),

    # Complete observation for informality outcomes
    informal_sample = as.integer(
      baseline_sample == 1L &
        informal_known == 1L
    )
  )
```

```{r}
# ============================================================
# 7. Validate the combined dataset
# ============================================================

expected_rows <- sum(
  nrow(analysis17_18),
  nrow(analysis18_19),
  nrow(analysis19_20),
  nrow(analysis20_21),
  nrow(analysis22),
  nrow(analysis23),
  nrow(analysis24),
  nrow(analysis25)
)


actual_rows <- nrow(
  did_data
)


cat(
  "Expected rows:",
  expected_rows,
  "\n"
)

cat(
  "Actual rows:",
  actual_rows,
  "\n"
)


stopifnot(
  expected_rows == actual_rows
)
```


